### 2014-12-20 paper/ms.tex
Altered the directory structure

\documentclass[apj]{emulateapj}

\usepackage{amsmath}
\usepackage{aas_macros}
\usepackage{natbib}
\usepackage{rotating}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{color}
%% \usepackage[section]{placeins}
%% \usepackage{caption}

%%% Definitions added by Manodeep and Andreas
\newcommand{\Msun}{M_{\odot}}
\newcommand{\hMsun}{h^{-1}M_{\odot}}
\newcommand{\Mvir}{M_\mathrm{vir}}
\newcommand{\Rvir}{\ensuremath{R_\mathrm{vir}}}
\newcommand{\Vvir}{V_\mathrm{vir}}
\newcommand{\rhoc}{\rho_{\rm crit}}
\newcommand{\dd}{{\rm {d}}}
\newcommand{\Ob}{\Omega_{\rm b}}
\newcommand{\Ol}{\Omega_\Lambda}
\newcommand{\Odm}{\Omega_{\rm dm}}
\newcommand{\Om}{\Omega_{\rm m}}
\newcommand{\sig}{{\sigma_8}}
\newcommand{\FOF}{{\small{FOF}}\xspace}
\newcommand{\FOFS}{{\small{FOF}}s\xspace}
\newcommand{\sdss}{{\small{SDSS}}\xspace}
\newcommand{\hod}{{\small{HOD}}\xspace}
\newcommand{\ngal}{\ensuremath{n_{gal}}\xspace}
\newcommand{\wprp}{\ensuremath{w_p(r_p)}\xspace}
\newcommand{\gmf}{\ensuremath{n(N)}\xspace}
\newcommand{\nsatavg}{\ensuremath{\langle N_{sat} \rangle}}
\newcommand{\Mmin}{\ensuremath{M_\mathrm{min}}}
\newcommand{\erf}{{\ensuremath{\mathrm{erf}}\xspace}}
\newcommand{\logMmin}{\ensuremath{\mathrm{\log M_{min}}}\xspace}
\newcommand{\siglogM}{\ensuremath{\sigma_{\log \mathrm{M}}}\xspace}
\newcommand{\siglogL}{\ensuremath{\sigma_{\log \mathrm{L}}}\xspace}
\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}\xspace}}
\newcommand{\hkpc}{\ensuremath{{h^{-1}kpc}\xspace}}
\newcommand*{\bigchi}{{\Large{\ensuremath{\chi}}}}
\newcommand*{\sizeforstatplots}{0.48}
\newcommand*{\sizeforhodparamplots}{0.49}
\newcommand{\emcee}{\texttt emcee}
\newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}

%%\setstretch{2.0}
%% \usepackage{etoolbox}
%% \makeatletter
%% \patchcmd{\@makecaption}
%%   {\scshape}
%%   {}
%%   {}
%%   {}
%% \makeatother

\begin{document}

\title{Cache is King: Writing a Blazing-fast Correlation Function Code}

\author{Manodeep Sinha\altaffilmark{1}}
\affiliation{Department of Physics and Astronomy, Vanderbilt University, Nashville, TN, 37235}
\altaffiltext{1}{manodeep.sinha@vanderbilt.edu}

\begin{abstract}
We have entered the era of `Big Data' where we are gathering a tremendous amount of data that
needs to be processed as well as modeled accurately. In particular, large galaxy surveys
like the Sloan Digitial Sky Survey requires computing correlation functions. While measuring the
correlation function is not a bottle-neck, modeling the observed galaxy distribution correctly
{\em} requires a MCMC chain and repeated measurements of $\xi(r)$ and/or $\xi(r_p,\pi)$.

The accompanying codes are publicly available at \url{https://bitbucket.org/manodeep/Corrfunc}. 
\end{abstract}

\keywords{methods: numerical}

\maketitle

\section{INTRODUCTION}

%% MS: I have not encountered this in the past -- but the footnotecounter needs to be reset. Otherwise, 
%% the numbering starts at Nauthors+1. Presumably \altaffiltext uses \footnote internally and does not
%% reset the counter like a responsible person!
\setcounter{footnote}{0}

\acknowledgments MS acknowledges the fantastic \texttt{largemem} nodes with 1 TB of RAM at Stampede (TACC) without which the project could not have been done. 
This work has been supported by the NSF Career Award (AST-1151650).

\bibliographystyle{apj}
\bibliography{master}

\begin{appendix}

\end{document}

### 2014-12-28 xi_theory/include/avx_calls.h
Created the code directory structure

@@ -1,0 +1,124 @@
+#pragma once
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <immintrin.h>
+
+
+#define PREFETCH(mem)        asm ("prefetcht0 %0"::"m"(mem))
+
+#if defined(__GNUC__) || defined(__GNUG__)
+#define AVX_BIT_COUNT_INT(X)                __builtin_popcount(X)
+#else
+#define AVX_BIT_COUNT_INT(X)                _popcnt32(X)
+#endif
+
+
+#define AVX_BIT_COUNT_LONG(X)               _popcnt64(X)
+#define AVX_BIT_COUNT_UNSIGNED_INT(X)       _mm_popcnt_u32(X)
+#define AVX_BIT_COUNT_UNSIGNED_LONG(X)      _mm_popcnt_u64(X)
+#define AVX_SET_INT(X)                      _mm256_set1_epi32(X)
+
+
+#ifndef DOUBLE_PREC
+
+#define AVX_INTS                         __m256i
+#define AVX_FLOATS                       __m256
+
+#define AVX_LOAD_FLOATS_UNALIGNED(X)     _mm256_loadu_ps(X)
+#define AVX_LOAD_FLOATS_ALIGNED(X)       _mm256_load_ps(X)
+#define AVX_MULTIPLY_FLOATS(X,Y)         _mm256_mul_ps(X,Y)
+#define AVX_DIVIDE_FLOATS(X,Y)           _mm256_div_ps(X,Y)
+#define AVX_SUBTRACT_FLOATS(X,Y)         _mm256_sub_ps(X,Y)
+#define AVX_ADD_FLOATS(X,Y)              _mm256_add_ps(X,Y)
+#define AVX_SQRT_FLOAT(X)                _mm256_sqrt_ps(X)
+#define AVX_SVML_SQRT_FLOAT(X)           _mm256_svml_sqrt_ps(X)
+#define AVX_TRUNCATE_FLOAT_TO_INT(X)     _mm256_cvttps_epi32(X)
+#define AVX_STORE_FLOATS_TO_MEMORY(X,Y)  _mm256_storeu_ps(X,Y)
+#define AVX_SQUARE_FLOAT(X)              _mm256_mul_ps(X,X)
+#define AVX_LOG_FLOAT(X)                 _mm256_log_ps(X)
+#define AVX_LOG10_FLOAT(X)                 _mm256_log10_ps(X)
[…]

### 2014-12-28 xi_theory/include/cellarray.h

@@ -1,0 +1,17 @@
+#pragma once
+
+#include <stdio.h>
+#include <stdlib.h>
+
+#include "function_precision.h" //definition of DOUBLE depending on Makefile flag -DOUBLE_PREC
+
+#define NLATMAX   100      /* maximum grid dimension in X-Y plane */
+
+typedef struct{
+  DOUBLE *x;
+  DOUBLE *y;
+  DOUBLE *z;
+  int nelements;
+} cellarray;
+
+

### 2014-12-28 xi_theory/io/ftread.h

@@ -1,0 +1,10 @@
+#ifndef __FTREAD_H
+#define __FTREAD_H
+#include <stdio.h>
+#include <stdlib.h>
+
+
+int ftread(void *ptr,size_t size,size_t nitems,FILE *stream);
+int my_ftread(void *ptr,size_t size,size_t nitems,FILE *stream);
+
+#endif

### 2014-12-28 xi_theory/logbins.c

@@ -1,0 +1,52 @@
+/* PROGRAM logbins
+
+   --- logbins xmin xmax nbin
+   --- Outputs the bins for a logarithmic bin spacing
+
+      * xmin   = inner radius of smallest bin
+      * xmax   = outer radius of largest bin
+      * nbin  = number of bins (logarithmically spaced in theta)
+      ----------------------------------------------------------------------
+*/
+
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+#define sqr(x) ((x)*(x))
+#define max(A,B) ((A) > (B) ? (A) : (B))
+#define min(A,B) ((A) < (B) ? (A) : (B))
+#define mabs(A) ((A) < 0.0 ? -(A) : (A))
+#define pi 3.1415926535898
+
+int main(int argc, char *argv[])
+{
+  /*---Arguments-------------------------*/
+  int nbin ;
+  double xmin,xmax,logxmin,logxmax,dlogx;
+
+  if(argc<4) {
+    fprintf(stderr,"Usage `%s' rmin rmax nbins \n",argv[0]);
+    exit(EXIT_FAILURE);
+  } else {
+    /*---Read-arguments-----------------------------------*/
+    sscanf(argv[1],"%lf",&xmin) ;
+    sscanf(argv[2],"%lf",&xmax) ;
+    sscanf(argv[3],"%d",&nbin) ;
+  }
+  /*---Setup-Pairs-arrays-------------------------------*/
+  logxmin = log10(xmin) ;
+  logxmax = log10(xmax) ;
+  dlogx = (logxmax-logxmin)/(double)nbin ;
+
+  double left_edge=logxmin;
+  for(int i=0;i<nbin;i++){
+    double right_edge=left_edge+dlogx; 
+    fprintf(stdout,"%20.12lf %20.12lf\n",pow(10.0,left_edge),pow(10.0,right_edge)) ;
+    left_edge=right_edge;
+  }
+  exit(EXIT_SUCCESS);
+}
+
+
+
+

### 2014-12-28 paper/ms.tex
Created the code directory structure

@@
 like the Sloan Digitial Sky Survey requires computing correlation functions. While measuring the
-correlation function is not a bottle-neck, modeling the observed galaxy distribution correctly
+correlation function in the data is not a bottle-neck, modeling the observed galaxy distribution correctly
 {\em} requires a MCMC chain and repeated measurements of $\xi(r)$ and/or $\xi(r_p,\pi)$.
@@
 
-\acknowledgments MS acknowledges the fantastic \texttt{largemem} nodes with 1 TB of RAM at Stampede (TACC) without which the project could not have been done. 
-This work has been supported by the NSF Career Award (AST-1151650).
+\section{}
 
@@
 
-\begin{appendix}
-
 \end{document}

### 2014-12-28 xi_theory/io/io.c

@@ -1,0 +1,160 @@
+#include "io.h"
+#include "ftread.h"
+#include "utils.h"
+
+#ifndef MEMORY_INCREASE_FAC
+#define MEMORY_INCREASE_FAC 1.1
+#endif
+
+unsigned int read_positions(const char *filename, const char *format, void **xpos, void **ypos, void **zpos, const size_t size)
+{
+	unsigned int np;
+	void *x,*y,*z;
+	if(strncmp(format,"f",1)==0) {
+		//read fast-food file
+		size_t bytes=sizeof(int) + sizeof(float)*9 + sizeof(int);//skip fdat
+		bytes += sizeof(int) + sizeof(float) + sizeof(int); //skip znow
+		int idat[5];
+		FILE *fp = my_fopen(filename,"r");
+		my_ftread(idat,sizeof(int),5,fp);
+		np = idat[1];
+
+		assert((size == 4 || size == 8) && "Size of each position element can be either 4 (float) or 8 (double)");
+		//TO DO -- cover the case where the file is written in the other precision. 
+
+		x = my_malloc(size,np);
+		y = my_malloc(size,np);
+		z = my_malloc(size,np);
+
+		my_fseek(fp,bytes,SEEK_CUR);
+		//Check that the file was written with the requested precision
+		unsigned int dummy;
+		my_fread(&dummy,sizeof(dummy), 1, fp);
+		//so rewind by 4 bytes  prepare for calls to ftread
+		my_fseek(fp, -sizeof(dummy), SEEK_CUR);
+		dummy /= np;
+		assert((dummy == 4 || dummy == 8) && "File must contain either 4 byte (float) or 8 byte(double) precision");
+		
+		if(dummy == size) {
+			my_ftread(x,size, np, fp);
+			my_ftread(y,size, np, fp);
+			my_ftread(z,size, np, fp);
+		} else {
+			fprintf(stderr,"WARNING: File was written in a different precision than requested (file precision = %u requested precision = %zu)\n",dummy,size);
+			//Okay so the file was written in a different precision.
+			//First, print a warning message and then read-in correctly with the
+			//requested precision
+			if(dummy == 4) {
[…]

### 2014-12-28 xi_theory/utils/gridlink.c

@@ -1,0 +1,144 @@
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+
+#include "defs.h"
+#include "gridlink.h"
+#include "utils.h"
+#include "function_precision.h"
+#include "cellarray.h"
+
+#define MEMORY_INCREASE_FAC   1.2
+
+double get_binsize(const double xmin,const double xmax, const double rmax, const int refine_factor, const int max_ncells, int *nlattice)  __attribute__((warn_unused_result));
+
+double get_binsize(const double xmin,const double xmax, const double rmax, const int refine_factor, const int max_ncells, int *nlattice)
+{
+  double xdiff = xmax-xmin;
+  int nmesh=(int)(refine_factor*xdiff/rmax) ;
+#ifdef PERIODIC
+  if (nmesh<(2*refine_factor+1))  {
+    fprintf(stderr,"linklist> ERROR:  nlattice = %d is so small that with periodic wrapping the same cells will be counted twice ....exiting\n",nmesh) ;
+    exit(EXIT_FAILURE) ;
+  }
+#endif
+  
+  if (nmesh>max_ncells)  nmesh=max_ncells;
+  double xbinsize = xdiff/nmesh;
+  *nlattice = nmesh;
+  return xbinsize;
+}
+
+cellarray * gridlink(const int np,
+					 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
+					 const DOUBLE xmin, const DOUBLE xmax,
+					 const DOUBLE ymin, const DOUBLE ymax,
+					 const DOUBLE zmin, const DOUBLE zmax,
+					 const DOUBLE rmax,
+					 const int xbin_refine_factor,
+					 const int ybin_refine_factor,
+					 const int zbin_refine_factor,
+					 int *nlattice_x,
+					 int *nlattice_y,
+					 int *nlattice_z)
+{
+  cellarray *lattice=NULL;
+  int ix,iy,iz;
+  int nmesh_x,nmesh_y,nmesh_z;
+  int *nallocated=NULL;
+  DOUBLE xdiff,ydiff,zdiff;
+  DOUBLE cell_volume,box_volume;
+  DOUBLE xbinsize,ybinsize,zbinsize;
+  int expected_n=0;
+  int64_t totncells;
[…]

### 2014-12-28 xi_theory/utils/utils.c

@@ -1,0 +1,702 @@
+/*
+A collection of C wrappers I use. Should be 
+very obvious. The ones that are not obvious
+have comments before the function itself. 
+
+Bugs:
+Please email me manodeep at gmail dot com
+
+Ver 1.0: Manodeep Sinha, 2nd April, 2012
+Ver 1.1: Manodeep Sinha, 14th June, 2012 - replaced 
+         check_string_copy with a "real" wrapper to
+		 snprintf.
+Ver 1.2: Manodeep Sinha, Jan 8, 2012 - replaced 
+         print_time with timeval and gettimeofday
+*/
+
+#include "utils.h"
+/* #define __USE_XOPEN2K */
+/* #define _XOPEN_SOURCE_EXTENDED */
+/* #define _GNU_SOURCE  */
+
+
+void  setup_bins(const char *fname,double *rmin,double *rmax,int *nbin,double **rupp)
+{
+  //set up the bins according to the binned data file
+  //the form of the data file should be <rlow  rhigh ....>
+  const int MAXBUFSIZE=1000;
+  char buf[MAXBUFSIZE];
+  FILE *fp=NULL;
+  double low,hi;
+  const char comment='#';
+  const int nitems=2;
+  int nread=0;
+  *nbin = ((int) getnumlines(fname,comment))+1;
+  *rupp = my_calloc(sizeof(double),*nbin+1);
+  
+  fp = my_fopen(fname,"r");
+  int index=1;
+  while(1) {
+    if(fgets(buf,MAXBUFSIZE,fp)!=NULL) {
+      nread=sscanf(buf,"%lf %lf",&low,&hi);
+      if(nread==nitems) {
+
+	if(index==1) {
+	  *rmin=low;
+	  (*rupp)[0]=low;
+	} 
+
+	(*rupp)[index] = hi;
+	index++;
+      }
+    } else {
+      break;
+    }
+  }
+  *rmax = (*rupp)[index-1];
+  fclose(fp);
+
+  (*rupp)[*nbin]=*rmax ;
+  (*rupp)[*nbin-1]=*rmax ;
+}  
+
+
+void  setup_bin_lookup(const char *fname,double *rmin,double *rmax,int *nbin,const size_t nbinlookup,double **rupp,int *binlookup)
+{
+  //set up the bins according to the binned data file
+  //the form of the data file should be 
+  const int MAXBUFSIZE=1000;
+  char buf[MAXBUFSIZE];
+  FILE *fp=NULL;
[…]

### 2014-12-28 xi_theory/xi_rp_pi/DDrppi.c

@@ -1,0 +1,222 @@
+/* PROGRAM DDrppi
+
+   --- DDrppi rpmin rpmax nrpbin data1 data2 [pimax] > wpfile
+   --- Measure the cross-correlation function xi(rp,pi) for two different
+       data files (or autocorrelation if data1=data2).
+
+      * rpmin   = inner radius of smallest bin (in Mpc/h)
+      * rpmax   = outer radius of largest bin
+      * nrpbin  = number of bins (logarithmically spaced in r)
+      * data1   = name of first data file
+      * data2   = name of second data file
+      *[pimax]  = maximum value of line-of-sight separation (default=40 Mpc/h)
+      > DDfile  = name of output file <logrp log(<rp>) pi pairs>
+      ----------------------------------------------------------------------
+*/
+
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+#include <assert.h>
+
+#include "../utils/utils.h"
+#include "../utils/io.h"
+#include "countpairs.h"
+
+int main(int argc, char *argv[])
+{
+
+  /*---Arguments-------------------------*/
+  int nrpbin;
+  char *file1=NULL,*format1=NULL;
+  char *file2=NULL,*format2=NULL;
+  DOUBLE pimax ;
+
+  /*---Data-variables--------------------*/
+  int ND1=0,ND2=0;
+
+  DOUBLE *x1=NULL,*y1=NULL,*z1=NULL;
+  DOUBLE *x2=NULL,*y2=NULL,*z2=NULL;//will point to x1/y1/z1 in case of auto-corr
+
+
+  /*---Corrfunc-variables----------------*/
+#ifndef USE_OMP
+	const char argnames[][30]={"file1","format1","file2","format2","binfile","pimax"};
+#else
+	int nthreads=2;
+	const char argnames[][30]={"file1","format1","file2","format2","binfile","pimax","Nthreads"};
+#endif
+  int nargs=sizeof(argnames)/(sizeof(char)*30);
+  
+  struct timeval t_end,t_start,t0,t1;
+  double read_time=0.0;
+  gettimeofday(&t_start,NULL);
+  
+  /*---Read-arguments-----------------------------------*/
+  if(argc< (nargs+1)) {
[…]

### 2014-12-28 xi_theory/xi_rp_pi/wprp.c

@@ -1,0 +1,219 @@
+/* PROGRAM wprp
+
+   --- wprp nrbin ND1 ND2 NR1 NR2 D1D2 D1R2 D2R1 R1R2 [pimax] > wpfile
+   --- Measure the cross-correlation function wp(rp) for two different
+       data and random files (or autocorrelation if D1=D2 and R1=R2).
+       Uses the Szalay estimator: xi(rp,pi) = (D1D2 - D1R2 - D2R1 - R1R2)/RR2
+
+      * nrpbin = number of bins (logarithmically spaced in r)
+      * ND1    = number of data points in data1
+      * ND2    = number of data points in data2
+      * NR1    = number of random points in random1
+      * NR2    = number of random points in random2
+      * D1D2   = data1-data2 pairs (output of corrfunc.c)
+      * D1R2   = data1-random2 pairs (output of corrfunc.c)
+      * D2R1   = data2-random1 pairs (output of corrfunc.c)
+      * R1R2   = random1-random2 pairs (output of corrfunc.c)
+      *[pimax] = maximum pi in integral to compute wp(rp) (default = 40 Mpc/h) 
+      > wpfile = name of output file <logrp rpavg wp(rp)>
+      ----------------------------------------------------------------------
+*/
+
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+
+#include "utils.h"
+void Printhelp(void);
+
+int main(int argc, char *argv[])
+{
+  int i,j;
+/*---Arguments-------------------------*/
+  int nrpbin,ND1,ND2,NR1,NR2 ;
+  double pimax ;
+  FILE *fpD1D2,*fpD1R2,*fpD2R1,*fpR1R2 ;
+/*---Data-variables--------------------*/
+  int ndat ;
+  double *logrp1,*ravg1,*pi1,*D1D2,*D1R2,*D2R1,*R1R2 ;
+  /*---Corrfunc-variables----------------*/
+  int npibin,indx,npimax ;
+  double dpi ;
+  double fN1,fN2,*xirppi,*wp,*logrp,*rpavg,*DDtot ;
+
+
+
[…]

### 2015-01-06 xi_theory/io/io.c
Changed nelements to int64_t. Made the loop variables same across all three correlation functions. Added some colored output to examples code. Python bindings is still not functional

@@ -6,58 +6,58 @@
 #define MEMORY_INCREASE_FAC 1.1
 #endif
 
-unsigned int read_positions(const char *filename, const char *format, void **xpos, void **ypos, void **zpos, const size_t size)
+int64_t read_positions(const char *filename, const char *format, void **xpos, void **ypos, void **zpos, const size_t size)
 {
-	unsigned int np;
-	void *x,*y,*z;
-	if(strncmp(format,"f",1)==0) {
-		//read fast-food file
-		size_t bytes=sizeof(int) + sizeof(float)*9 + sizeof(int);//skip fdat
-		bytes += sizeof(int) + sizeof(float) + sizeof(int); //skip znow
-		int idat[5];
-		FILE *fp = my_fopen(filename,"r");
-		my_ftread(idat,sizeof(int),5,fp);
-		np = idat[1];
-
-		assert((size == 4 || size == 8) && "Size of each position element can be either 4 (float) or 8 (double)");
-		//TO DO -- cover the case where the file is written in the other precision. 
-
-		x = my_malloc(size,np);
-		y = my_malloc(size,np);
-		z = my_malloc(size,np);
-
-		my_fseek(fp,bytes,SEEK_CUR);
-		//Check that the file was written with the requested precision
-		unsigned int dummy;
-		my_fread(&dummy,sizeof(dummy), 1, fp);
-		//so rewind by 4 bytes  prepare for calls to ftread
-		my_fseek(fp, -sizeof(dummy), SEEK_CUR);
-		dummy /= np;
-		assert((dummy == 4 || dummy == 8) && "File must contain either 4 byte (float) or 8 byte(double) precision");
-		
-		if(dummy == size) {
-			my_ftread(x,size, np, fp);
-			my_ftread(y,size, np, fp);
-			my_ftread(z,size, np, fp);
-		} else {
-			fprintf(stderr,"WARNING: File was written in a different precision than requested (file precision = %u requested precision = %zu)\n",dummy,size);
-			//Okay so the file was written in a different precision.
-			//First, print a warning message and then read-in correctly with the
-			//requested precision
+  int64_t np;
+  void *x,*y,*z;
[…]

### 2015-01-06 xi_theory/python_bindings/call_correlation_functions.py

@@ -12,7 +12,6 @@ import numpy as np
 import _countpairs
 
 
-
 file="../tests/data/gals_Mr19.txt"
 binfile="../tests/bins"
 dtype='float' ### Make sure the precision agrees with the definition in ../common.mk
@@ -25,6 +24,7 @@ boxsize=np.max(array)
 nthreads=4
 pimax=40.0
 results = _countpairs.countpairs_wp(boxsize,pimax,nthreads,binfile,x,y,z)
+
 
 
 

### 2015-01-06 xi_theory/utils/gridlink.c

@@ -29,13 +29,13 @@ double get_binsize(const double xmin,const double xmax, const double rmax, const
   return xbinsize;
 }
 
-void get_max_min(const int ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
-								 DOUBLE *min_x, DOUBLE *min_y, DOUBLE *min_z, DOUBLE *max_x, DOUBLE *max_y, DOUBLE *max_z)
+void get_max_min(const int64_t ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
+				 DOUBLE *min_x, DOUBLE *min_y, DOUBLE *min_z, DOUBLE *max_x, DOUBLE *max_y, DOUBLE *max_z)
 {
-	DOUBLE xmin = *min_x, ymin = *min_y, zmin=*min_z;
-	DOUBLE xmax = *max_x, ymax = *max_y, zmax=*max_z;
+  DOUBLE xmin = *min_x, ymin = *min_y, zmin=*min_z;
+  DOUBLE xmax = *max_x, ymax = *max_y, zmax=*max_z;
 	
-  for(int i=0;i<ND1;i++) {
+  for(int64_t i=0;i<ND1;i++) {
     if(X1[i] < xmin) xmin=X1[i];
     if(Y1[i] < ymin) ymin=Y1[i];
     if(Z1[i] < zmin) zmin=Z1[i];
@@ -50,29 +50,29 @@ void get_max_min(const int ND1, const DOUBLE * restrict X1, const DOUBLE * restr
 }
 
 
-cellarray * gridlink(const int np,
-										 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
-										 const DOUBLE xmin, const DOUBLE xmax,
-										 const DOUBLE ymin, const DOUBLE ymax,
-										 const DOUBLE zmin, const DOUBLE zmax,
-										 const DOUBLE max_x_size,
-										 const DOUBLE max_y_size,
-										 const DOUBLE max_z_size,
-										 const int xbin_refine_factor,
-										 const int ybin_refine_factor,
-										 const int zbin_refine_factor,
-										 int *nlattice_x,
-										 int *nlattice_y,
-										 int *nlattice_z)
+cellarray * gridlink(const int64_t np,
[…]

### 2015-01-06 xi_theory/wp/countpairs_wp.h

@@ -15,13 +15,13 @@ typedef struct{
 	int nbin;
 } results_countpairs_wp;
 
-results_countpairs_wp *countpairs_wp(const int ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
-																		 const double boxsize, 
+results_countpairs_wp *countpairs_wp(const int64_t ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
+									 const double boxsize, 
 #ifdef USE_OMP
-																		 const int numthreads,
+									 const int numthreads,
 #endif
-																		 const char *binfile,
-																		 const double pimax) __attribute__((warn_unused_result));
+									 const char *binfile,
+									 const double pimax) __attribute__((warn_unused_result));
 
 
 void free_results_wp(results_countpairs_wp **results);

### 2015-01-06 xi_theory/wp/wp.c

@@ -15,9 +15,10 @@
 */
 
 #include <stdio.h>
-#include <math.h>
 #include <stdlib.h>
+#include <math.h>
 #include <assert.h>
+#include <inttypes.h>
 
 #include "defs.h" //for ADD_DIFF_TIME
 #include "function_precision.h" //definition of DOUBLE
@@ -39,7 +40,7 @@ int main(int argc, char *argv[])
   DOUBLE pimax ;
 	
   /*---Data-variables--------------------*/
-  int ND1=0,ND2=0;
+  int64_t ND1=0,ND2=0;
 
   DOUBLE *x1=NULL,*y1=NULL,*z1=NULL;
 
@@ -158,7 +159,7 @@ int main(int argc, char *argv[])
 	free_results_wp(&results);
 	
   gettimeofday(&t_end,NULL);
-  fprintf(stderr,"wp> Done -  ND1=%d ND2=%d. Time taken = %6.2lf seconds. read-in time = %6.2lf seconds pair-counting time = %6.2lf sec\n",
+  fprintf(stderr,"wp> Done -  ND1=%12"PRId64" ND2=%12"PRId64". Time taken = %6.2lf seconds. read-in time = %6.2lf seconds pair-counting time = %6.2lf sec\n",
 	  ND1,ND2,ADD_DIFF_TIME(t_start,t_end),read_time,pair_time);
   return EXIT_SUCCESS;
 }

### 2015-01-06 xi_theory/xi_of_r/countpairs.c

@@ -38,13 +38,13 @@ void free_results(results_countpairs **results)
 	tmp = NULL;
 }
 
-results_countpairs * countpairs(const int ND1, const DOUBLE * const X1, const DOUBLE * const Y1, const DOUBLE  * const Z1,
-																const int ND2, const DOUBLE * const X2, const DOUBLE * const Y2, const DOUBLE  * const Z2,
+results_countpairs * countpairs(const int64_t ND1, const DOUBLE * const X1, const DOUBLE * const Y1, const DOUBLE  * const Z1,
+								const int64_t ND2, const DOUBLE * const X2, const DOUBLE * const Y2, const DOUBLE  * const Z2,
 #ifdef USE_OMP
-																const int numthreads,
+								const int numthreads,
 #endif
-																const int autocorr,
-																const char *binfile)
+								const int autocorr,
+								const char *binfile)
 {
 	
   int bin_refine_factor=1;
@@ -86,9 +86,9 @@ results_countpairs * countpairs(const int ND1, const DOUBLE * const X1, const DO
 	get_max_min(ND1, X1, Y1, Z1, &xmin, &ymin, &zmin, &xmax, &ymax, &zmax);
 	
 	if(autocorr==0) {
-		fprintf(stderr,"ND1 = %8d [xmin,ymin,zmin] = [%lf,%lf,%lf], [xmax,ymax,zmax] = [%lf,%lf,%lf]\n",ND1,xmin,ymin,zmin,xmax,ymax,zmax);
+		fprintf(stderr,"ND1 = %12"PRId64" [xmin,ymin,zmin] = [%lf,%lf,%lf], [xmax,ymax,zmax] = [%lf,%lf,%lf]\n",ND1,xmin,ymin,zmin,xmax,ymax,zmax);
 		get_max_min(ND2, X2, Y2, Z2, &xmin, &ymin, &zmin, &xmax, &ymax, &zmax);
-		fprintf(stderr,"ND2 = %8d [xmin,ymin,zmin] = [%lf,%lf,%lf], [xmax,ymax,zmax] = [%lf,%lf,%lf]\n",ND2,xmin,ymin,zmin,xmax,ymax,zmax);    
+		fprintf(stderr,"ND2 = %12"PRId64" [xmin,ymin,zmin] = [%lf,%lf,%lf], [xmax,ymax,zmax] = [%lf,%lf,%lf]\n",ND2,xmin,ymin,zmin,xmax,ymax,zmax);    
 	}
[…]

### 2015-01-06 xi_theory/xi_of_r/countpairs.h

@@ -14,13 +14,13 @@ typedef struct{
 	int nbin;
 } results_countpairs;
 
-results_countpairs * countpairs(const int ND1, const DOUBLE * const X1, const DOUBLE * const Y1, const DOUBLE  * const Z1,
-																const int ND2, const DOUBLE * const X2, const DOUBLE * const Y2, const DOUBLE  * const Z2,
+results_countpairs * countpairs(const int64_t ND1, const DOUBLE * const X1, const DOUBLE * const Y1, const DOUBLE  * const Z1,
+								const int64_t ND2, const DOUBLE * const X2, const DOUBLE * const Y2, const DOUBLE  * const Z2,
 #ifdef USE_OMP
-																const int numthreads,
+								const int numthreads,
 #endif
-																const int autocorr,
-																const char *binfile) __attribute__((warn_unused_result));
+								const int autocorr,
+								const char *binfile) __attribute__((warn_unused_result));
 
 
 void free_results(results_countpairs **results);

### 2015-01-06 xi_theory/xi_rp_pi/DDrppi.c

@@ -15,9 +15,10 @@
 */
 
 #include <stdio.h>
-#include <math.h>
 #include <stdlib.h>
+#include <math.h>
 #include <assert.h>
+#include <inttypes.h>
 
 #include "defs.h" //for ADD_DIFF_TIME
 #include "function_precision.h" //definition of DOUBLE
@@ -39,7 +40,7 @@ int main(int argc, char *argv[])
 	
 	
   /*---Data-variables--------------------*/
-  int ND1=0,ND2=0;
+  int64_t ND1=0,ND2=0;
 
   DOUBLE *x1=NULL,*y1=NULL,*z1=NULL;
   DOUBLE *x2=NULL,*y2=NULL,*z2=NULL;//will point to x1/y1/z1 in case of auto-corr
@@ -160,13 +161,13 @@ int main(int argc, char *argv[])
 
 	const DOUBLE dpi = pimax/(DOUBLE)results->npibin ;
 	for(int i=1;i<results->nbin;i++) {
-		const double logrp = LOG10(results->rupp[i]);
-    for(int j=0;j<npibin;j++) {
+	  const double logrp = LOG10(results->rupp[i]);
+	  for(int j=0;j<npibin;j++) {
       int index = i*(npibin+1) + j;
 #ifdef OUTPUT_RPAVG			
       fprintf(stdout,"%10"PRIu64" %20.8lf %20.8lf  %20.8lf \n",results->npairs[index],results->rpavg[index],logrp,(j+1)*dpi);
 #else			
-			fprintf(stdout,"%10"PRIu64" %20.8lf %20.8lf  %20.8lf \n",results->npairs[index],0.0,logrp,(j+1)*dpi);
+	  fprintf(stdout,"%10"PRIu64" %20.8lf %20.8lf  %20.8lf \n",results->npairs[index],0.0,logrp,(j+1)*dpi);
 #endif			
     }
   }
@@ -175,7 +176,7 @@ int main(int argc, char *argv[])
 	free_results_rp_pi(&results);
 	
   gettimeofday(&t_end,NULL);
-  fprintf(stderr,"xi_rp_pi> Done -  ND1=%d ND2=%d. Time taken = %6.2lf seconds. read-in time = %6.2lf seconds pair-counting time = %6.2lf sec\n",
[…]

### 2015-01-06 paper/ms.tex
Changed nelements to int64_t. Made the loop variables same across all three correlation functions. Added some colored output to examples code. Python bindings is still not functional

@@
 \usepackage{color}
-%% \usepackage[section]{placeins}
-%% \usepackage{caption}
 
-%%% Definitions added by Manodeep and Andreas
+%%% Definitions added by Manodeep 
 \newcommand{\Msun}{M_{\odot}}
 \newcommand{\hMsun}{h^{-1}M_{\odot}}
-\newcommand{\Mvir}{M_\mathrm{vir}}
-\newcommand{\Rvir}{\ensuremath{R_\mathrm{vir}}}
-\newcommand{\Vvir}{V_\mathrm{vir}}
-\newcommand{\rhoc}{\rho_{\rm crit}}
-\newcommand{\dd}{{\rm {d}}}
-\newcommand{\Ob}{\Omega_{\rm b}}
-\newcommand{\Ol}{\Omega_\Lambda}
-\newcommand{\Odm}{\Omega_{\rm dm}}
-\newcommand{\Om}{\Omega_{\rm m}}
-\newcommand{\sig}{{\sigma_8}}
-\newcommand{\FOF}{{\small{FOF}}\xspace}
-\newcommand{\FOFS}{{\small{FOF}}s\xspace}
 \newcommand{\sdss}{{\small{SDSS}}\xspace}
-\newcommand{\hod}{{\small{HOD}}\xspace}
-\newcommand{\ngal}{\ensuremath{n_{gal}}\xspace}
-\newcommand{\wprp}{\ensuremath{w_p(r_p)}\xspace}
-\newcommand{\gmf}{\ensuremath{n(N)}\xspace}
-\newcommand{\nsatavg}{\ensuremath{\langle N_{sat} \rangle}}
-\newcommand{\Mmin}{\ensuremath{M_\mathrm{min}}}
-\newcommand{\erf}{{\ensuremath{\mathrm{erf}}\xspace}}
-\newcommand{\logMmin}{\ensuremath{\mathrm{\log M_{min}}}\xspace}
-\newcommand{\siglogM}{\ensuremath{\sigma_{\log \mathrm{M}}}\xspace}
-\newcommand{\siglogL}{\ensuremath{\sigma_{\log \mathrm{L}}}\xspace}
 \newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}\xspace}}
 \newcommand{\hkpc}{\ensuremath{{h^{-1}kpc}\xspace}}
-\newcommand*{\bigchi}{{\Large{\ensuremath{\chi}}}}
-\newcommand*{\sizeforstatplots}{0.48}
-\newcommand*{\sizeforhodparamplots}{0.49}
 \newcommand{\emcee}{\texttt emcee}
@@
 
-\title{Cache is King: Writing a Blazing-fast Correlation Function Code}
+\title{Cache is King: Presenting a Suite of Blazing-Fast, Correlation Function Code}
 
@@
 correlation function in the data is not a bottle-neck, modeling the observed galaxy distribution correctly
-{\em} requires a MCMC chain and repeated measurements of $\xi(r)$ and/or $\xi(r_p,\pi)$.
+{\em} requires a MCMC chain and repeated measurements of $\xi(r)$ and/or $\xi(r_p,\pi)$. However, 
+the architecture of modern CPU's means that best performance is only obtained when cache misses 
+are kept to a minimum. Here, I present a suite of 3 correlation function codes that exploit 
+current CPU architecture and hand-written AVX
 
-The accompanying codes are publicly available at \url{https://bitbucket.org/manodeep/Corrfunc}. 
+The accompanying codes are publicly available at \url{https://bitbucket.org/manodeep/Corrfunc/}. 
 \end{abstract}
@@
 \section{INTRODUCTION}
-
 %% MS: I have not encountered this in the past -- but the footnotecounter needs to be reset. Otherwise, 
@@
 \setcounter{footnote}{0}
+The large-scale structure of the Universe can be now measured using $\mathcal{O}(million)$ galaxies 
+using data from current surveys like SDSS/BOSS surveys. Upcoming surveys like LSST will probe even 
+deeper and wider and aim to target on $\mathcal{O}(10's of million)$ galaxies. With such a large galaxy data, 
+we can measure the galaxy density field fairly accurately in the data. While the number of galaxies observed 
+in these current and upcoming surveys is large, we have to compute the galaxy density fields {\em only once}. Thus, 
+even a slow, correlation function code will be fine when measuring the data correlation function. 
+However, predicting this galaxy  density field data will require an even larger number of model galaxies -- increasing the computational 
+load of determining the density field even further. Typically, the modeling process also involves an MCMC -- 
+requiring $\sim$ millions of evaluations of the correlation function.\footnote{I will assume that creating the models themselves 
+is much faster compared to computing the correlation functions.}
+
+Modern cpu's have a hierarchy of memory locations; the smallest and fastest are closest to the cores while the largest and slowest are the
+farthest. All cpu instructions need to be carried out from cpu registers - there are $\sim 10-30$ registers typically available and the 
+access times can be thought of as instantaneous. Next up is the $L1$ cache divided into $L1D$ for data and $L1I$ for instructions cache. Since 
+the cpu always necessarily executes instructions that are close together, we will ignore the instruction cache from now on. Typical $L1$ cache 
+sizes range from 64KB to 128 KB. Next level up is the $L2$ cache, typically $\sim$ 256KB to 1 MB. The last level cache or the $L3$ cache 
+is usually shared across all cores on the socket and can be 10 MB to 40 MB.
+
+\section{Methods}
+We need to compute pairwise distances to get the correlation function. A naive implementation of a correlation function would compute {\em all possible} 
+pairwise separations with a complexity $\mathcal{O}(N^2)$. However, for almost all correlation functions, we are only interested in separations less than 
+a certain \rmax, where \rmax is much smaller than the domain of the point distribution itself. We can then immediately see a way to prune pairs that can not
+{\em possibly} be within \rmax. If we impose a 3-d grid, with cell-size \rmax, then two points separated by more than one cell size (\rmax) in any one 
+dimension can not be within \rmax of each other. Thus, given one point which is the target galaxy and a grid with cell-size \rmax, 
+immediately allows us to prune {\em all} of the points that are not within 1 cell offset in each dimension. However, even with this pruning, the actual 
+implementation of the algorithm matters. For instance, a linked-list in each cell performs $\sim 60x$ worse than the algorithm described here.\footnote{The 
+usual linked-list is very cache-unfriendly. Each dereference requires a read from a new region of memory and an almost guaranteed cache miss. }
+
+\subsection{How to Maintain Cache Locality within the Grid}
+For all pairs around a given target galaxy, we need to compute distances to all points within all neighbouring 3-d cells. 
+We ensure that the particle locations are contiguous by moving them into the following \texttt{C struct} in the order in which they arrive.
+
+\begin{code}
+struct cell
+{
+  DOUBLE *x;
+  DOUBLE *y;
+  DOUBLE *z;
+  int64_t nelements;
+}
+\end{code}\footnote{Most of the code assumes that the number of galaxies is within INT_MAX $\sim$ 2.1 billion. So, it would appear 
+to make no sense that the field inside the struct cell has `nelements' of type int64_t. The reason I have 
+chosen to use int64_t is that under C alignment rules 4 bytes of padding would have appended anyway to the struct cell. So, I chose 
+to declare `nelements' as an int64_t and used those 4 bytes. Saves some future conversion/coding. }
+
+\subsection{Hand-written Vectorization Support}
+Advanced Vector Extensions (AVX) has been available in cpu's more recent than 2011. AVX allows the processing of 8 floats or 4 double simultaneously -- thus, 
+potentially increasing the throughput by a factor of 8x/4x. However, automatic vectorization is not always possible by the compiler and in those cases 
+we can write AVX vector intrinsics to directly manipulate 8 floats/4 doubles.
+
+\subsection{Comparison to Tree Algorithms}
+
+\section{Results}
+
+\section{Benchmarks \& Scaling}
+
+\section{Conclusions}
+
+\acknowledgements
 

### 2015-01-06 xi_theory/xi_of_r/DD.c

@@ -2,6 +2,7 @@
 #include <math.h>
 #include <stdlib.h>
 #include <assert.h>
+#include <inttypes.h>
 
 #include "defs.h" //for ADD_DIFF_TIME
 #include "function_precision.h" //definition of DOUBLE
@@ -15,7 +16,7 @@ int main(int argc, char *argv[])
 {
 
   /*---Data-variables--------------------*/
-  int ND1,ND2 ;
+  int64_t ND1,ND2 ;
 
   DOUBLE *x1=NULL,*y1=NULL,*z1=NULL;
   DOUBLE *x2=NULL,*y2=NULL,*z2=NULL;
@@ -142,7 +143,7 @@ int main(int argc, char *argv[])
 	//free the memory in the results structx
 	free_results(&results);
   gettimeofday(&t_end,NULL);
-  fprintf(stderr,"xi> Done -  ND1=%d ND2=%d. Time taken = %6.2lf seconds. read-in time = %6.2lf seconds sec pair-counting time = %6.2lf sec\n",
+  fprintf(stderr,"xi> Done -  ND1=%"PRId64" ND2=%"PRId64". Time taken = %6.2lf seconds. read-in time = %6.2lf seconds sec pair-counting time = %6.2lf sec\n",
 		  ND1,ND2,ADD_DIFF_TIME(t_start,t_end),read_time,pair_time);
   return EXIT_SUCCESS;
 }

### 2015-01-06 xi_theory/python_bindings/_countpairs.c
Changed nelements to int64_t. Made the loop variables same across all three correlation functions. Added some colored output to examples code. Python bindings is still not functional

@@ -94,8 +94,8 @@ static PyObject *countpairs_countpairs(PyObject *self, PyObject *args)
 
 	
   /* How many data points are there? */
-	const int ND1 = (int)PyArray_DIM(x1_array, 0);
-	const int ND2 = (int)PyArray_DIM(x2_array, 0);
+	const int64_t ND1 = (int64_t)PyArray_DIM(x1_array, 0);
+	const int64_t ND2 = (int64_t)PyArray_DIM(x2_array, 0);
 
 	/* Get pointers to the data as C-types. */
 	DOUBLE *x1 = (DOUBLE *)PyArray_DATA(x1_array);

### 2015-01-06 xi_theory/utils/gridlink.h

@@ -3,20 +3,20 @@
 #include "function_precision.h"
 #include "cellarray.h"
 
-cellarray * gridlink(const int np,
-										 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
-										 const DOUBLE xmin, const DOUBLE xmax,
-										 const DOUBLE ymin, const DOUBLE ymax,
-										 const DOUBLE zmin, const DOUBLE zmax,
-										 const DOUBLE max_x_size,
-										 const DOUBLE max_y_size,
-										 const DOUBLE max_z_size,
-										 const int xbin_refine_factor,
-										 const int ybin_refine_factor,
-										 const int zbin_refine_factor,
-										 int *nlattice_x,
-										 int *nlattice_y,
-										 int *nlattice_z) __attribute__((warn_unused_result)); 
+cellarray * gridlink(const int64_t np,
+					 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
+					 const DOUBLE xmin, const DOUBLE xmax,
+					 const DOUBLE ymin, const DOUBLE ymax,
+					 const DOUBLE zmin, const DOUBLE zmax,
+					 const DOUBLE max_x_size,
+					 const DOUBLE max_y_size,
+					 const DOUBLE max_z_size,
+					 const int xbin_refine_factor,
+					 const int ybin_refine_factor,
+					 const int zbin_refine_factor,
+					 int *nlattice_x,
+					 int *nlattice_y,
+					 int *nlattice_z) __attribute__((warn_unused_result)); 
 
-void get_max_min(const int ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
-								 DOUBLE *min_x, DOUBLE *min_y, DOUBLE *min_z, DOUBLE *max_x, DOUBLE *max_y, DOUBLE *max_z);
+void get_max_min(const int64_t ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
+				 DOUBLE *min_x, DOUBLE *min_y, DOUBLE *min_z, DOUBLE *max_x, DOUBLE *max_y, DOUBLE *max_z);

### 2015-01-06 xi_theory/wp/countpairs_wp.c

@@ -45,7 +45,7 @@ void free_results_wp(results_countpairs_wp **results)
 }
 
 
-results_countpairs_wp *countpairs_wp(const int ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
+results_countpairs_wp *countpairs_wp(const int64_t ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
 																		 const double boxsize, 
 #ifdef USE_OMP
 																		 const int numthreads,
@@ -207,16 +207,16 @@ results_countpairs_wp *countpairs_wp(const int ND1, const DOUBLE * restrict X1, 
 						const DOUBLE *y = first->y;
 						const DOUBLE *z = first->z;
 						
-						for(int k=0;k<first->nelements;k++) {
-							const DOUBLE x1pos = x[k] + off_xwrap;
-							const DOUBLE y1pos = y[k] + off_ywrap;
-							const DOUBLE z1pos = z[k] + off_zwrap;
+						for(int64_t i=0;i<first->nelements;i++) {
+							const DOUBLE x1pos = x[i] + off_xwrap;
+							const DOUBLE y1pos = y[i] + off_ywrap;
+							const DOUBLE z1pos = z[i] + off_zwrap;
 
 #ifndef USE_AVX							
-							for(int i=0;i<second->nelements;i++) {
-								const DOUBLE dx = x2[i]-x1pos;
-								const DOUBLE dy = y2[i]-y1pos;
-								const DOUBLE dz = z2[i]-z1pos;
+							for(int64_t j=0;j<second->nelements;j++) {
+								const DOUBLE dx = x2[j]-x1pos;
+								const DOUBLE dy = y2[j]-y1pos;
+								const DOUBLE dz = z2[j]-z1pos;
 								if(dz < 0 ) {
 									continue;
 								} else if(dz >= pimax) {
@@ -261,18 +261,18 @@ results_countpairs_wp *countpairs_wp(const int ND1, const DOUBLE * restrict X1, 
 							union float8 union_mDperp;
 #endif							
 							
-							int i;
[…]

### 2015-01-06 xi_theory/xi_rp_pi/countpairs_rp_pi.c

@@ -37,14 +37,14 @@ void free_results_rp_pi(results_countpairs_rp_pi **results)
 }
 
 
-results_countpairs_rp_pi * countpairs_rp_pi(const int ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
-																						const int ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
+results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
+											const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
 #ifdef USE_OMP
-																						const int numthreads,
+											const int numthreads,
 #endif
-																						const int autocorr,
-																						const char *binfile,
-																						const double pimax)
+											const int autocorr,
+											const char *binfile,
+											const double pimax)
 {
   int bin_refine_factor=1;
 	int zbin_refine_factor=2;
@@ -84,9 +84,9 @@ results_countpairs_rp_pi * countpairs_rp_pi(const int ND1, const DOUBLE *X1, con
 	get_max_min(ND1, X1, Y1, Z1, &xmin, &ymin, &zmin, &xmax, &ymax, &zmax);
 
 	if(autocorr==0) {
-		fprintf(stderr,"ND1 = %8d [xmin,ymin,zmin] = [%lf,%lf,%lf], [xmax,ymax,zmax] = [%lf,%lf,%lf]\n",ND1,xmin,ymin,zmin,xmax,ymax,zmax);
+		fprintf(stderr,"ND1 = %12"PRId64" [xmin,ymin,zmin] = [%lf,%lf,%lf], [xmax,ymax,zmax] = [%lf,%lf,%lf]\n",ND1,xmin,ymin,zmin,xmax,ymax,zmax);
 		get_max_min(ND2, X2, Y2, Z2, &xmin, &ymin, &zmin, &xmax, &ymax, &zmax);
-		fprintf(stderr,"ND2 = %8d [xmin,ymin,zmin] = [%lf,%lf,%lf], [xmax,ymax,zmax] = [%lf,%lf,%lf]\n",ND2,xmin,ymin,zmin,xmax,ymax,zmax);
[…]

### 2015-01-13 paper/ms.tex
Trying to fix the dynamic path issue with the python library

@@
 \usepackage{color}
+\usepackage{listings}
 
@@
 %% \makeatother
+
+\lstset{
+  language=C,                % choose the language of the code
+  numbers=left,                   % where to put the line-numbers
+  stepnumber=1,                   % the step between two line-numbers.
+  numbersep=5pt,                  % how far the line-numbers are from the code
+  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
+  showspaces=false,               % show spaces adding particular underscores
+  showstringspaces=false,         % underline spaces within strings
+  showtabs=false,                 % show tabs within strings adding particular underscores
+  tabsize=2,                      % sets default tabsize to 2 spaces
+  captionpos=b,                   % sets the caption-position to bottom
+  breaklines=true,                % sets automatic line breaking
+  breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
+  title=\lstname,                 % show the filename of files included with \lstinputlisting;
+}
 
@@
 
-\begin{code}
-struct cell
-{
-  DOUBLE *x;
-  DOUBLE *y;
-  DOUBLE *z;
-  int64_t nelements;
-}
-\end{code}\footnote{Most of the code assumes that the number of galaxies is within INT_MAX $\sim$ 2.1 billion. So, it would appear 
+\footnote{Most of the code assumes that the number of galaxies is within INT_MAX $\sim$ 2.1 billion. So, it would appear 
 to make no sense that the field inside the struct cell has `nelements' of type int64_t. The reason I have 

### 2015-01-16 paper/ms.tex
Updating the Makefile + latest version of the paper

@@
-\documentclass[apj]{emulateapj}
+\documentclass[preprint,authoryear,12pt]{elsarticle}
 
@@
 \usepackage{aas_macros}
-\usepackage{natbib}
 \usepackage{rotating}
@@
 \usepackage{listings}
+\usepackage{caption}
+\usepackage{url}
+\journal{Astronomy and Computing}
 
@@
 \newcommand{\hkpc}{\ensuremath{{h^{-1}kpc}\xspace}}
-\newcommand{\emcee}{\texttt emcee}
+\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
 \newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
-
-%%\setstretch{2.0}
-%% \usepackage{etoolbox}
-%% \makeatletter
-%% \patchcmd{\@makecaption}
-%%   {\scshape}
-%%   {}
-%%   {}
-%%   {}
-%% \makeatother
 
@@
 
-\title{Cache is King: Presenting a Suite of Blazing-Fast, Correlation Function Code}
+\begin{frontmatter}
 
-\author{Manodeep Sinha\altaffilmark{1}}
-\affiliation{Department of Physics and Astronomy, Vanderbilt University, Nashville, TN, 37235}
-\altaffiltext{1}{manodeep.sinha@vanderbilt.edu}
+\title{Cache is King: Presenting a Suite of Fast, Correlation Function Codes}
+
+\author[ms]{Manodeep Sinha}
+\ead{manodeep.sinha@vanderbilt.edu}
+\address[ms]{6902 Stevenson Center, Department of Physics \& Astronomy, Vanderbilt University, Nashville, TN 37235}
 
@@
 
-\keywords{methods: numerical}
+\begin{keyword}
+methods: numerical \sep (cosmology:) large-scale structure of universe \sep
+\end{keyword}
 
-\maketitle
+\end{frontmatter}
 
 \section{INTRODUCTION}
-%% MS: I have not encountered this in the past -- but the footnotecounter needs to be reset. Otherwise, 
-%% the numbering starts at Nauthors+1. Presumably \altaffiltext uses \footnote internally and does not
-%% reset the counter like a responsible person!
-\setcounter{footnote}{0}
-The large-scale structure of the Universe can be now measured using $\mathcal{O}(million)$ galaxies 
+%% %% MS: I have not encountered this in the past -- but the footnotecounter needs to be reset. Otherwise, 
+%% %% the numbering starts at Nauthors+1. Presumably \altaffiltext uses \footnote internally and does not
+%% %% reset the counter like a responsible person!
+%% \setcounter{footnote}{0}
+The large-scale structure of the Universe can be now measured using $\mathcal{O}$(million) galaxies 
 using data from current surveys like SDSS/BOSS surveys. Upcoming surveys like LSST will probe even 
-deeper and wider and aim to target on $\mathcal{O}(10's of million)$ galaxies. With such a large galaxy data, 
+deeper and wider and aim to target on $\mathcal{O}$(10's of million) galaxies. With such a large galaxy data, 
 we can measure the galaxy density field fairly accurately in the data. While the number of galaxies observed 
@@
 immediately allows us to prune {\em all} of the points that are not within 1 cell offset in each dimension. However, even with this pruning, the actual 
-implementation of the algorithm matters. For instance, a linked-list in each cell performs $\sim 60x$ worse than the algorithm described here.\footnote{The 
+implementation of the algorithm matters. For instance, a linked-list in each cell performs $\sim 60\times$ worse than the algorithm described here.\footnote{The 
 usual linked-list is very cache-unfriendly. Each dereference requires a read from a new region of memory and an almost guaranteed cache miss. }
+
+\subsection{Partitioning the Particles based on \rmax}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[clip=true]{tikz_grid}
+\caption{A 2-D grid showing the bin-lattice partitioning scheme. The bigger square show the entire 
+domain, the red circles show a random distribution of 100 particles. Let's say we want to compute all pairs 
+for the target blue point, then we would only have to consider red points that are within one cell (the dark shaded region). 
+A circle with radius \rmax is also drawn to shown the actual pairs that will eventually count in the correlation function.} 
+\end{figure}
 
@@
 
-\footnote{Most of the code assumes that the number of galaxies is within INT_MAX $\sim$ 2.1 billion. So, it would appear 
-to make no sense that the field inside the struct cell has `nelements' of type int64_t. The reason I have 
-chosen to use int64_t is that under C alignment rules 4 bytes of padding would have appended anyway to the struct cell. So, I chose 
-to declare `nelements' as an int64_t and used those 4 bytes. Saves some future conversion/coding. }
+%% \begin{lstlisting}
+%% \end{lstlisting}
+
+%% \lstinputlisting{cellarray.h}
+
+%% \footnote{Most of the code assumes that the number of galaxies is within INT_MAX $\sim$ 2.1 billion. So, it would appear 
+%% to make no sense that the field inside the struct cell has `nelements' of type \verbatim{int64_t}. The reason I have 
+%% chosen to use \verbatim{int64_t} is that under C alignment rules 4 bytes of padding would have appended anyway to the struct cell. So, I chose 
+%% to declare `nelements' as an \verbatim{int64_t} and used those 4 bytes. Saves some future conversion/coding. }
 
@@
 
-\acknowledgements
+\citet{foreman-mackey_etal_13}
 
-\section{}
+\section*{Acknowledgements}
 
-\bibliographystyle{apj}
+\bibliographystyle{elsarticle-harv}
 \bibliography{master}

### 2015-01-21 paper/ms.tex
The C link line was required
Updated the paper structure + some text

@@
 \newcommand{\hkpc}{\ensuremath{{h^{-1}kpc}\xspace}}
+\newcommand{\emcee}{\texttt emcee}
 \newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
+\newcommand{\xir}{\ensuremath{{\xi(r)}}\xspace}
+\newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
+\newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
 \newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
@@
   stepnumber=1,                   % the step between two line-numbers.
-  numbersep=5pt,                  % how far the line-numbers are from the code
+  numbersep=10pt,                  % how far the line-numbers are from the code
   backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
@@
   breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
-  title=\lstname,                 % show the filename of files included with \lstinputlisting;
 }
@@
 
-%% \begin{lstlisting}
-%% \end{lstlisting}
+\lstinputlisting[firstline=11]{../xi_theory/include/cellarray.h}
 
-%% \lstinputlisting{cellarray.h}
+Since the typical particle data is small ($\sim$ 20-30 MB), duplicating the entire particle distribution does not impose a strong 
+constraint on the memory requirements. However, this duplication allows us to store the particles contiguously and produces fewer 
+cache misses while looping over particles in a cell. 
+The entire 3-D particle distribution is then deposited on the uniform grid. Now, for each particle we need to visit 27 total cells to 
+compute all possible pairs within \rmax.
 
-%% \footnote{Most of the code assumes that the number of galaxies is within INT_MAX $\sim$ 2.1 billion. So, it would appear 
-%% to make no sense that the field inside the struct cell has `nelements' of type \verbatim{int64_t}. The reason I have 
-%% chosen to use \verbatim{int64_t} is that under C alignment rules 4 bytes of padding would have appended anyway to the struct cell. So, I chose 
-%% to declare `nelements' as an \verbatim{int64_t} and used those 4 bytes. Saves some future conversion/coding. }
+\section{The pair-counting algorithm}
+
+\subsection{\xir}
+\subsection{\xirppi}
+\subsection{\wprp}
 
@@
 
-\subsection{Comparison to Tree Algorithms}
+\section{Benchmarks \& Scaling}
+In this section we present the runtimes and OpenMP scaling for the different codes.
 
-\section{Results}
-
-\section{Benchmarks \& Scaling}
+\subsection{\xir}
+\subsection{\xirppi}
+\subsection{\wprp}
 

### 2015-01-21 xi_theory/include/cellarray.h
Moved the scalar sorting into the actual cells and OpenMP'ized them (if enabled). Produces much better scaling with nthreads -- updated timings and plot as well

@@ -1,0 +1,18 @@
+#pragma once
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <inttypes.h>
+
+#include "function_precision.h" //definition of DOUBLE depending on Makefile flag -DOUBLE_PREC
+
+#define NLATMAX   100      /* maximum grid dimension in X-Y plane */
+
+typedef struct{
+  DOUBLE *x;
+  DOUBLE *y;
+  DOUBLE *z;
+  int64_t nelements;
+} cellarray;
+
+

### 2015-01-21 xi_theory/xi_rp_pi/countpairs_rp_pi.h

@@ -1,0 +1,25 @@
+#pragma once
+
+#include "function_precision.h" //for definition of DOUBLE
+#include <inttypes.h> //for uint64_t
+
+//define the results structure
+typedef struct{
+	uint64_t *npairs;
+	DOUBLE *rupp;
+	DOUBLE *rpavg;
+	DOUBLE pimax;
+	int nbin;
+	int npibin;
+} results_countpairs_rp_pi;
+
+results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
+											const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
+#ifdef USE_OMP
+											const int numthreads,
+#endif
+											const int autocorr,
+											const char *binfile,
+											const double pimax)  __attribute__((warn_unused_result));
+
+void free_results_rp_pi(results_countpairs_rp_pi **results);

### 2015-01-21 xi_theory/logbins.c
Moved the scalar sorting into the actual cells and OpenMP'ized them (if enabled). Produces much better scaling with nthreads -- updated timings and plot as well

@@ -1,0 +1,50 @@
+/* PROGRAM logbins
+
+   --- logbins xmin xmax nbin
+   --- Outputs the bins for a logarithmic bin spacing
+
+      * xmin   = inner radius of smallest bin
+      * xmax   = outer radius of largest bin
+      * nbin  = number of bins (logarithmically spaced in theta)
+      ----------------------------------------------------------------------
+*/
+
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+#include <assert.h>
+
+int main(int argc, char *argv[])
+{
+  /*---Arguments-------------------------*/
+  int nbin ;
+  double xmin,xmax,logxmin,logxmax,dlogx;
+
+  if(argc<4) {
+    fprintf(stderr,"Usage `%s' rmin rmax nbins \n",argv[0]);
+    exit(EXIT_FAILURE);
+  } else {
+    /*---Read-arguments-----------------------------------*/
+    sscanf(argv[1],"%lf",&xmin) ;
+    sscanf(argv[2],"%lf",&xmax) ;
+    sscanf(argv[3],"%d",&nbin) ;
+  }
+	assert(xmin > 0 && "rmin has to be non-zero, otherwise log is undefined");
+
+  /*---Setup-Pairs-arrays-------------------------------*/
+  logxmin = log10(xmin) ;
+  logxmax = log10(xmax) ;
+  dlogx = (logxmax-logxmin)/(double)nbin ;
+
+  double left_edge=logxmin;
+  for(int i=0;i<nbin;i++){
+    double right_edge=left_edge+dlogx; 
+    fprintf(stdout,"%20.12lf %20.12lf\n",pow(10.0,left_edge),pow(10.0,right_edge)) ;
+    left_edge=right_edge;
+  }
+  exit(EXIT_SUCCESS);
+}
+
+
+
+

### 2015-01-21 xi_theory/python_bindings/_countpairs.c

@@ -1,0 +1,310 @@
+#include <Python.h>
+#include <arrayobject.h>
+
+#include "countpairs.h"
+#include "countpairs_rp_pi.h"
+#include "countpairs_wp.h"
+
+//Docstrings for the methods
+static char module_docstring[]             =	"This module provides an interface for calculating correlation functions using C.";
+static char countpairs_docstring[]         =	"Calculate the 3-D \\xi auto/cross-correlation function given two sets of X1/Y1/Z1 and X2/Y2/Z2 arrays.";
+static char countpairs_rp_pi_docstring[]   =	"Calculate the 2-D DD(rp,pi) auto/cross-correlation function given two sets of X1/Y1/Z1 and X2/Y2/Z2 arrays.";
+static char countpairs_wp_docstring[]      =	"Calculate the projected auto-correlation function wp (assumes PERIODIC) given one set of X1/Y1/Z1 arrays.";
+
+/* function proto-type*/
+static PyObject *countpairs_countpairs(PyObject *self, PyObject *args);
+static PyObject *countpairs_countpairs_rp_pi(PyObject *self, PyObject *args);
+static PyObject *countpairs_countpairs_wp(PyObject *self, PyObject *args);
+PyMODINIT_FUNC init_countpairs(void);
+
+static PyMethodDef module_methods[] = {
+	{"countpairs"       ,countpairs_countpairs       ,METH_VARARGS,countpairs_docstring},
+	{"countpairs_rp_pi" ,countpairs_countpairs_rp_pi ,METH_VARARGS,countpairs_rp_pi_docstring},
+	{"countpairs_wp"    ,countpairs_countpairs_wp    ,METH_VARARGS,countpairs_wp_docstring},
+	{NULL, NULL, 0, NULL}
+};
+
+
+PyMODINIT_FUNC init_countpairs(void)
+{
+	PyObject *m = Py_InitModule3("_countpairs", module_methods, module_docstring);
+	if (m == NULL)
+		return;
+
[…]

### 2015-01-21 xi_theory/utils/gridlink.h

@@ -1,0 +1,22 @@
+#pragma once
+
+#include "function_precision.h"
+#include "cellarray.h"
+
+cellarray * gridlink(const int64_t np,
+					 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
+					 const DOUBLE xmin, const DOUBLE xmax,
+					 const DOUBLE ymin, const DOUBLE ymax,
+					 const DOUBLE zmin, const DOUBLE zmax,
+					 const DOUBLE max_x_size,
+					 const DOUBLE max_y_size,
+					 const DOUBLE max_z_size,
+					 const int xbin_refine_factor,
+					 const int ybin_refine_factor,
+					 const int zbin_refine_factor,
+					 int *nlattice_x,
+					 int *nlattice_y,
+					 int *nlattice_z) __attribute__((warn_unused_result)); 
+
+void get_max_min(const int64_t ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
+				 DOUBLE *min_x, DOUBLE *min_y, DOUBLE *min_z, DOUBLE *max_x, DOUBLE *max_y, DOUBLE *max_z);

### 2015-01-21 xi_theory/wp/countpairs_wp.c
Removed sglib header from countpairs_rp_pi

@@ -1,0 +1,473 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <assert.h>
+#include <stdint.h>
+#include <inttypes.h>
+
+#include "function_precision.h"
+#include "cellarray.h" //definition of struct cellarray
+#include "utils.h" //all of the utilities
+#include "gridlink.h"//function proto-type for gridlink
+#include "countpairs_wp.h" //function proto-type
+
+#include "sglib.h"
+
+#ifdef USE_AVX
+#include "avx_calls.h"
+#endif
+
+#ifdef USE_OMP
+#include <omp.h>
+#endif
+
+
+#ifndef PERIODIC
+#warning "wp is only valid for PERIODIC boundary conditions. Ignoring the Makefile (non)-definition of PERIODIC"
+#endif
+
+void free_results_wp(results_countpairs_wp **results)
+{
+	if(results == NULL)
+		return;
+
+	if(*results == NULL)
+		return;
+
+	results_countpairs_wp *tmp = *results;
+	free(tmp->npairs);
+	free(tmp->rupp);
+	free(tmp->wp);
+	free(tmp->rpavg);
+	free(tmp);
+	tmp = NULL;
+}
+
+
+results_countpairs_wp *countpairs_wp(const int64_t ND1, DOUBLE * restrict X1, DOUBLE * restrict Y1, DOUBLE * restrict Z1,
+																		 const double boxsize, 
+#ifdef USE_OMP
+																		 const int numthreads,
+#endif
+																		 const char *binfile,
+																		 const double pimax)
+{
+
+	int bin_refine_factor=2,zbin_refine_factor=1;
+	int nmesh_x, nmesh_y, nmesh_z;
+	
+
+#ifdef USE_OMP
+	if(numthreads == 1) {
+		bin_refine_factor=2;
+		zbin_refine_factor=1;
+	} else {
+		//seems redundant - but I can not discount the possibility that some
+		//combination of these refine factors will be faster on a different architecture.
+		bin_refine_factor=2;
+		zbin_refine_factor=1;
+	}
+#endif
+
[…]

### 2015-01-21 xi_theory/xi_rp_pi/DDrppi.c

@@ -1,0 +1,233 @@
+/* PROGRAM DDrppi
+
+   --- DDrppi rpmin rpmax nrpbin data1 data2 [pimax] > wpfile
+   --- Measure the cross-correlation function xi(rp,pi) for two different
+       data files (or autocorrelation if data1=data2).
+
+      * rpmin   = inner radius of smallest bin (in Mpc/h)
+      * rpmax   = outer radius of largest bin
+      * nrpbin  = number of bins (logarithmically spaced in r)
+      * data1   = name of first data file
+      * data2   = name of second data file
+      *[pimax]  = maximum value of line-of-sight separation (default=40 Mpc/h)
+      > DDfile  = name of output file <logrp log(<rp>) pi pairs>
+      ----------------------------------------------------------------------
+*/
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <assert.h>
+#include <inttypes.h>
+
+#include "defs.h" //for ADD_DIFF_TIME
+#include "function_precision.h" //definition of DOUBLE
+#include "countpairs_rp_pi.h" //function proto-type for countpairs
+#include "io.h" //function proto-type for file input
+#include "utils.h" //general utilities
+
+
+void Printhelp(void);
+
+int main(int argc, char *argv[])
+{
+
+  /*---Arguments-------------------------*/
+	char *file1=NULL,*file2=NULL;
+	char *fileformat1=NULL,*fileformat2=NULL;
+	char *binfile=NULL;
+  DOUBLE pimax ;
+	
+	
+  /*---Data-variables--------------------*/
+  int64_t ND1=0,ND2=0;
+
+  DOUBLE *x1=NULL,*y1=NULL,*z1=NULL;
+  DOUBLE *x2=NULL,*y2=NULL,*z2=NULL;//will point to x1/y1/z1 in case of auto-corr
+
+
+  /*---Corrfunc-variables----------------*/
+#ifndef USE_OMP
+	const char argnames[][30]={"file1","format1","file2","format2","binfile","pimax"};
+#else
+	int nthreads=2;
+	const char argnames[][30]={"file1","format1","file2","format2","binfile","pimax","Nthreads"};
+#endif
+  int nargs=sizeof(argnames)/(sizeof(char)*30);
+  
[…]

### 2015-01-21 xi_theory/xi_rp_pi/countpairs_rp_pi.c

@@ -1,0 +1,501 @@
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+
+#include "countpairs_rp_pi.h" //function proto-type
+#include "gridlink.h"//function proto-type for gridlink
+#include "cellarray.h" //definition of struct cellarray
+#include "utils.h" //all of the utilities
+
+
+#ifdef USE_AVX
+#include "avx_calls.h"
+#endif
+
+#ifdef USE_OMP
+#include <omp.h>
+#endif
+
+
+void free_results_rp_pi(results_countpairs_rp_pi **results)
+{
+	if(results==NULL)
+		return;
+
+	if(*results==NULL)
+		return;
+	
+	results_countpairs_rp_pi *tmp = *results;
+
+	free(tmp->npairs);
+	free(tmp->rupp);
+	free(tmp->rpavg);
+	free(tmp);
+	tmp = NULL;
+}
+
+
+results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
+											const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
+#ifdef USE_OMP
+											const int numthreads,
+#endif
+											const int autocorr,
+											const char *binfile,
+											const double pimax)
+{
+  int bin_refine_factor=1;
+	int zbin_refine_factor=2;
+	if(autocorr==1) {
+		bin_refine_factor=2;
+		zbin_refine_factor=2;
+	} else {
+		bin_refine_factor=1;
+		zbin_refine_factor=1;
+	}
+#ifdef USE_OMP
+	if(numthreads > 1) {
+		if(autocorr==1) {
+			bin_refine_factor=2;
+			zbin_refine_factor=2;
+		} else {
+			bin_refine_factor=1;
+			zbin_refine_factor=1;
+		}
+	}
+#endif
+	const int npibin = (int) pimax;
+  /***********************
+   *initializing the  bins
+   ************************/
+	double *rupp;
+	int nrpbin ;
+	double rpmin,rpmax;
[…]

### 2015-01-21 xi_theory/xi_rp_pi/countpairs_rp_pi.c
Removed sglib header from countpairs_rp_pi

@@ -8,8 +8,6 @@
 #include "utils.h" //all of the utilities
 
 
-#include "sglib.h"
-
 #ifdef USE_AVX
 #include "avx_calls.h"
 #endif

### 2015-01-21 xi_theory/include/defs.h
Removed sglib header from countpairs_rp_pi

@@ -1,0 +1,6 @@
+#pragma once
+
+
+#define ADD_DIFF_TIME(t0,t1)     ((t1.tv_sec - t0.tv_sec) + 1e-6*(t1.tv_usec - t0.tv_usec))
+
+

### 2015-01-21 xi_theory/io/io.c

@@ -1,0 +1,160 @@
+#include "io.h"
+#include "ftread.h"
+#include "utils.h"
+
+#ifndef MEMORY_INCREASE_FAC
+#define MEMORY_INCREASE_FAC 1.1
+#endif
+
+int64_t read_positions(const char *filename, const char *format, void **xpos, void **ypos, void **zpos, const size_t size)
+{
+  int64_t np;
+  void *x,*y,*z;
+  if(strncmp(format,"f",1)==0) {
+	//read fast-food file
+	size_t bytes=sizeof(int) + sizeof(float)*9 + sizeof(int);//skip fdat
+	bytes += sizeof(int) + sizeof(float) + sizeof(int); //skip znow
+	int idat[5];
+	FILE *fp = my_fopen(filename,"r");
+	my_ftread(idat,sizeof(int),5,fp);
+	np = (int64_t) idat[1]; //idat[1] is int
+	
+	assert((size == 4 || size == 8) && "Size of each position element can be either 4 (float) or 8 (double)");
+	//TO DO -- cover the case where the file is written in the other precision. 
+	
+	x = my_malloc(size,np);
+	y = my_malloc(size,np);
+	z = my_malloc(size,np);
+	
+	my_fseek(fp,bytes,SEEK_CUR);
+	//Check that the file was written with the requested precision
+	unsigned int dummy;
+	my_fread(&dummy,sizeof(dummy), 1, fp);
+	//so rewind by 4 bytes  prepare for calls to ftread
+	my_fseek(fp, -sizeof(dummy), SEEK_CUR);
+	dummy /= np;
+	assert((dummy == 4 || dummy == 8) && "File must contain either 4 byte (float) or 8 byte(double) precision");
+	
+	if(dummy == size) {
+	  my_ftread(x,size, np, fp);
+	  my_ftread(y,size, np, fp);
+	  my_ftread(z,size, np, fp);
+	} else {
+	  fprintf(stderr,"WARNING: File was written in a different precision than requested (file precision = %u requested precision = %zu)\n",dummy,size);
+	  //Okay so the file was written in a different precision.
+	  //First, print a warning message and then read-in correctly with the
+	  //requested precision
[…]

### 2015-01-22 paper/ms.tex
Removed sglib header from countpairs_rp_pi

@@
 \section{Benchmarks \& Scaling}
-In this section we present the runtimes and OpenMP scaling for the different codes.
+In this section we present the runtimes and OpenMP scaling for the three codes.
 
-\subsection{\xir}
-\subsection{\xirppi}
-\subsection{\wprp}
+\subsection{Scaling with OpenMP threads}
+\begin{figure}[htbp]
+\includegraphics[clip=true,width=0.5\linewidth]{xi_of_r_scaling_Mr19}%
+\includegraphics[clip=true,width=0.5\linewidth]{xi_rp_pi_scaling_Mr19}\\
+\includegraphics[clip=true,width=0.5\linewidth]{wp_scaling_Mr19}%
+\caption{OpenMP scaling for the \xir, \xirppi, \wprp }
+\end{figure}
 

### 2015-01-23 paper/ms.tex
Added an users-guide

--- 2015-01-21 paper/ms.tex
@@
   language=C,                % choose the language of the code
+  basicstyle=\ttfamily,      % sets font
   numbers=left,                   % where to put the line-numbers
+  numberstyle=\small, 
   stepnumber=1,                   % the step between two line-numbers.
@@
 
-\lstinputlisting[firstline=11]{../xi_theory/include/cellarray.h}
+\begin{lstlisting}
+typedef struct{
+  DOUBLE *x;
+  DOUBLE *y;
+  DOUBLE *z;
+  int64_t nelements;
+} cellarray;
+\end{lstlisting}
 
@@
 potentially increasing the throughput by a factor of 8x/4x. However, automatic vectorization is not always possible by the compiler and in those cases 
-we can write AVX vector intrinsics to directly manipulate 8 floats/4 doubles.
+we can write AVX vector intrinsics to directly manipulate 8 floats/4 doubles\footnote{Another option would be to use the vectorclass written by Agner Fog here: \url{http://agner.org/optimize/vectorclass/}}.
 
 \section{Benchmarks \& Scaling}
-In this section we present the runtimes and OpenMP scaling for the different codes.
+In this section we present the runtimes and scalings for different number of particles, \rmax and OpenMP threads for the codes. 
+For all of the scaling tests, only an auto-correlation calculation was used and the fiducial catalog contains $\sim 1.2$ million 
+galaxies on a periodic cube of side 420 \hMpc.
 
-\subsection{\xir}
-\subsection{\xirppi}
-\subsection{\wprp}
+\subsection{Scaling with Number of Particles}
+In Fig.~\ref{fig:scaling_numpart}, we show the scaling for the three codes with the number of particles. For this scaling, we subsampled 
+the fiducial mock to attain 10 logarithmic steps in particle number ranging from $1.2\times10^4$ to $1.2\times10^6$.
+
+\begin{figure}[htbp]
+\includegraphics[clip=true,width=\linewidth]{timings_Mr19_numpart}%
+\caption{Scaling with particle number for \xir, \xirppi, \wprp }
+\label{fig:scaling_numpart}
+\end{figure}
+
+\subsection{Scaling with \rmax}
+\begin{figure}[htbp]
+\includegraphics[clip=true,width=\linewidth]{timings_Mr19_rmax}%
+\caption{Scaling with \rmax for \xir, \xirppi, \wprp }
+\label{fig:scaling_rmax}
+\end{figure}
+
+\subsection{Scaling with OpenMP threads}
+\begin{figure}[htbp]
+\includegraphics[clip=true,width=\linewidth]{timings_Mr19_openmp}%
+\caption{OpenMP scaling for \xir, \xirppi, \wprp }
+\label{fig:scaling_openmp}
+\end{figure}
 
 \section{Conclusions}
+I have presented a suite of three fast correlation function codes that take advantage of the underlying hardware. The reasons why the codes 
+are much faster for typical workloads are:
+\begin{itemize}
+\item A `bin-lattice' scheme is used to first partition the computational domain into 3-D cells that can then be used to prune majority of the volume. 
+\item To obtain better cache-locality, the particle list is duplicated into a contiguous array for each dimension. Thus, all particles that fall into the
+same 3-D cell are stored in a contiguous array. 
+\item With the AVX instruction set, modern CPU's can process 8 floats/4 doubles simultaneously. The codes contain hand-written AVX intrinsics that offer a 
+factor of few speedup compared to the compiler generated scalar code. 
+\end{itemize}
 

### 2015-01-23 paper/users-guide.tex
Added an users-guide

\documentclass[12pt,a4paper]{article}

\usepackage{amsmath}
\usepackage{aas_macros}
\usepackage{rotating}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{color}
\usepackage{listings}
\usepackage{caption}
%\usepackage{url}
\usepackage{hyperref}
\usepackage{dirtree}
\usepackage{adjustbox}

\let\stdsection\section
\renewcommand\section{\newpage\stdsection}

\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
\newcommand{\xir}{\ensuremath{{\xi(r)}}\xspace}
\newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
\newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
\newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
\newcommand{\clang}{{\texttt{clang}}\xspace}
\newcommand{\icc}{{\texttt{icc}}\xspace}
\newcommand{\gcc}{{\texttt{gcc}}\xspace}

\newcommand\DotFill{\leavevmode\xleaders\hbox{.}\hfill\kern0pt}
\newcolumntype{M}[1]{>{\raggedright\arraybackslash}m{#1}}

\lstset{
  language=C,                % choose the language of the code
  basicstyle=\ttfamily,      % sets font
  numbers=left,                   % where to put the line-numbers
  numberstyle=\small, 
  stepnumber=1,                   % the step between two line-numbers.
  numbersep=10pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  tabsize=2,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
}

\begin{document}

\title{User Guide for Corrfunc}
\author{Manodeep Sinha}

\maketitle

\pagenumbering{Roman}
\tableofcontents
\newpage
\pagenumbering{arabic}

\section{Introduction}

\section{Installation}
The only requirements for the code to install is a valid C compiler, with OpenMP support. The AVX
instruction set can only be used for CPU's later than 2011 (Intel Sandy Bridge/ AMD Bulldozer or later).

\subsection{Getting the Source}
You can obtain the source in two ways: i) Clone the mercurial repo (\texttt{hg clone} \url{https://bitbucket.org/manodeep/corrfunc/}) 
or ii) Download the tar archive (\href{https://bitbucket.org/manodeep/corrfunc/downloads/corrfunc.1.0.0.tar.gz}{corrfunc.\$MAJOR.0.\$MINOR.tar.gz}) and 
unpack it in the directory where you wish to keep the files (\texttt{tar xvzf corrfunc.\$MAJOR.0.\$MINOR.tar.gz}). Here, \$MAJOR and \$MINOR
refer to the major and minor release versions (current \$MAJOR=1, \$MINOR=0). I will only change the \$MAJOR version if the API breaks.

The directory structure for the code looks like this:
\dirtree{%
.1 corrfunc.
.2 paper.
.2 xi\_theory.
.3 benchmarks\DotFill
  \begin{minipage}[t]{8cm}
    IDL scripts to run benchmarks{.} 
  \end{minipage}.
.3 bin\DotFill 
  \begin{minipage}[t]{8cm}
    Will be created to copy executable files when you run `make install'{.} 
  \end{minipage}.
.3 examples\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for example C bindings using the static libraries{.} 
  \end{minipage}.
.3 include\DotFill 
  \begin{minipage}[t]{8cm}
    Header files for static libraries{.} 
  \end{minipage}.
.3 io\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for reading in data{.} 
  \end{minipage}.
.3 lib\DotFill 
  \begin{minipage}[t]{8cm}
    Will be created to copy static libraries and python library after you run `make'{.} 
  \end{minipage}.
.3 python\_bindings\DotFill 
  \begin{minipage}[t]{8cm}
    Source files to generate python bindings{.} 
  \end{minipage}.
.3 tests\DotFill 
  \begin{minipage}[t]{8cm}
    Correct outputs for tests \textcolor{blue}{(work in progress)}{.} 
  \end{minipage}.
.4 data\DotFill 
  \begin{minipage}[t]{8cm}
    Mock galaxy catalogs for tests{.} 
  \end{minipage}.
.3 utils\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for creating 3-D grid and helper routines{.} 
  \end{minipage}.
.3 wp\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $w_p(r_p)${.} 
  \end{minipage}.
.3 xi\_of\_r\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $\xi(r)${.} 
  \end{minipage}.
.3 xi\_rp\_pi\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $\xi(r_p,\pi)${.} 
  \end{minipage}.
}

\subsection{Code Options}
There are a few code options that control both the Science case and the code compilation. All of these 
options are located in `common.mk' in the base directory (`corrfunc'). Edit the first few lines to set these 
options:

\begin{table}
\begin{center}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{ccccM{4in}} 
\toprule
\multicolumn{1}{c}{\textbf{Option Type}}   &
\multicolumn{1}{c}{\textbf{Option Name}}   &
\multicolumn{1}{c}{\textbf{Default State}} &
\multicolumn{1}{c}{\textbf{Requires}}      &
\multicolumn{1}{c}{\textbf{Notes}}      \\
\midrule
                 & PERIODIC            & Enabled  & None            & Enables periodic boundary conditions. \\
\textbf{Science} & OUTPUT\_RPAVG       & Disabled & DOUBLE\_PREC    & Outputs the average pair-separation in each bin. \xir and \wprp can be slower by more than $2\times$, \xirppi is less affected. \\
\midrule
                 & DOUBLE\_PREC        & Disabled & None                               & Computations are done using double precision. Slower and requires more RAM. \\
\textbf{Code}    & USE\_AVX            & Enabled  & CPU and compiler with AVX support  & CPUs later than 2011 have AVX support. Code will run much faster with this option. \\
                 & USE\_OMP            & Enabled  & OpenMP capable compiler            & Since \clang does not support OpenMP yet, \texttt{common.mk} will stop compilation with \clang when this flag is enabled. \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{center}
\caption{\footnotesize }
\label{table:sims}
\end{table}
Depending on your Science use-case and the cpu/compiler, you will want to set the different options. Once you set those options, you should set the 
C compiler, CC (available options are \icc, \gcc, \clang). Once you have set the compiler, installing should be as simple as typing `make' and 
`make install' in the \texttt{xi\_theory} directory. However, you may have to do more on Mac OSX - so I will outline some of the scenarios in 
Section~\ref{section:mac}. 
\subsection{Linux}
If the installation went well, you should have an executable called \texttt{run\_correlations} in the \texttt{examples} directory. Type \texttt{./run\_correlations} 
in the \texttt{examples} directory and you should see the code in action. The C source file \texttt{run\_correlations.c} also serves as an example to 
use the \xir, \xirppi and \wprp libraries in C.

\subsection{Mac OSX}\label{section:mac}
There can be two issues on MACs. One is that the default \gcc assembler supplied by \texttt{XCode} or \texttt{macports} is too old and does not support AVX instructions 
even when the CPU does. One way to get around this is by using the \clang assembler even when compiling with \gcc. The easiest way to do it is by replacing the 
default assembler with this following script (taken from \href{https://gist.github.com/ancapdev/8059572}{here}:

\lstinputlisting[language=sh]{as}

I have included the \texttt{as} script in the paper directory - copy it to the appropriate directory (\texttt{/opt/local/bin/} for me since I use \texttt{macports} \gcc 
on my laptop).

Another problem might come with running the python example codes in the \texttt{python\_bindings} directory. If you get an error message 
\texttt{Fatal Python error: PyThreadState\_Get: no current thread} when you run \texttt{python call\_correlation\_functions.py}, then the following 
steps might fix the problem (these are also noted in the FAQ). This error occurs when the python library used at compile time is not the same 
as the runtime python library. In all cases that I have seen, this error occurs when using the \texttt{conda} package manager for python\footnote{This 
behaviour is by design according to \texttt{conda}}. 
\begin{itemize}

\item Change the relative path for the shared python library \_countpairs.so. You can change the relative path by issuing the command: \\
{\scriptsize\texttt{install\_name\_tool -change libpython2.7.dylib  \`{}python-config --prefix\`{}/lib/libpython2.7.dylib \_countpairs.so}}

\item Add to the fallback library path environment variable. \\
{\scriptsize\texttt{export DYLD\_FALLBACK\_LIBRARY\_PATH=\`{}python-config --prefix\`{}/lib:\$DYLD\_FALLBACK\_LIBRARY\_PATH}}

\item If both of the above methods fail, then create a symbolic link \\
{\scriptsize\texttt{ln -s \`{}python-config --prefix\`{}/lib/libpython2.7.dylib}}

\end{itemize}

\section{Code Design}
\subsection{Input Files}
\subsection{C bindings}
The \texttt{examples} contains the files \texttt{run\_correlations.c} that shows how to use the three types of correlation function libraries from C. 
\subsection{Python Bindings}
The \texttt{python\_bindings} directory contains python bindings for python 2.x. Note that python3 is not supported out of the box\footnote{In the future, I might switch to 
cython to cover both python2 and python3}. If all went well, then typing \texttt{python call\_correlation\_functions.py} should run the example python code. If you get 
an error (and you are on a MAC), then refer to Section~\ref{section:mac} or the FAQ.

\section{Extending the Code}
\subsection{Different Type of Input Data File}
\subsection{Using SSE instead of AVX}
\subsection{Correlation function with weights}
The default mode of the correlation function calculations assume identical unit weights for all points. However, it is fairly 
straightforward to extend the code to support weights for individual points.

\section*{Acknowledgements}

%% \bibliographystyle{apj}
%% \bibliography{master}

\end{document}

### 2015-01-27 paper/users-guide.tex
Added a CSV file reader to io.c. Plus, the users guide is almost complete

@@
-\documentclass[12pt,a4paper]{article}
+\documentclass[12pt,titlepage]{article}
 
@@
 \usepackage{caption}
-%\usepackage{url}
-\usepackage{hyperref}
+\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
 \usepackage{dirtree}
 \usepackage{adjustbox}
+\usepackage{bytefield}
+\usepackage{chngcntr}
 
+%% stdsection is the old section -> no new page
+%% section produces a new page before a new section
 \let\stdsection\section
@@
 
+%% Some of the short-cuts I use
 \newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
@@
 
+%% Fill out space with dots
 \newcommand\DotFill{\leavevmode\xleaders\hbox{.}\hfill\kern0pt}
+
+%% New column type in tables
 \newcolumntype{M}[1]{>{\raggedright\arraybackslash}m{#1}}
 
+%% Code paper - so there are lots of code listings
 \lstset{
@@
   basicstyle=\ttfamily,      % sets font
-  numbers=left,                   % where to put the line-numbers
+  numbers=none,                   % where to put the line-numbers
   numberstyle=\small, 
@@
 }
+%% Change the listings counter to show Section number + listings number (rather than just the listings number)
+\AtBeginDocument{\counterwithin{lstlisting}{section}}
+
+%% I dislike paragraph indents
+\setlength{\parindent}{0pt}
 
@@
 \title{User Guide for Corrfunc}
-\author{Manodeep Sinha}
+%% Abuse the author macro to show all the info I want.
+\author{Manodeep Sinha, \\
+  Department of Physics \& Astronomy,\\
+  Vanderbilt University,\\
+  Nashville, TN 37235.\\
+  \href{mailto:manodeep@gmail.com}{\texttt{manodeep@gmail.com}}
+}
 
@@
 \tableofcontents
-\newpage
 \pagenumbering{arabic}
@@
 \section{Introduction}
+Correlation functions are a statistical measure of a density field and are widely used in large-scale structure formation. Generally, 
+the measurements are done {\em once}\footnote{which is why I have not bothered with releasing the codes to measure correlation 
+functions on data} on survey data and compared with model predictions in a Monte-Carlo Markov Chain. As such, the 
+correlation functions have to be measured repeatedly during an MCMC. The codes presented here are meant to cover the typical scenarios 
+of measuring correlation functions in theory-land. The primary consideration in writing these codes is speed\footnotemark -- the codes 
+presented here should outperform any other CPU based correlation functions codes by a wide-margin.
+
+\footnotetext{The secondary consideration was maintainability and ease of use for others. I have versions of these codes that are 
+even faster but are potentially gibberish for anyone other than me!}
 
@@
   \begin{minipage}[t]{8cm}
-    Correct outputs for tests \textcolor{blue}{(work in progress)}{.} 
+    Correct outputs for tests{.}
   \end{minipage}.
@@
 options are located in `common.mk' in the base directory (`corrfunc'). Edit the first few lines to set these 
-options:
+options (see Table.~\ref{table:options} for details) :
+\begin{itemize}
+\item Science options -- PERIODIC, OUTPUT\_RPAVG
+\item Code options -- DOUBLE\_PREC, USE\_AVX and USE\_OMP
+\end{itemize}
 
@@
 \end{center}
-\caption{\footnotesize }
-\label{table:sims}
+\caption{\footnotesize List of compilations options, what the options mean and their dependencies for the codes. }
+\label{table:options}
 \end{table}
@@
 C compiler, CC (available options are \icc, \gcc, \clang). Once you have set the compiler, installing should be as simple as typing `make' and 
-`make install' in the \texttt{xi\_theory} directory. However, you may have to do more on Mac OSX - so I will outline some of the scenarios in 
+`make install' in the \texttt{xi\_theory} directory. All the libraries are intentionally chosen to be \texttt{static} libraries just to avoid 
+any path conflicts. However, on MAC OSX, you may have to do more to get the library to work -- so I have outlined some of the scenarios in 
 Section~\ref{section:mac}. 
@@
 even when the CPU does. One way to get around this is by using the \clang assembler even when compiling with \gcc. The easiest way to do it is by replacing the 
-default assembler with this following script (taken from \href{https://gist.github.com/ancapdev/8059572}{here}:
+default assembler with the \texttt{as} script in the \texttt{paper} directory (taken from \href{https://gist.github.com/ancapdev/8059572}{this url}). Copy this
+\texttt{as} script to the appropriate directory (\texttt{/opt/local/bin/} for me since I use \texttt{macports} \gcc on my laptop).
 
-\lstinputlisting[language=sh]{as}
-
-I have included the \texttt{as} script in the paper directory - copy it to the appropriate directory (\texttt{/opt/local/bin/} for me since I use \texttt{macports} \gcc 
-on my laptop).
-
-Another problem might come with running the python example codes in the \texttt{python\_bindings} directory. If you get an error message 
-\texttt{Fatal Python error: PyThreadState\_Get: no current thread} when you run \texttt{python call\_correlation\_functions.py}, then the following 
+Another problem might come with running the python example codes in the \texttt{python\_bindings} directory. If you get an error message:
+\begin{itemize} 
+\item \texttt{Fatal Python error: PyThreadState\_Get: no current thread} 
+\end{itemize}
+when you run \texttt{python call\_correlation\_functions.py}, then the following 
 steps might fix the problem (these are also noted in the FAQ). This error occurs when the python library used at compile time is not the same 
@@
 {\scriptsize\texttt{ln -s \`{}python-config --prefix\`{}/lib/libpython2.7.dylib}}
+\end{itemize}
+If all went well, then you should be able to run the \texttt{run\_correlations} code in the \texttt{examples} directory as well as execute \texttt{python call\_correlation\_functions.py} 
+in the \texttt{python\_bindings} directory. In all of the above examples, I have assumed that the relevant python library is \texttt{libpython2.7.dylib} (the default 
+under \texttt{conda}) -- you may have to replace it with your python library version.
 
+\section{Running the Codes}
+\subsection{Input Files}
+The codes currently can handle these types of input data files:
+\begin{itemize}
+\item \texttt{ascii} -- White-space separated columns, format code is `a'.
+\item \texttt{csv}   -- Comma-separated values, format code is `c'.
+\item \texttt{fast-food} -- Fast-food, fortran binary format, format code is `f'.
+The fast-food file format is described in detail in Section~\ref{section:fastfood}.
+\end{itemize}
+For the \texttt{ascii} and \texttt{csv} files, the code reads in the first three columns as the co-moving \texttt{X/Y/Z} arrays. Note, that more columns 
+can be present but the code will ignore those columns.
+
+\subsubsection{The fast-food file format}\label{section:fastfood}
+The fast-food format is a fortran binary format -- all fields are surrounded with 4 bytes padding. These value of these padding bytes 
+is the number of bytes of data contained in between the padding bytes. For example, to write out \texttt{20 bytes of data} in 
+a fast-food file format would require a total of $4+20+4=28$ bytes. The first and last 4 bytes of the file will contain the value $20$ -- 
+showing that 20 bytes of real data are contained in between the two paddings.
+
+\flushleft The \texttt{fast-food} file consists of a header:
+\begin{lstlisting}[language=C,numbers=none]
+int idat[5];
+float fdat[9];
+float znow;
+\end{lstlisting}
+For the purposes of these correlation function codes, the only useful quantity is \texttt{idat[1]} which contains \texttt{N} -- the number of particles 
+in the data file.
+
+\vspace{0.2in}
+%% \begin{enumerate}
+%% \item
+\begin{bytefield}{28}
+  \bitheader{0-27} \\
+  \bitbox{4}{20} & \bitbox{4}{Int} & \bitbox{4}{\textcolor{red}{\texttt{N}}} \bitbox{12}{3 Integers}& \bitbox{4}{20}\\
+\end{bytefield}\\
+%% \item
+%% \begin{bytefield}{44}
+%%   \bitheader{0-43} \\
+%%  \bitbox{4}{36} & \bitbox{36}{9 Floats} & \bitbox{4}{36} \\
+%% \end{bytefield}
+%% %% \item
+%% \begin{bytefield}{12}
+%%   \bitheader{0-11} \\
+%%   \bitbox{4}{4} & \bitbox{4}{Float} & \bitbox{4}{4}\\
+%% \end{bytefield}
+%% \end{enumerate}
+The next 56 bytes contain two other fields (10 floats + 4 padding bytes of 4 bytes each) and their corresponding paddings. Since these bytes do not contain any data 
+that are useful in the context of these codes we will skip over the contents of these bytes. In \texttt{io.c}, I actually \texttt{fseek} over these bytes.
+
+\flushleft After this header, the actual \texttt{X/Y/Z} values are stored. The first 4 bytes after the header contains \texttt{N*4, for float precision} or \texttt{N*8, for 
+double precision} where \texttt{N=idat[1]}, is the number of particles in the file. After all the \texttt{X} values there will be another 4 bytes containing 
+\texttt{N*4} or \texttt{N*8}. {\em Note, that even when the \texttt{X/Y/Z} arrays are written out in double-precision, the padding is still 4 bytes.}
+\begin{minipage}[t]{\textwidth}
+\centerline{\underline{\bfseries Byte-structure of the X/Y/Z arrays in a fast-food file}}
+\vspace{0.2in}
+\begin{minipage}[t]{0.5\textwidth}
+\centering
+\textbf{floats}\\[3ex]
+\begin{bytefield}{4}
+  \bitheader{0-3} \\
+  \bitbox{4}{\texttt{N*4}} \\
+  \wordbox{1}{X[0]} \\
+  \wordbox{1}{X[1]} \\
+  \wordbox[]{1}{$\vdots$} \\[1ex]
+  \wordbox{1}{X[N-1]} \\
+  \bitbox{4}{\texttt{N*4}}
+\end{bytefield} 
+\end{minipage}%
+\begin{minipage}[t]{0.5\textwidth}
+\centering
+\textbf{doubles}\\[3ex]
+\begin{bytefield}{8}
+  \bitheader{0-7} \\
+  \bitbox{4}{\texttt{N*8}} \\
+  \wordbox{1}{X[0]} \\
+  \wordbox{1}{X[1]} \\
+  \wordbox[]{1}{$\vdots$} \\[1ex]
+  \wordbox{1}{X[N-1]} \\
+  \bitbox{4}{\texttt{N*8}}
+\end{bytefield}
+\end{minipage}
+\end{minipage}
+\subsection{Specifying the radial bins}\label{section:bins}
+The codes were intentionaly designed to read in a set of (somewhat) arbitrary\footnote{I assume bins are non-overlapping} set of bin specifications from a file.
+This way, you can specify disjoint bin-edges as well as
+use $0.0$ as a bin edge (which would be impossible if $\log$ bins are assumed). The bins are to be specified in a (white-space separated) text file in this manner:
+\begin{verbatim}
+r_low[0]  r_high[0]
+r_low[1]  r_high[1]
+.
+.
+.
+r_low[nbins-1]  r_high[nbins-1]
+\end{verbatim}
+where, \texttt{r\_low[i]} and \texttt{r\_high[i]} are the left and right edges of the \texttt{i}'th bin respectively. The text files should contains as many lines as the number of 
+bins desired. The \texttt{logbins} executable can be used to create such a text file containing $\log$ bins. The syntax for running \texttt{logbins} is:
+\begin{lstlisting}[language=sh,numbers=none]
+./logbins <rmin> <rmax> <nbins>   > <text filename>
+\end{lstlisting}
+An example of such a file with radial bins is the file \texttt{bins} in the \texttt{tests} directory.
+
+{\em Note, all of the three codes print the correlation function to \texttt{stdout} -- so be sure to redirect \texttt{stdout} to an output file. }
+
+\subsection{Running \texorpdfstring{\xir}{xi(r)}}
+To run the correlation function in 3-D, you will need to run the executable \texttt{DD} from either the \texttt{bin} or \texttt{xi\_of\_r} directory. The inputs 
+to \texttt{DD} are:
+\begin{itemize}
+\item \texttt{file1}   -- the file name for the first file.
+\item \texttt{format1} -- the file format for the first file.
+\item \texttt{file2}   -- the file name for the second file.
+\item \texttt{format2} -- the file format for the second file.
+\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
+\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
+\end{itemize}
+Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different).
+Out of the box,
+some sample \texttt{DD} call directory) can be:
+\begin{itemize}
+\item auto-correlation --  {\scriptsize \texttt{./DD  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 4 > Mr19\_output.DD}}
+\item cross-correlation -- {\scriptsize \texttt{./DD ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 4 > cmass\_output.DR}}
+\end{itemize}
+The output (printed to stdout) has \texttt{nbins} columns \texttt{<Npairs> <rpavg> <rmin> <rmax>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
+Makefile option OUTPUT\_RPAVG has been enabled. Here, each line of the output represents the \texttt{i}'th radial bin.
+
+\subsection{Running \texorpdfstring{\xirppi}{xi(rp,pi)}}
+To run the correlation function in 2-D for \xirppi, you will need to run the executable \texttt{DDrppi} from either the \texttt{bin} or \texttt{xi\_rp\_pi} directory. The inputs 
+to \texttt{DDrppi} are:
+\begin{itemize}
+\item \texttt{file1}   -- the file name for the first file.
+\item \texttt{format1} -- the file format for the first file.
+\item \texttt{file2}   -- the file name for the second file.
+\item \texttt{format2} -- the file format for the second file.
+\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
+\item \texttt{pimax}   -- the maximum distance to consider in the $\pi$ direction. The code uses \texttt{Z} axis as the $\pi$ (line-of-sight) direction. 
+\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
+\end{itemize}
+Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different).
+Out of the box,
+some sample \texttt{DDrppi} call directory) can be:
+\begin{itemize}
+\item auto-correlation --  {\scriptsize \texttt{./DDrppi  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 40.0 4 > Mr19\_output\_rppi.DD}}
+\item cross-correlation -- {\scriptsize \texttt{./DDrppi ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 80.0 4 > cmass\_output\_rppi.DR}}
+\end{itemize}
+The output (printed to stdout) has \texttt{nbins} columns \texttt{<Npairs> <rpavg> <log(rmin)> <pi\_upper>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
+Makefile option OUTPUT\_RPAVG has been enabled. The code bins in 1 Mpc/h in the $\pi$ direction by default. Therefore, 
+the total number of bins in the output file will be $nbins \times \pi_{\rm max}$, where each radial bin is further broken into $\pi_{\rm max}$ bins along the 
+$\pi$ direction.
+
+There is a code \texttt{wprp} in the \texttt{xi\_rp\_pi} directory that can combine the output of \texttt{DDrppi} for \texttt{DD}, \texttt{DR} and \texttt{RR} counts 
+and use Landy-Szalay estimator to produce a projected correlation function.
+
+\subsection{Running \texorpdfstring{\wprp}{wp(rp)}}
+To run the projected correlation function, you will need to run the executable \texttt{wp} from either the \texttt{bin} or \texttt{wp} directory. The inputs 
+to \texttt{wp} are:
+\begin{itemize}
+\item \texttt{boxsize} -- the boxsize for the periodic box. 
+\item \texttt{file}    -- file name.
+\item \texttt{format}  -- file format.
+\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
+\item \texttt{pimax}   -- the maximum distance to consider in the $\pi$ direction. The code uses \texttt{Z} axis as the $\pi$ (line-of-sight) direction. 
+\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
+\end{itemize}
+The \texttt{wp} code only computes an auto-correlation with PERIODIC boundary conditions, irrespective of the options set in \texttt{common.mk}. 
+\begin{itemize}
+\item {\scriptsize \texttt{./wp  420.0 ../tests/data/gals\_Mr19.ff f ../tests/bins 40.0 4 > Mr19\_output.wp}}
+\end{itemize}
+The output (printed to stdout) has \texttt{nbins} columns \texttt{<wp> <rpavg> <rmin> <rmax> <Npairs>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
+Makefile option OUTPUT\_RPAVG has been enabled.
+
+\section{Code Design}
+The following sections are taken directly from the associated \href{http://adsabs.harvard.edu/abs/2014arXiv1405.5832L}{paper}. The fundamental idea 
+behind the code can be broken down into the following steps:
+\begin{itemize}
+\item Given the max. separation, \rmax, for the correlation function, grid the entire the particle distribution with cell width \rmax. In case of 
+\xirppi and \wprp, the \texttt{X/Y} bins are \rmax while the \texttt{Z} bins is $\pi_{\rm max}$. 
+\item In each 3-D cell, store all the particles that are located in that cell in contiguous \texttt{X/Y/Z} arrays. This is contained in a structure 
+defined in \ref{code:celldefn}. 
+\item Loop over all particles in every cell (this loop uses OpenMP parallelization if USE\_OMP is enabled). This is the \texttt{index1} loop in 
+\texttt{countpairs.c, countpairs\_rp\_pi.c and countpairs\_wp.c}. The loop variable for the loop over each particle is \texttt{i}.
+\item Now, for every particle in the input list, we only need to look at particles in the neighbouring cells (see Fig.~\ref{fig:grid}). 
+\item Loop over neighbouring cells in all 3 dimensions. Corresponding variables are \texttt{iiix, iiiy, iiiz}. These variables are constructed from the 
+loop variables \texttt{iix, iiy, iiz} to account for periodic boundary conditions. \footnote{I apologize for this atrocious variable naming scheme.}
+\item Once we have a triplet of \texttt{iiix, iiiy, iiiz}, we can construct the index for the neighbouring cell. Variable \texttt{index2} gives the index 
+for the neighbouring cell. 
+\item Compute the distances between target particle and some particle bunch (where a chunk is 4 double or 8 floats) using AVX intrinsics. The slower, non-AVX 
+version of the code computes the distances serially and does not take advantage of the CPU architecture. This is the loop with loop variable \texttt{j}
+in \texttt{countpairs.c, countpairs\_rp\_pi.c, countpairs\_wp.c}. 
+\item Check if any of the distances are less than \rmax. If not, continue processing the data. If yes, then update the histograms using 
+AVX \texttt{bit-masks} and \texttt{popcount}. 
+\item Continue until all particles in neighbouring cell are done. 
 \end{itemize}
 
-\section{Code Design}
-\subsection{Input Files}
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth,clip=true]{tikz_grid}
+\caption{A 2-D grid showing the bin-lattice partitioning scheme. The bigger square show the entire
+domain, the red circles show a random distribution of 100 particles. Let's say we want to compute all pairs
+for the target blue point, then we would only have to consider red points that are within one cell (the dark shaded region).
+A circle with radius \rmax is also drawn to shown the actual pairs that will eventually count in the correlation function.}
+\label{fig:grid}
+\end{figure}
+
+\subsection{How to Maintain Cache Locality within the Grid}
+For all pairs around a given target galaxy, we need to compute distances to all points within all neighbouring 3-d cells.
+We ensure that the particle locations are contiguous by moving them into the following \texttt{C struct} in the order in which they arrive.
+
+\begin{lstlisting}[language=C,numbers=none,label={code:celldefn},caption={Definition of the cellarray structure. This 
+structure contains the \texttt{X/Y/Z} positions of all the particles that are in one 3-D cell. }]
+typedef struct{
+  DOUBLE *x;
+  DOUBLE *y;
+  DOUBLE *z;
+  int64_t nelements;
+} cellarray;
+\end{lstlisting}
+
+\stdsection{Calling the C Libraries}
+All of the correlation function codes create a corresponding \texttt{static} library rather than a \texttt{dynamic/shared} library. This was a design decision intended 
+to minimize path-issues for the end-user. After the libraries have been created, it is fairly straightforward to use them in an external \texttt{C/python} code. 
 \subsection{C bindings}
-The \texttt{examples} contains the files \texttt{run\_correlations.c} that shows how to use the three types of correlation function libraries from C. 
+The \texttt{examples} contains the files \texttt{run\_correlations.c} that shows how to use the three types of correlation function libraries from C. Essentially, the process 
+consists of including the appropriate header file and passing the appropriate arrays in the functions.
+
+\subsubsection{API for \texorpdfstring{\xir}{xi(r)}}
+\begin{lstlisting}[language=C,numbers=none,label={code:API_DD},basicstyle=\scriptsize]
+typedef struct{
+  uint64_t *npairs;
+  DOUBLE *rupp;
+  DOUBLE *rpavg;
+  int nbin;
+} results_countpairs;
+
+results_countpairs * countpairs(
+const int64_t ND1, const DOUBLE * const X1, const DOUBLE * const Y1, const DOUBLE  * const Z1,
+const int64_t ND2, const DOUBLE * const X2, const DOUBLE * const Y2, const DOUBLE  * const Z2,
+#ifdef USE_OMP
+const int numthreads,
+#endif
+const int autocorr,
+const char *binfile) __attribute__((warn_unused_result));
+
+void free_results(results_countpairs **results);
+\end{lstlisting}
+
+\subsubsection{API for \texorpdfstring{\xirppi}{xi(rp,pi)}}
+\begin{lstlisting}[language=C,numbers=none,label={code:API_DDrppi},basicstyle=\scriptsize]
+typedef struct{
+  uint64_t *npairs;
+  DOUBLE *rupp;
+  DOUBLE *rpavg;
+  DOUBLE pimax;
+  int nbin;
+  int npibin;
+} results_countpairs_rp_pi;
+
+results_countpairs_rp_pi * countpairs_rp_pi(
+const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
+const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
+#ifdef USE_OMP
+const int numthreads,
+#endif
+const int autocorr,
+const char *binfile,
+const double pimax)  __attribute__((warn_unused_result));
+
+void free_results_rp_pi(results_countpairs_rp_pi **results);
+\end{lstlisting}
+
+\subsubsection{API for \texorpdfstring{\wprp}{wp(rp)}}
+\begin{lstlisting}[language=C,numbers=none,label={code:API_wp},basicstyle=\scriptsize]
+typedef struct{
+  uint64_t *npairs;
+  DOUBLE *wp;
+  DOUBLE *rupp;
+  DOUBLE *rpavg;
+  DOUBLE pimax;
+  int nbin;
+} results_countpairs_wp;
+
+results_countpairs_wp *countpairs_wp(
+const int64_t ND1, DOUBLE * restrict X1, DOUBLE * restrict Y1, DOUBLE * restrict Z1,
+const double boxsize, 
+#ifdef USE_OMP
+const int numthreads,
+#endif
+const char *binfile,
+const double pimax) __attribute__((warn_unused_result));
+
+void free_results_wp(results_countpairs_wp **results);
+\end{lstlisting}
+
 \subsection{Python Bindings}
@@
 
-\section{Extending the Code}
+If you edit the \texttt{common.mk} file and compile for \texttt{double precision} arithmetic, then be sure to change the line:
+\begin{lstlisting}[language=python,numbers=none]
+dtype=np.float32
+\end{lstlisting}
+to 
+\begin{lstlisting}[language=python,numbers=none]
+dtype=np.float64.
+\end{lstlisting}
+Otherwise, you will get a \texttt{TypeError} at runtime.
+
+\stdsection{Extending the Code}
 \subsection{Different Type of Input Data File}
+All of the codes use \texttt{io.c} in the \texttt{io} sub-directory to read-in the data. If you want to specify a different file format, the easiest way would be to edit 
+\texttt{io.c}. Decide on the file format code and add another \texttt{strncmp} case in \texttt{io.c}. Remember that the \texttt{x/y/z} are declared as \texttt{void} pointers, 
+so you can not directly reference the \texttt{x/y/z} pointers. If you do add support for a different file-type, please submit a pull request and I will be happy to merge it into the 
+code-base.
+
+\subsection{Computing a different type of correlation function}
+Let's say, you want to compute a marked correlation function. Now, you will need to read-in/create the marks for each individual point. And then you will have to add an 
+appropriate field to the \texttt{cellarray} structure (see \ref{code:celldefn}).
+
 \subsection{Using SSE instead of AVX}
-\subsection{Correlation function with weights}
-The default mode of the correlation function calculations assume identical unit weights for all points. However, it is fairly 
-straightforward to extend the code to support weights for individual points.
-
-\section*{Acknowledgements}
+If your CPU is too old and does not support AVX, then you can still use SSE instrinsics to compute the correlation functions. However, this will require replacing all of the 
+AVX sections with corresponding SSE intrinsics. \href{mailto:manodeep@gmail.com}{Email me} and I will guide you through the conversion process.
 

### 2015-01-27 xi_theory/io/io.c

@@ -10,7 +10,7 @@ int64_t read_positions(const char *filename, const char *format, void **xpos, vo
 {
   int64_t np;
   void *x,*y,*z;
-  if(strncmp(format,"f",1)==0) {
+  if(strncmp(format,"f",1)==0) { /*Read-in fast-food file*/
 	//read fast-food file
 	size_t bytes=sizeof(int) + sizeof(float)*9 + sizeof(int);//skip fdat
 	bytes += sizeof(int) + sizeof(float) + sizeof(int); //skip znow
@@ -20,7 +20,6 @@ int64_t read_positions(const char *filename, const char *format, void **xpos, vo
 	np = (int64_t) idat[1]; //idat[1] is int
 	
 	assert((size == 4 || size == 8) && "Size of each position element can be either 4 (float) or 8 (double)");
-	//TO DO -- cover the case where the file is written in the other precision. 
 	
 	x = my_malloc(size,np);
 	y = my_malloc(size,np);
@@ -44,117 +43,174 @@ int64_t read_positions(const char *filename, const char *format, void **xpos, vo
 	  //Okay so the file was written in a different precision.
 	  //First, print a warning message and then read-in correctly with the
 	  //requested precision
-			if(dummy == 4) {
-			  assert(size == 8 && "Expected to be storing to doubles");
-			  float *tmp = my_malloc(dummy,np);
-			  double *tmp_x = (double *) x;
-			  double *tmp_y = (double *) y;
-			  double *tmp_z = (double *) z;
-			  //read-in x
-			  my_ftread(tmp, dummy, np, fp);
-			  for(unsigned int i=0;i<np;i++) tmp_x[i] = tmp[i];
+	  if(dummy == 4) {
+		assert(size == 8 && "Expected to be storing to doubles");
+		float *tmp = my_malloc(dummy,np);
+		double *tmp_x = (double *) x;
+		double *tmp_y = (double *) y;
+		double *tmp_z = (double *) z;
+		//read-in x
[…]

### 2015-01-27 xi_theory/python_bindings/call_correlation_functions.py
Added code to the python script to read in the ascii file using pandas (if pandas is installed)

@@ -4,6 +4,8 @@ routines from python. (The codes are written in C)
 
 Author: Manodeep Sinha <manodeep@gmail.com>
 
+Requires: numpy
+
 """
 from __future__ import print_function
 import os
@@ -11,14 +13,27 @@ import sys
 import numpy as np
 import _countpairs
 
+try:
+    import pandas as pd
+except ImportError:
+    pd = None
+
 file="../tests/data/gals_Mr19.txt"
 ### Make sure the precision agrees with the definition in ../common.mk.
 ### Otherwise, you will get a runtime error -- 
 ### TypeError TypeError: array cannot be safely cast to required type
 dtype=np.float32
-x,y,z = np.genfromtxt(file,dtype=dtype,unpack=True)
-array = np.hstack((x,y,z))
-boxsize=np.max(array)
+
+### Check if pandas is available - much faster to read in the data through pandas
+if pd is not None:
+    df = pd.read_csv(file,header=None,engine="c",dtype={"x":dtype,"y":dtype,"z":dtype},delim_whitespace=True)
+    x = np.asarray(df[0],dtype=dtype)
+    y = np.asarray(df[1],dtype=dtype)
+    z = np.asarray(df[2],dtype=dtype)
+else:
+    x,y,z = np.genfromtxt(file,dtype=dtype,unpack=True)
+
+boxsize=420.0
 nthreads=4
 pimax=40.0
 binfile="../tests/bins"

### 2015-01-28 paper/users-guide.tex
Added code to the python script to read in the ascii file using pandas (if pandas is installed)
Almost complete users-guide.

@@
-\documentclass[12pt,titlepage]{article}
+\documentclass[12pt,titlepage,justified]{article}
 
@@
 \usepackage{chngcntr}
+\usepackage{alltt}
 
@@
 \let\stdsection\section
-\renewcommand\section{\newpage\stdsection}
+%% \renewcommand\section{\newpage\stdsection}
 
@@
 \newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
+\newcommand{\pimax}{\ensuremath{{\pi_{\rm max}}}\xspace}
+\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}\xspace}}
 \newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
@@
   language=C,                % choose the language of the code
-  basicstyle=\ttfamily,      % sets font
+  basicstyle=\scriptsize\ttfamily,      % sets font
   numbers=none,                   % where to put the line-numbers
-  numberstyle=\small, 
-  stepnumber=1,                   % the step between two line-numbers.
-  numbersep=10pt,                  % how far the line-numbers are from the code
+%%   numberstyle=\small, 
+%%   stepnumber=1,                   % the step between two line-numbers.
+%%   numbersep=10pt,                  % how far the line-numbers are from the code
   backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
@@
   tabsize=2,                      % sets default tabsize to 2 spaces
-  captionpos=b,                   % sets the caption-position to bottom
   breaklines=true,                % sets automatic line breaking
   breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
+  captionpos=t,  
+  xleftmargin=17pt,
+  framexleftmargin=17pt,
+  framexrightmargin=17pt,
+  framexbottommargin=5pt,
+  framextopmargin=5pt,
+  showstringspaces=false,
+  frame=top,frame=bottom,
+  escapeinside={(*@}{@*)},
 }
@@
 \AtBeginDocument{\counterwithin{lstlisting}{section}}
+
+\DeclareCaptionFormat{listing}{\rule{\dimexpr\textwidth+17pt\relax}{0.4pt}\par\vskip1pt#1#2#3}
+\captionsetup[lstlisting]{format=listing,singlelinecheck=false, margin=0pt, font={sf},labelsep=space,labelfont=bf}
+
+%%
+\renewcommand{\thetable}{\Roman{table}}
+\captionsetup[table]{margin=0pt,labelsep=period,labelfont=bf,textfont={it}}
+\captionsetup[figure]{margin=0pt,labelsep=period,labelfont=bf,textfont={it}}
 
@@
 of measuring correlation functions in theory-land. The primary consideration in writing these codes is speed\footnotemark -- the codes 
-presented here should outperform any other CPU based correlation functions codes by a wide-margin.
+presented here should outperform any other CPU based correlation functions codes by a wide margin.
+
+Cache locality and hand-written AVX intrinsics are the reasons why the code is very fast. However, that also means that the code is 
+not very portable. I have tried my best to ensure that the codes work on Linux and MAC OSX. If it does not work for you, particularly 
+if you are on a reasonable Linux install, please \href{mailto:manodeep@gmail.com}{email me}. The API for the codes can be considered 
+frozen; I will not change the API without changing the MAJOR release version.
+
+The paper associated with the codes is being (slowly) written by me. I hope to submit the paper to Astronomy \& Computing and release 
+the codes simultaneously. Once I submit the paper to arXiv, there will be an actual paper arXiv link \href{http://adsabs.harvard.edu/abs/2014arXiv1405.5832L}{here}.
 
 \footnotetext{The secondary consideration was maintainability and ease of use for others. I have versions of these codes that are 
-even faster but are potentially gibberish for anyone other than me!}
+even faster but are much harder to modify/maintain by any one other than me !}
 
@@
   \end{minipage}.
-.3 wp\DotFill 
-  \begin{minipage}[t]{8cm}
-    Source files for $w_p(r_p)${.} 
-  \end{minipage}.
 .3 xi\_of\_r\DotFill 
@@
     Source files for $\xi(r_p,\pi)${.} 
+  \end{minipage}.
+.3 wp\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files for $w_p(r_p)${.} 
   \end{minipage}.
@@
 \begin{table}
-\begin{center}
+\centering
+\caption{\footnotesize List of compilations options, what the options mean and their dependencies for the codes. }
 \begin{adjustbox}{max width=\textwidth}
@@
 \end{adjustbox}
-\end{center}
-\caption{\footnotesize List of compilations options, what the options mean and their dependencies for the codes. }
 \label{table:options}
@@
 
-\section{Running the Codes}
+\subsection{Running the tests}
+If installation went fine, then run \texttt{make tests} to run a suite of tests. If any of the tests fail, then please \href{mailto:manodeep@gmail.com}{email me}. I have 
+never seen the tests fail unless I made some coding error while modifying the source. 
+Once the tests pass successfully, type \texttt{make install} to install the binaries into the \texttt{bin} directory.
+
+\stdsection{Running the Codes}
+The codes should run straight out of the box. Note, the \xir and \xirppi can compute both auto and cross-correlations with and without PERIODIC boundary conditions whereas the \wprp code 
+only computes the auto-correlation with PERIODIC boundary conditions. Also, since both the \xir and \xirppi codes compute cross-correlations, pairs are {\em double-counted} in both 
+\xir and \xirppi codes. You could, in theory, make the auto-correlation bit faster by only computing unique pairs (as is done in \wprp, see Section~\ref{section:pair_counting_wp} 
+for details). However, since the total run-time will dominated by the cross-correlation (the number of randoms is typically an order of magnitude larger than the number 
+of data points), I have not implemented those optimizations for the auto-correlation calculations. 
 \subsection{Input Files}
@@
 \flushleft The \texttt{fast-food} file consists of a header:
-\begin{lstlisting}[language=C,numbers=none]
+\begin{lstlisting}[language=C,numbers=none,caption={The header format for \texttt{fast-food} files}]
 int idat[5];
@@
 \vspace{0.2in}
-%% \begin{enumerate}
-%% \item
 \begin{bytefield}{28}
@@
 \end{bytefield}\\
-%% \item
-%% \begin{bytefield}{44}
-%%   \bitheader{0-43} \\
-%%  \bitbox{4}{36} & \bitbox{36}{9 Floats} & \bitbox{4}{36} \\
-%% \end{bytefield}
-%% %% \item
-%% \begin{bytefield}{12}
-%%   \bitheader{0-11} \\
-%%   \bitbox{4}{4} & \bitbox{4}{Float} & \bitbox{4}{4}\\
-%% \end{bytefield}
-%% \end{enumerate}
 The next 56 bytes contain two other fields (10 floats + 4 padding bytes of 4 bytes each) and their corresponding paddings. Since these bytes do not contain any data 
@@
 
-\flushleft After this header, the actual \texttt{X/Y/Z} values are stored. The first 4 bytes after the header contains \texttt{N*4, for float precision} or \texttt{N*8, for 
-double precision} where \texttt{N=idat[1]}, is the number of particles in the file. After all the \texttt{X} values there will be another 4 bytes containing 
-\texttt{N*4} or \texttt{N*8}. {\em Note, that even when the \texttt{X/Y/Z} arrays are written out in double-precision, the padding is still 4 bytes.}
+After this header, the actual \texttt{X/Y/Z} values are stored. The first 4 bytes after the header contains \texttt{N*4} for float precision or \texttt{N*8} for 
+double precision where \texttt{N=idat[1]}, is the number of particles in the file. After all the \texttt{X} values there will be another 4 bytes containing 
+\texttt{N*4} or \texttt{N*8}. {\em Note, that even when the \texttt{X/Y/Z} arrays are written out in double-precision, the padding is still 4 bytes.} The 
+blocks for \texttt{Y/Z} follow after the \texttt{X} block.
+
+\vspace{0.2in}
 \begin{minipage}[t]{\textwidth}
@@
 use $0.0$ as a bin edge (which would be impossible if $\log$ bins are assumed). The bins are to be specified in a (white-space separated) text file in this manner:
-\begin{verbatim}
-r_low[0]  r_high[0]
-r_low[1]  r_high[1]
-.
-.
-.
-r_low[nbins-1]  r_high[nbins-1]
-\end{verbatim}
+\begin{alltt}
+r\_low[0]  r\_high[0]
+r\_low[1]  r\_high[1]
+\vdots
+r\_low[nbins-1]  r\_high[nbins-1]
+\end{alltt}
 where, \texttt{r\_low[i]} and \texttt{r\_high[i]} are the left and right edges of the \texttt{i}'th bin respectively. The text files should contains as many lines as the number of 
-bins desired. The \texttt{logbins} executable can be used to create such a text file containing $\log$ bins. The syntax for running \texttt{logbins} is:
-\begin{lstlisting}[language=sh,numbers=none]
-./logbins <rmin> <rmax> <nbins>   > <text filename>
-\end{lstlisting}
-An example of such a file with radial bins is the file \texttt{bins} in the \texttt{tests} directory.
-
+bins desired. The \texttt{logbins} executable can be used to create such a text file containing $\log$ bins. The syntax for running \texttt{logbins} is:\\
+\vspace{0.2in}
+\texttt{./logbins <rmin> <rmax> <nbins>   > <text filename>}
+\vspace{0.2in}
+\\An example of such a file with radial bins is the file \texttt{bins} in the \texttt{tests} directory. 
 {\em Note, all of the three codes print the correlation function to \texttt{stdout} -- so be sure to redirect \texttt{stdout} to an output file. }
@@
 \subsection{Running \texorpdfstring{\xir}{xi(r)}}
-To run the correlation function in 3-D, you will need to run the executable \texttt{DD} from either the \texttt{bin} or \texttt{xi\_of\_r} directory. The inputs 
-to \texttt{DD} are:
+To run the correlation function in 3-D, you will need to run the executable \texttt{DD} from either the \texttt{bin} or \texttt{xi\_of\_r} directory. Note, that 
+\xir double-counts the pairs (as does \xirppi). The inputs to \texttt{DD} are:
 \begin{itemize}
@@
 \end{itemize}
-The output (printed to stdout) has \texttt{nbins} columns \texttt{<Npairs> <rpavg> <rmin> <rmax>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
+The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<Npairs> <rpavg> <rmin> <rmax>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
 Makefile option OUTPUT\_RPAVG has been enabled. Here, each line of the output represents the \texttt{i}'th radial bin.
@@
 \subsection{Running \texorpdfstring{\xirppi}{xi(rp,pi)}}
-To run the correlation function in 2-D for \xirppi, you will need to run the executable \texttt{DDrppi} from either the \texttt{bin} or \texttt{xi\_rp\_pi} directory. The inputs 
+To run the correlation function in 2-D for \xirppi, you will need to run the executable \texttt{DDrppi} from either the \texttt{bin} or \texttt{xi\_rp\_pi} directory. Note, that 
+\xirppi double-counts the pairs (as does \xir).The inputs 
 to \texttt{DDrppi} are:
@@
 \end{itemize}
-The output (printed to stdout) has \texttt{nbins} columns \texttt{<Npairs> <rpavg> <log(rmin)> <pi\_upper>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
+The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<Npairs> <rpavg> <log(rmin)> <pi\_upper>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the
 Makefile option OUTPUT\_RPAVG has been enabled. The code bins in 1 Mpc/h in the $\pi$ direction by default. Therefore, 
-the total number of bins in the output file will be $nbins \times \pi_{\rm max}$, where each radial bin is further broken into $\pi_{\rm max}$ bins along the 
+the total number of bins in the output file will be \texttt{nbins$\times$\pimax}, where each radial bin is further broken into \pimax bins along the 
 $\pi$ direction.
@@
 \end{itemize}
-The output (printed to stdout) has \texttt{nbins} columns \texttt{<wp> <rpavg> <rmin> <rmax> <Npairs>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
+The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<wp> <rpavg> <rmin> <rmax> <Npairs>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
 Makefile option OUTPUT\_RPAVG has been enabled.
 
-\section{Code Design}
+\stdsection{Code Design}
 The following sections are taken directly from the associated \href{http://adsabs.harvard.edu/abs/2014arXiv1405.5832L}{paper}. The fundamental idea 
@@
 \item Given the max. separation, \rmax, for the correlation function, grid the entire the particle distribution with cell width \rmax. In case of 
-\xirppi and \wprp, the \texttt{X/Y} bins are \rmax while the \texttt{Z} bins is $\pi_{\rm max}$. 
-\item In each 3-D cell, store all the particles that are located in that cell in contiguous \texttt{X/Y/Z} arrays. This is contained in a structure 
-defined in \ref{code:celldefn}. 
+\xirppi and \wprp, the \texttt{X/Y} bins are \rmax while the \texttt{Z} bins is \pimax.
+\item In each 3-D cell, store all the particles that are located in that cell in contiguous \texttt{X/Y/Z} arrays (see Listing~\ref{code:assign_to_cellarray}).
+The particle data are contained in a structure
+defined in Listing~\ref{code:celldefn}.
 \item Loop over all particles in every cell (this loop uses OpenMP parallelization if USE\_OMP is enabled). This is the \texttt{index1} loop in 
@@
 \item Compute the distances between target particle and some particle bunch (where a chunk is 4 double or 8 floats) using AVX intrinsics. The slower, non-AVX 
-version of the code computes the distances serially and does not take advantage of the CPU architecture. This is the loop with loop variable \texttt{j}
-in \texttt{countpairs.c, countpairs\_rp\_pi.c, countpairs\_wp.c}. 
+version of the code computes the distances serially and does not take advantage of the CPU architecture. This is the \texttt{j} loop in 
+\texttt{countpairs.c, countpairs\_rp\_pi.c, countpairs\_wp.c}. 
 \item Check if any of the distances are less than \rmax. If not, continue processing the data. If yes, then update the histograms using 
@@
 
+\subsection{Partitioning the Particles based on \texorpdfstring{\rmax}{rmax}}
 \begin{figure}
@@
 We ensure that the particle locations are contiguous by moving them into the following \texttt{C struct} in the order in which they arrive.
-
 \begin{lstlisting}[language=C,numbers=none,label={code:celldefn},caption={Definition of the cellarray structure. This 
@@
 \end{lstlisting}
+The code \texttt{gridlink.c} takes in an input list of 3 arrays \texttt{X/Y/Z} and grids them into a regular 3-D grid using the specified bins parameter
+\texttt{max\_x\_size} or \texttt{max\_y\_size} or \texttt{max\_z\_size} for the \texttt{X/Y/Z} axes respectively. \footnote{In practice, the bins may be further 
+subdivided using the corresponding bin refine factors. These bin refine factors seem to influence runtime the most, you should experiment with a few values of 
+\texttt{bin\_refine\_factor} and \texttt{zbin\_refine\_factor} to see what produces the best runtimes for your typical scenario.}
+Once the number of grid cells along each axes has been determined, we allocate memory for the \texttt{struct lattice}. This \texttt{struct lattice} is a 
+declared as an 1-D array; the conversion from the three indices in 3-D (\texttt{ix,iy,iz}) to a single index (\texttt{index}) happens through the last 
+line in Listing~\ref{code:onedim_index}. Such an 1-D array for \texttt{struct lattice} gives much better OpenMP scaling.   
+\begin{lstlisting}[language=C,numbers=none,label={code:onedim_index},caption={Accessing as \texttt{lattice[index]} rather than \texttt{lattice[ix][iy][iz]}.},basicstyle=\scriptsize]
+int64_t totncells  = nmesh_x * nmesh_y * nmesh_z;
+cellarray *lattice  = my_malloc(sizeof(cellarray), totncells);
+int64_t *nallocated = my_malloc(sizeof(*nallocated), totncells);
+int64_t index = ix*nmesh_y*nmesh_z + iy*nmesh_z + iz;
+\end{lstlisting}
+Now that we know the total number of cells in the entire domain, we need to allocated memory to store the particles in each cell. However, there is no way to 
+know the exact number of particles in each cell without processing the entire data-set; so, we pre-allocate with an estimate of the expected number of particles from the 
+volume of each 3-D cell (assuming a random distribution). Then, we can allocate memory for each of the \texttt{X/Y/Z} arrays inside each 3-D cell:
+\begin{lstlisting}[language=C,numbers=none,label={code:malloc_cellarray},caption={Pre-allocating memory for the \texttt{X/Y/Z} arrays in \texttt{struct cellarray}.},basicstyle=\scriptsize]
+for (int64_t index=0;index<totncells;index++) {
+  lattice[index].x = my_malloc(sizeof(DOUBLE),expected_n);
+  lattice[index].y = my_malloc(sizeof(DOUBLE),expected_n);
+  lattice[index].z = my_malloc(sizeof(DOUBLE),expected_n);
+  lattice[index].nelements=0;
+  nallocated[index] = expected_n;
+}
+\end{lstlisting}
+Here, \texttt{nallocated} is an array that keeps track of the amount of memory alread allocated for each cell. After this step, all cells have been allocated memory for \texttt{expected\_n} 
+particles. Now, we can begin to process the individual particles and assigning them to the 3-D cells (if enough memory has already been allocated to assign the new particle). 
+\begin{lstlisting}[language=C,numbers=none,label={code:assign_to_cellarray},caption={Assigning the particles to the \texttt{struct cellarray} in the cell.},basicstyle=\scriptsize]
+for (int64_t i=0;i<np;i++)  {
+  ix=(int)((x[i]-xmin)*xinv) ;
+  iy=(int)((y[i]-ymin)*yinv) ;
+  iz=(int)((z[i]-zmin)*zinv) ;
 
-\stdsection{Calling the C Libraries}
+  int64_t index = ix*nmesh_y*nmesh_z + iy*nmesh_z + iz;
+
+  if(lattice[index].nelements == nallocated[index]) {
+    expected_n = nallocated[index]*MEMORY_INCREASE_FAC;
+
+    lattice[index].x = my_realloc(lattice[index].x ,sizeof(DOUBLE),expected_n,''lattice.x'');
+    lattice[index].y = my_realloc(lattice[index].y ,sizeof(DOUBLE),expected_n,''lattice.y'');
+    lattice[index].z = my_realloc(lattice[index].z ,sizeof(DOUBLE),expected_n,''lattice.z'');
+
+    nallocated[index] = expected_n;
+  }
+  int64_t ipos=lattice[index].nelements;
+  lattice[index].x[ipos] = x[i];
+  lattice[index].y[ipos] = y[i];
+  lattice[index].z[ipos] = z[i];
+  lattice[index].nelements++;
+}
+\end{lstlisting}
+The loop goes over all of the particles and calculates the corresponding 3-D cell indices --- \texttt{ix,iy,iz}. With these 3 variables, the corresponding 1-D index, \texttt{index}, can be 
+calculated. The next lines check (and reallocate memory, if necessary) to ensure that enough memory has been allocated to the \texttt{struct lattice[index]} to accommodate this new particle. 
+The last 5 lines are simply assigning the particle into the appropriate \texttt{struct lattice[index]}. Once all the particles have been assigned, the cells contain contiguous \texttt{X/Y/Z} 
+arrays describing the original particle distribution.
+
+\subsection{The Pair-Counting Algorithms}
+After running through \texttt{gridlink}, the particle distribution is stored in contiguous \texttt{X/Y/Z} inside the 1-D array of \texttt{struct cellarray}. To find all 
+possible pairs, we first need to loop over all particles in the first data-set. 
+\begin{lstlisting}[language=C,numbers=none,label={code:loop_over_first_cells},caption={Looping over all cells in the first data-set.},basicstyle=\scriptsize]
+for (int64_t index1=0;index1<totncells;index1++) {
+  const cellarray *first = &lattice1[index1];
+  const DOUBLE *x1 = first->x;
+  const DOUBLE *y1 = first->y;
+  const DOUBLE *z1 = first->z;
+}
+\end{lstlisting}
+Now, we have the target cell pointer and the associated \texttt{x1/y1/z1} array pointers. Next, we need to get the indices for the neighbouring cells in 3-D. In order to do that, first 
+the 1-D index, \texttt{index1} needs to be converted into a set of three 3-D indices, \texttt{ix, iy, iz}. Listing~\ref{code:finding_3D_index_first} shows the conversion from 
+the 1-D index to the corresponding 3-D indices. 
+\begin{lstlisting}[language=C,numbers=none,label={code:finding_3D_index_first},caption={Reconstructing 3-D index for \texttt{first} cell in the first data-set.},basicstyle=\scriptsize]
+const int iz = index1 % nmesh_z ;
+const int ix = index1 / (nmesh_z * nmesh_y) ;
+const int iy = (index1 - iz - ix*nmesh_z*nmesh_y)/nmesh_z ;
+\end{lstlisting}
+After executing the code segment in Listing~\ref{code:finding_3D_index_first}, we have the full 3-D indices for the target cell. Now, we have to find all of the indices 
+for the neighbouring cells that can potentially satisfy the \rmax constraint. This requires considering all 3-D cells that are located within \texttt{bin\_refine\_factor} of 
+the target cell (for each dimension). Since the target cell, \texttt{first} has a 3-D X index of \texttt{ix}, this means all cells that have X indices in the range 
+\texttt{ix $\pm$ bin\_refine\_factor} can potentially have pairs that satisfy the \rmax constraint. However, in case of PERIODIC boundary conditions, we also have 
+to ensure that the indices (and the actual particle positions) wrap around on the other side of the cube. Listing~\ref{code:looping_over_neighbouring_cells} shows 
+how the looping over neighbouring cells is done for the X dimension. Similar segments follow in the actual code for the Y/Z dimensions. 
+\begin{lstlisting}[language=C,numbers=none,label={code:looping_over_neighbouring_cells},caption={Looping over all the neighbouring cells and taking care 
+of PERIODIC boundary conditions},basicstyle=\scriptsize]
+for(int iix=-bin_refine_factor;iix<=bin_refine_factor;iix++){
+  int iiix;
+#ifdef PERIODIC
+  DOUBLE off_xwrap=0.0;
+  if(ix + iix >= nmesh_x) {
+    off_xwrap = -xdiff;
+  } else if (ix + iix < 0) {
+    off_xwrap = xdiff;
+  }
+  iiix=(ix+iix+nmesh_x)%nmesh_x;
+#else
+  iiix = iix+ix;
+  if(iiix < 0 || iiix >= nmesh_x) {
+    continue;
+  }
+#endif
+  (*@{\Large$\vdots$}@*)
+  (*@{\normalsize \texttt{Similar chunks of code for Y/Z.}}@*)
+  (*@{\Large$\vdots$}@*)
+  const int64_t index2 = iiix*nmesh_y*nmesh_z + iiiy*nmesh_z + iiiz;
+\end{lstlisting}
+Once all of the three 3-D indices for the neighbouring cell has been determined, we can reconstruct the 1-D index, \texttt{index2} for that cell. With this 1-D index, we can 
+create a pointer, \texttt{second}, that contains the \texttt{cellarray pointer} to the neighbouring cell. In Listing~\ref{code:dereference_second_cell}, we show 
+how the neighbouring cell and the associated \texttt{x2/y2/z2} array pointers are defined. 
+\begin{lstlisting}[language=C,numbers=none,label={code:dereference_second_cell},caption={Dereferencing the pointers for the neighbouring (\texttt{second}) cell under consideration.},basicstyle=\scriptsize]
+const cellarray *second = &lattice2[index2];
+const DOUBLE *x2 = second->x;
+const DOUBLE *y2 = second->y;
+const DOUBLE *z2 = second->z;
+\end{lstlisting}
+At this point, we have a set of \texttt{x1/y/1/z1} arrays with \texttt{first->nelements} elements representing the first data-set.
+We also have another set of \texttt{x2/y2/z2} arrays with \texttt{second->nelements}
+elements representing the second data-set.
+
+\begin{lstlisting}[language=C,numbers=none,label={code:looping_over_first_particles},caption={Looping over all particles in the \texttt{first} cell and accounting for PERIODIC boundary conditions.},basicstyle=\scriptsize]
+for(int64_t i=0;i<first->nelements;i++) {
+  DOUBLE x1pos=x1[i];
+  DOUBLE y1pos=y1[i];
+  DOUBLE z1pos=z1[i];
+#ifdef PERIODIC
+  x1pos += off_xwrap;
+  y1pos += off_ywrap;
+  z1pos += off_zwrap;
+#endif
+\end{lstlisting}
+
+\begin{lstlisting}[language=C,numbers=none,label={code:looping_over_second_particles},caption={AVX intrinsics for looping over all particles in the \texttt{second} cell.
+PERIODIC boundary conditions have
+already been accounted for in \texttt{x1pos,y1pos,z1pos} variables.},basicstyle=\scriptsize]
+const AVX_FLOATS m_x1pos = AVX_SET_FLOAT(x1pos);
+const AVX_FLOATS m_y1pos = AVX_SET_FLOAT(y1pos);
+const AVX_FLOATS m_z1pos = AVX_SET_FLOAT(z1pos);
+
+int64_t j;
+for(j=0;j<=(second->nelements-NVEC);j+=NVEC) {
+  const AVX_FLOATS x2pos = AVX_LOAD_FLOATS_UNALIGNED(&x2[j]);
+  const AVX_FLOATS y2pos = AVX_LOAD_FLOATS_UNALIGNED(&y2[j]);
+  const AVX_FLOATS z2pos = AVX_LOAD_FLOATS_UNALIGNED(&z2[j]);
+\end{lstlisting}  
+The three codes \xir, \xirppi and \wprp diverge somewhat after this point. For \xir, we need to calculate the full 3-D separation, whereas for \xirppi and \wprp 
+the separation is the projected distance. I will discuss further implementations in the following sub-sections dedicated to each code. 
+\subsubsection{Pair-counting for \texorpdfstring{\xir}{xi(r)}}\label{section:pair_counting_r}
+\subsubsection{Pair-counting for \texorpdfstring{\xirppi}{xi(rp,pi)}}\label{section:pair_counting_rp_pi}
+\subsubsection{Pair-counting for \texorpdfstring{\wprp}{wp(rp)}}\label{section:pair_counting_wp}
+Since the pair-counting in \wprp always assumes PERIODIC boundary conditions {\em and} only computes an auto-correlation, some more optimizations are possible in the
+\wprp calculation. First, each individual cell in the \texttt{struct lattice} elements are sorted based on their \texttt{Z} arrays. The \texttt{X/Y} arrays are 
+also re-ordered simultaneously. Once all the \texttt{Z} values inside a cell have been sorted, we can avoid double-counting the pairs by changing the \texttt{iiz loop}
+to be:\\
+\vspace{0.2in}
+\texttt{for(int iiz={\bf 0};iiz<=zbin\_refine\_factor;iiz++)} 
+\vspace{0.2in}
+\\instead of\\
+\vspace{0.2in}
+\texttt{for(int iiz={\bf -zbin\_refine\_factor};iiz<=zbin\_refine\_factor;iiz++)} 
+\vspace{0.2in}
+
+Another advantage of sorting the \texttt{Z} values is earlier termination inside the actual calculation. The Listing~\ref{code:calculating_separations_wp} shows 
+how to break early from the \texttt{j-loop}. Since the \texttt{z2} values are always stored in increasing order, if {\em all} values of \texttt{zdiff:=z2[j:j+NVEC-1]-z1}
+are greater than \pimax, then none of the \texttt{zdiff} values in future iterations of the \texttt{j-loop} can be smaller than \pimax. When the code encounters 
+such a scenario, it updates the \texttt{j} variable to \texttt{second->nelements} to ensure that the serial section of the code is not executed and breaks out the 
+AVX \texttt{j-loop}. 
+\begin{lstlisting}[label={code:calculating_separations_wp},caption={AVX intrinsics for calculating separations in \wprp and checking for early termination.}]
+const AVX_FLOATS m_zdiff = AVX_SUBTRACT_FLOATS(m_z2,m_zpos);//z2[j:j+NVEC-1] - z1
+AVX_FLOATS m_mask_pimax = AVX_COMPARE_FLOATS(m_zdiff,m_pimax,_CMP_LT_OS);
+const int test = AVX_TEST_COMPARISON(m_mask_pimax);
+if(test == 0) {
+  j = second->nelements;
+  break;
+}
+\end{lstlisting}
+
+\subsection{AVX intrinsics to update the \texttt{npairs} histogram}
+This section explains the AVX intrinsics used to update the pair-counts histogram, \texttt{npairs}.\footnote{The thread-local version in \wprp is called \texttt{local\_npair}.}
+If the code execution reaches this loop, at least one (squared) pair separation falls within the range \texttt{sqr\_rpmin} and \texttt{sqr\_rpmax}, where \texttt{sqr\_rpmin} 
+is the squared lower radial limit of the first bin and \texttt{sqr\_rpmax} is the squared upper limit of the last bin (equivalent to \rmax$^2$ in this user-guide). 
+Since the pairs are more likely to occur in the largest separation bins, the loop goes backwards from the last bin to the first and uses early loop-termination in case
+all possible pairs have already been accounted for. 
+\begin{lstlisting}[label={code:updating_histogram},caption={AVX intrinsics for updating the \texttt{npairs} histgram for \xir and \wprp.},basicstyle=\scriptsize]
+for(int kbin=nrpbin-1;kbin>=1;kbin--) {
+  const AVX_FLOATS m1 = AVX_COMPARE_FLOATS(r2,m_rupp_sqr[kbin-1],_CMP_GE_OS);
+  const AVX_FLOATS m_bin_mask = AVX_BITWISE_AND(m1,m_mask_left);
+  m_mask_left = AVX_COMPARE_FLOATS(r2,m_rupp_sqr[kbin-1],_CMP_LT_OS);
+  const int test2  = AVX_TEST_COMPARISON(m_bin_mask);
+  npairs[kbin] += AVX_BIT_COUNT_INT(test2);
+#ifdef OUTPUT_RPAVG
+  m_rpbin = AVX_BLEND_FLOATS_WITH_MASK(m_rpbin,m_kbin[kbin], m_bin_mask);
+#endif
+  const int test3 = AVX_TEST_COMPARISON(m_mask_left);
+  if(test3 == 0) break;
+}
+\end{lstlisting}
+Note that \texttt{rupp\_sqr} (and its AVX equivalent, \texttt{m\_rupp\_sqr}) contains the squared upper limits for the bins. Thus, when considering bin \texttt{kbin}, 
+\texttt{m\_rupp\_sqr[kbin-1]} gives the squared lower radial limit for the bin while \texttt{m\_rupp\_sqr[kbin]} gives the squared upper limit for bin \texttt{kbin}. 
+Here, \texttt{m1} is the mask that contains all separations that satisfy \texttt{r2 $\geq$ rupp\_sqr[kbin-1]}, while the mask \texttt{m\_mask\_left} contains 
+those squared separations that satisfy \texttt{r2 < rupp\_sqr[kbin]}. Note, that \texttt{m\_mask\_left} is either computed before entering the \texttt{kbin} 
+loop or during a previous iteration of the same \texttt{kbin} loop. The AVX variable \texttt{m\_bin\_mask} then contains the bitwise and of \texttt{m1} and 
+\texttt{m\_mask\_left} -- the mask for the squared separations that fall into \texttt{kbin}. The variable \texttt{test2} contains an integer composed of 
+the upper set-bits of the mask \texttt{m\_bin\_mask} -- thus, contains only 4/8 useful bits for \texttt{float/double} precision calculations respectively. The 
+\texttt{npairs} pair-counts is then updated using a hardware \texttt{popcnt} instruction. The last two lines check if there are any more pairs left that satisfy 
+the lower bin-ranges; if not, the loop is terminated with a \texttt{break} statement.
+
+\section{Calling the C Libraries}
 All of the correlation function codes create a corresponding \texttt{static} library rather than a \texttt{dynamic/shared} library. This was a design decision intended 
 to minimize path-issues for the end-user. After the libraries have been created, it is fairly straightforward to use them in an external \texttt{C/python} code.  
+Be absolutely sure to pass arrays for the correct type -- float arrays if you did not use DOUBLE\_PREC (default) or double arrays if you did use DOUBLE\_PREC. 
+Also, only include headers from the \texttt{xi\_theory/include} directory -- those headers correspond to the actual static library. {\em DO NOT include} 
+the header files from the \texttt{xi\_of\_r}/\texttt{xi\_rp\_pi}/\texttt{wp} directories -- these headers do not contain the correct function signature for 
+the compilation flags used to generate the static libraries.
+
 \subsection{C bindings}
 The \texttt{examples} contains the files \texttt{run\_correlations.c} that shows how to use the three types of correlation function libraries from C. Essentially, the process 
-consists of including the appropriate header file and passing the appropriate arrays in the functions.
+consists of including the appropriate header file and passing the arrays ({\em of the correct \texttt{float/double} type}) into the functions. Make sure to include the static 
+library in the linking step to create a stand-alone executable (see the \texttt{Makefile} in the \texttt{examples} directory).
 
 \subsubsection{API for \texorpdfstring{\xir}{xi(r)}}
-\begin{lstlisting}[language=C,numbers=none,label={code:API_DD},basicstyle=\scriptsize]
+The interface for the 3-D correlation function, \xir, is through the \texttt{countpairs} function. Here is 
+the corresponding function signature:
+\begin{lstlisting}[language=C,numbers=none,label={code:API_DD},basicstyle=\scriptsize,caption={API for the 3-D \xir}]
+results_countpairs * countpairs(
+const int64_t ND1, const DOUBLE * const X1, const DOUBLE * const Y1, const DOUBLE  * const Z1,
+const int64_t ND2, const DOUBLE * const X2, const DOUBLE * const Y2, const DOUBLE  * const Z2,
+#ifdef USE_OMP
+const int numthreads,
+#endif
+const int autocorr,
+const char *binfile);
+\end{lstlisting}
+
+The parameters to the function are:
+\begin{itemize}
+\item \texttt{ND1} -- number of elements in the first data-set.
+\item \texttt{X1}  -- the array of X-values in the first data-set.
+\item \texttt{Y1}  -- the array of Y-values in the first data-set.
+\item \texttt{Z1}  -- the array of Z-values in the first data-set.
+\item \texttt{ND2} -- number of elements in the second data-set.
+\item \texttt{X2}  -- the array of X-values in the second data set.
+\item \texttt{Y2}  -- the array of Y-values in the second data set.
+\item \texttt{Z2}  -- the array of Z-values in the second data set.
+\item \texttt{numthreads} -- the number of threads to use (if USE\_OMP is enabled in \texttt{common.mk}).
+\item \texttt{autocorr} -- if an auto-correlation is being calculated (1 implies auto-correlation, flag is used for some runtime optimizations).
+\item \texttt{binfile} -- file name that contains the radial bins. See Section~\ref{section:bins} for details on how to create this file.
+\end{itemize}
+
+The output from \texttt{countpairs} is contained in \texttt{struct results\_countpairs}. The structure
+definition is:
+\begin{lstlisting}[language=C,numbers=none,label={code:API_DD_struct},basicstyle=\scriptsize,caption={Structure definition for the output of \xir}]
 typedef struct{
@@
 
-results_countpairs * countpairs(
-const int64_t ND1, const DOUBLE * const X1, const DOUBLE * const Y1, const DOUBLE  * const Z1,
-const int64_t ND2, const DOUBLE * const X2, const DOUBLE * const Y2, const DOUBLE  * const Z2,
+void free_results(results_countpairs **results);
+\end{lstlisting}
+
+The fields in the structure correspond to:
+\begin{itemize}
+\item \texttt{npairs} -- array containing the pair-counts. 
+\item \texttt{rupp}   -- array containing the upper limits of the bins. \texttt{rupp[0]} gives the lower-limit of the first radial bin. 
+\item \texttt{rpavg}  -- array containing the average value of the separations for all the pairs that fell into the bin. Will contain 
+meaningful values if OUTPUT\_RPAVG is defined in \texttt{common.mk}; identically \texttt{0.0} otherwise. 
+\item \texttt{nbin}   -- the number of radial bins used. Note that the actual pair-counts are stored in the index range \texttt{[1,nbin-1]}. The
+zero'th bin contains garbage for all of the arrays in this \texttt{struct results\_countpairs}.
+\end{itemize}
+After the results structure has been used, use \texttt{free\_results(\&results\_countpairs)} to free allocated memory.
+
+\subsubsection{API for \texorpdfstring{\xirppi}{xi(rp,pi)}}
+The interface for the 2-D correlation function, \xirppi, is through the \texttt{countpairs\_rp\_pi} function. Here is 
+the corresponding function signature:
+\begin{lstlisting}[language=C,numbers=none,label={code:API_DDrppi},basicstyle=\scriptsize,caption={API for the 2-D \xirppi}]
+results_countpairs_rp_pi * countpairs_rp_pi(
+const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
+const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
 #ifdef USE_OMP
@@
 const int autocorr,
-const char *binfile) __attribute__((warn_unused_result));
-
-void free_results(results_countpairs **results);
+const char *binfile,
+const double pimax);
 \end{lstlisting}
 
-\subsubsection{API for \texorpdfstring{\xirppi}{xi(rp,pi)}}
-\begin{lstlisting}[language=C,numbers=none,label={code:API_DDrppi},basicstyle=\scriptsize]
+The parameters to the function are:
+\begin{itemize}
+\item \texttt{ND1} -- number of elements in the first data-set.
+\item \texttt{X1}  -- the array of X-values in the first data-set.
+\item \texttt{Y1}  -- the array of Y-values in the first data-set.
+\item \texttt{Z1}  -- the array of Z-values in the first data-set.
+\item \texttt{ND2} -- number of elements in the second data-set.
+\item \texttt{X2}  -- the array of X-values in the second data set.
+\item \texttt{Y2}  -- the array of Y-values in the second data set.
+\item \texttt{Z2}  -- the array of Z-values in the second data set.
+\item \texttt{numthreads} -- the number of threads to use (if USE\_OMP is enabled in \texttt{common.mk}).
+\item \texttt{autocorr} -- if an auto-correlation is being calculated (1 implies auto-correlation, flag is used for some runtime optimizations).
+\item \texttt{binfile} -- file name that contains the radial bins. See Section~\ref{section:bins} for details on how to create this file.
+\item \texttt{pimax} -- the maximum line-of-sight distance (assumed to be the \texttt{Z} axis) to use. 
+\end{itemize}
+
+The output from \texttt{countpairs\_rp\_pi} is contained in \texttt{struct results\_countpairs\_rp\_pi}. The structure
+definition is:
+\begin{lstlisting}[language=C,numbers=none,label={code:API_DDrppi_struct},basicstyle=\scriptsize,,caption={Structure definition for the output of \xirppi}]
 typedef struct{
@@
 
-results_countpairs_rp_pi * countpairs_rp_pi(
-const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
-const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
+void free_results_rp_pi(results_countpairs_rp_pi **results);
+\end{lstlisting}
+
+The fields in the structure correspond to:
+\begin{itemize}
+\item \texttt{npairs} -- array containing the pair-counts.  Number of elements in the array is \texttt{(npibin+1)$\times$(nbin+1)}.
+\item \texttt{rupp}   -- array containing the upper limits of the bins. \texttt{rupp[0]} gives the lower-limit of the first radial bin. Number of elements 
+in the array is \texttt{nbin}. 
+\item \texttt{rpavg}  -- array containing the average value of the separations for all the pairs that fell into the bin. Will contain 
+meaningful values if OUTPUT\_RPAVG is defined in \texttt{common.mk}; identically \texttt{0.0} otherwise. Number of elements in the array is \texttt{(npibin+1)$\times$(nbin+1)}.
+\item \texttt{pimax}  -- the maximum line-of-sight distance used in the bins. 
+\item \texttt{nbin}   -- the number of radial bins used. Note that the actual pair-counts are stored in the index range \texttt{[1,nbin-1]}. The
+zero'th bin contains garbage for all of the arrays in this \texttt{struct results\_countpairs\_rp\_pi}.
+\item \texttt{npibin} -- the number of $\pi$ bins used. The total number of elements in the arrays in the \texttt{struct countpair\_rp\_pi} is 
+\texttt{(npibin+1)$\times$(nbin+1)}. As usual, meaningful data is only contained in the radial bin range \texttt{[1,nbin-1]}. 
+\end{itemize}
+After the results structure has been used, use \texttt{free\_results\_rp\_pi(\&results\_countpairs\_rp\_pi)} to free allocated memory.
+
+\subsubsection{API for \texorpdfstring{\wprp}{wp(rp)}}
+The interface for the projected correlation function, \wprp, is through the \texttt{countpairs\_wp} function. The code 
+{\em always} uses PERIODIC boundary conditions, irrespective of the settings in \texttt{common.mk}. Here is 
+the corresponding function signature\footnote{The \texttt{X1/Y1/Z1} arrays are not 
+declared with \texttt{const} qualifiers because I sort those arrays on \texttt{Z} in the 
+\texttt{countpairs\_wp} function.}:
+\begin{lstlisting}[language=C,numbers=none,label={code:API_wp},basicstyle=\scriptsize,caption={API for the \wprp}]
+results_countpairs_wp *countpairs_wp(
+const int64_t ND1, DOUBLE * restrict X1, DOUBLE * restrict Y1, DOUBLE * restrict Z1,
+const double boxsize, 
 #ifdef USE_OMP
@@
 #endif
-const int autocorr,
 const char *binfile,
-const double pimax)  __attribute__((warn_unused_result));
-
-void free_results_rp_pi(results_countpairs_rp_pi **results);
+const double pimax);
 \end{lstlisting}
 
-\subsubsection{API for \texorpdfstring{\wprp}{wp(rp)}}
-\begin{lstlisting}[language=C,numbers=none,label={code:API_wp},basicstyle=\scriptsize]
+The parameters to the function are:
+\begin{itemize}
+\item \texttt{ND1} -- number of elements in the data-set.
+\item \texttt{X1}  -- the array of X-values in the data-set.
+\item \texttt{Y1}  -- the array of Y-values in the data-set.
+\item \texttt{Z1}  -- the array of Z-values in the data-set.
+\item \texttt{boxsize}  -- the boxsize that fully contains the \texttt{X/Y/Z} values. 
+\item \texttt{numthreads} -- the number of threads to use (if USE\_OMP is enabled in \texttt{common.mk}).
+\item \texttt{binfile} -- file name that contains the radial bins. See Section~\ref{section:bins} for details on how to create this file.
+\item \texttt{pimax} -- the maximum line-of-sight distance (assumed to be the \texttt{Z} axis) to use. 
+\end{itemize}
+
+The output from \texttt{countpairs\_wp} is contained in \texttt{struct results\_countpairs\_wp}. The structure definition is:
+\begin{lstlisting}[language=C,numbers=none,label={code:API_wp_struct},basicstyle=\scriptsize,caption={Structure definition for the output of \wprp}]
 typedef struct{
@@
 
-results_countpairs_wp *countpairs_wp(
-const int64_t ND1, DOUBLE * restrict X1, DOUBLE * restrict Y1, DOUBLE * restrict Z1,
-const double boxsize, 
-#ifdef USE_OMP
-const int numthreads,
-#endif
-const char *binfile,
-const double pimax) __attribute__((warn_unused_result));
-
 void free_results_wp(results_countpairs_wp **results);
 \end{lstlisting}
+
+The fields in the structure correspond to:
+\begin{itemize}
+\item \texttt{npairs} -- array containing the pair-counts.  Number of elements in the array is \texttt{nbin}.
+\item \texttt{wp}     -- array containing the actual \texttt{wp} values. 
+\item \texttt{rupp}   -- array containing the upper limits of the bins. \texttt{rupp[0]} gives the lower-limit of the first radial bin. Number of elements 
+in the array is \texttt{nbin}. 
+\item \texttt{rpavg}  -- array containing the average value of the separations for all the pairs that fell into the bin. Will contain 
+meaningful values if OUTPUT\_RPAVG is defined in \texttt{common.mk}; identically \texttt{0.0} otherwise. Number of elements in the array is \texttt{nbin}.
+\item \texttt{pimax}  -- the maximum line-of-sight distance used in the bins. 
+\item \texttt{nbin}   -- the number of radial bins used. Note that the actual pair-counts are stored in the index range \texttt{[1,nbin-1]}. The
+zero'th bin contains garbage for all of the arrays in this \texttt{struct results\_countpairs\_wp}.
+\end{itemize}
+After the results structure has been used, use \texttt{free\_results\_wp(\&results\_countpairs\_wp)} to free allocated memory.
 
@@
 an error (and you are on a MAC), then refer to Section~\ref{section:mac} or the FAQ. 
+If you edit the \texttt{common.mk} file and compile for \texttt{double precision} arithmetic, then be sure to change the line \texttt{dtype=np.float32} to 
+\texttt{dtype=np.float64}. Otherwise, you will get a \texttt{TypeError} at runtime.
 
-If you edit the \texttt{common.mk} file and compile for \texttt{double precision} arithmetic, then be sure to change the line:
-\begin{lstlisting}[language=python,numbers=none]
-dtype=np.float32
-\end{lstlisting}
-to 
-\begin{lstlisting}[language=python,numbers=none]
-dtype=np.float64.
-\end{lstlisting}
-Otherwise, you will get a \texttt{TypeError} at runtime.
+\stdsection{Benchmarks \& Scaling}
+In this section we present the runtimes and scalings for different number of particles, \rmax and OpenMP threads for the codes. 
+For all of the scaling tests, only an auto-correlation calculation was used and the fiducial catalog contains $\sim 1.2$ million 
+galaxies on a periodic cube of side 420 \hMpc.
 
+\subsection{Scaling with Number of Particles}
+In Fig.~\ref{fig:scaling_numpart}, we show the scaling for the three codes with the number of particles. For this scaling, we subsampled 
+the fiducial mock to attain 10 logarithmic steps in particle number ranging from $1.2\times10^4$ to $1.2\times10^6$.
+
+\begin{figure}[htbp]
+\includegraphics[clip=true,width=\linewidth]{timings_Mr19_numpart}%
+\caption{Scaling with particle number for \xir, \xirppi, \wprp }
+\label{fig:scaling_numpart}
+\end{figure}
+
+\subsection{Scaling with \texorpdfstring{\rmax}{rmax}}
+\begin{figure}[htbp]
+\includegraphics[clip=true,width=\linewidth]{timings_Mr19_rmax}%
+\caption{Scaling with \rmax for \xir, \xirppi, \wprp }
+\label{fig:scaling_rmax}
+\end{figure}
+
+\subsection{Scaling with OpenMP threads}
+\begin{figure}[htbp]
+\includegraphics[clip=true,width=\linewidth]{timings_Mr19_openmp}%
+\caption{OpenMP scaling for \xir, \xirppi, \wprp }
+\label{fig:scaling_openmp}
+\end{figure}
+
+\clearpage
 \stdsection{Extending the Code}
@@
 \subsection{Computing a different type of correlation function}
-Let's say, you want to compute a marked correlation function. Now, you will need to read-in/create the marks for each individual point. And then you will have to add an 
-appropriate field to the \texttt{cellarray} structure (see \ref{code:celldefn}).
+Let's say, you want to compute a marked correlation function. Now, you will need to read-in/create the marks for each individual point. Here are the steps 
+you will need to create your custom correlation function code:
+\begin{enumerate}
+\item Add the data fields into the cellarray structure definition (see Listing~\ref{code:celldefn}.
+\item Add in the memory allocation for the fields in \texttt{gridlink.c} after the \texttt{malloc} for \texttt{X/Y/Z} pointers (see Listing~\ref{code:malloc_cellarray}).
+\item Extend \texttt{gridlink.c} to accept additional arrays and assign those arrays into \texttt{struct lattice} (see Listing~\ref{code:assign_to_cellarray}).
+\item Enable OUTPUT\_RPAVG and DOUBLE\_PREC (this is not required, but probably the easiest way to create a custom correlation function calculation).
+\item Declare and zero-initialize a \texttt{results} array that will contain your custom correlation function. Follow the implementation in the code for calculating \texttt{rpavg}.
+\item Add in the AVX arrays that load your custom data fields in the \texttt{j loop} in the \texttt{countpairs*} functions.
+\item Add your custom correlation function weight into the \texttt{results} array. Combine thread-local arrays into a global one in case USE\_OMP is selected. 
+\item Add the custom correlation function field to the corresponding \texttt{struct results\_countpairs*}. Assign the \texttt{results} array into the \texttt{struct results\_countpairs*}.
+\item Add a call to \texttt{free(your field)} in the corresponding function \texttt{free\_results\_countpairs*}.
+\end{enumerate}
 
@@
 
+\section{License}
+The code has been released under the MIT License. 
+\begin{verbatim}
+Permission is hereby granted, free of charge, to any person 
+obtaining a copy of this software and associated documentation 
+files (the ``Software''), to deal in the Software without 
+restriction, including without limitation the rights to use, 
+copy, modify, merge, publish, distribute, sublicense, and/or
+sell copies of the Software, and to permit persons to whom 
+the Software is furnished to do so, subject to the following 
+conditions:
+
+The above copyright notice and this permission notice shall 
+be included in all copies or substantial portions of the 
+Software.
+
+THE SOFTWARE IS PROVIDED ``AS IS'', WITHOUT WARRANTY OF ANY 
+KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE 
+WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR 
+PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS 
+OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR 
+OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT 
+OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION 
+WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN 
+THE SOFTWARE.
+\end{verbatim}
+
 \end{document}

### 2015-01-28 xi_theory/utils/gridlink.c

@@ -99,17 +99,17 @@ cellarray * gridlink(const int64_t np,
   lattice    = (cellarray *) my_malloc(sizeof(cellarray), totncells);
   nallocated = (int64_t *)       my_malloc(sizeof(*nallocated)      , totncells);
 
-  for (int i=0;i<nmesh_x;i++) {
-    for (int j=0;j<nmesh_y;j++) {
-      for (int k=0;k<nmesh_z;k++) {
-		int64_t index = i*nmesh_y*nmesh_z + j*nmesh_z + k;
-		lattice[index].x = my_malloc(sizeof(DOUBLE),expected_n);//This allocates extra and is wasteful
-		lattice[index].y = my_malloc(sizeof(DOUBLE),expected_n);
-		lattice[index].z = my_malloc(sizeof(DOUBLE),expected_n);
-		lattice[index].nelements=0;
-		nallocated[index] = expected_n;
-      }
-    }
+  /*
+	Allocate memory for each of the fields in cellarray. Since we haven't processed the data yet, 
+	expected_n is a reasonable guess as to the number of points in the cell. 
+   */
+  for (int64_t index=0;index<totncells;index++) {
+	lattice[index].x = my_malloc(sizeof(DOUBLE),expected_n);//This allocates extra and is wasteful
+	lattice[index].y = my_malloc(sizeof(DOUBLE),expected_n);
+	lattice[index].z = my_malloc(sizeof(DOUBLE),expected_n);
+	//allocate new fields in cellarray here (if you are creating a custom correlation function)
+	lattice[index].nelements=0;
+	nallocated[index] = expected_n;
   }
 
   DOUBLE xinv=1.0/xbinsize;
@@ -153,6 +153,8 @@ cellarray * gridlink(const int64_t np,
       nallocated[index] = expected_n;
     }
     assert(lattice[index].nelements < nallocated[index] && "Ensuring that number of particles in a cell doesn't corrupt memory");
+	//Index is the 1-D index for the 3-D cell. 
+	//ipos is the ipos'th particle in that 3-D cell.
     int64_t ipos=lattice[index].nelements;
     lattice[index].x[ipos] = x[i];
     lattice[index].y[ipos] = y[i];
@@ -160,6 +162,8 @@ cellarray * gridlink(const int64_t np,
     lattice[index].nelements++;
   }
   free(nallocated);
+
[…]

### 2015-01-28 xi_theory/wp/countpairs_wp.c

@@ -210,6 +210,8 @@ results_countpairs_wp *countpairs_wp(const int64_t ND1, DOUBLE * restrict X1, DO
 			const int iy = (index1 - iz - ix*nmesh_z*nmesh_y)/nmesh_z;
 			assert( ((iz + nmesh_z*iy + nmesh_z*nmesh_y*ix) == index1) && "Index reconstruction is wrong");
 	  
+			//Loop in the neighbouring cells in X. 
+			//iiix accounts for periodic boundary conditions
 			for(int iix=-bin_refine_factor;iix<=bin_refine_factor;iix++) {
 				const int iiix=(ix+iix+nmesh_x)%nmesh_x;
 				DOUBLE off_xwrap=0.0,off_ywrap=0.0,off_zwrap=0.0;
@@ -220,6 +222,8 @@ results_countpairs_wp *countpairs_wp(const int64_t ND1, DOUBLE * restrict X1, DO
 					off_xwrap = -side;
 				}
 				
+				//Loop in the neighbouring cells in Y. 
+				//iiiy accounts for periodic boundary conditions
 				for(int iiy=-bin_refine_factor;iiy<=bin_refine_factor;iiy++) {
 					const int iiiy=(iy+iiy+nmesh_y)%nmesh_y;
 					off_ywrap=0.0;
@@ -229,6 +233,11 @@ results_countpairs_wp *countpairs_wp(const int64_t ND1, DOUBLE * restrict X1, DO
 						off_ywrap = -side;
 					}
 
+					//Loop in the neighbouring cells in Z. 
+					//iiiz accounts for periodic boundary conditions
+					//However, since we are always doing an auto-correlation
+					//we can avoid double-counting by only looping forward
+					//in Z -> i.e., iiz loop starts from 0 rather than -zbin_refine_factor.
 					for(int iiz=0;iiz<=zbin_refine_factor;iiz++) {
 						const int iiiz=(iz+iiz+nmesh_z)%nmesh_z ;
 						off_zwrap = 0.0;
@@ -312,7 +321,7 @@ results_countpairs_wp *countpairs_wp(const int64_t ND1, DOUBLE * restrict X1, DO
 								const AVX_FLOATS m_y2 = AVX_LOAD_FLOATS_UNALIGNED(&y2[j]);
 								const AVX_FLOATS m_z2 = AVX_LOAD_FLOATS_UNALIGNED(&z2[j]);
 					
[…]

### 2015-01-28 paper/users-guide.tex
Latest version. Made some changes to tests suite in case of failures

@@
-\documentclass[12pt,titlepage,justified]{article}
+\documentclass[12pt,titlepage]{article}
 
+\usepackage{geometry}
 \usepackage{amsmath}
@@
 \usepackage{caption}
-\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
+\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue,breaklinks=true]{hyperref}
 \usepackage{dirtree}
@@
 \usepackage{alltt}
+
+\setlength{\textwidth}{6.5in}     %%                                   %%
+\setlength{\oddsidemargin}{0in}   %% (It is recommended that you       %%
+\setlength{\evensidemargin}{0in}  %%  not change these parameters,     %%
+\setlength{\textheight}{8.5in}    %%  at the risk of having your       %%
+\setlength{\topmargin}{0in}       %%  proposal dismissed on the basis  %%
+\setlength{\headheight}{0in}      %%  of incorrect formatting!!!)      %%
+\setlength{\headsep}{0in}         %%                                   %%
+\setlength{\footskip}{.5in}       %%                                   %%
 
@@
   language=C,                % choose the language of the code
-  basicstyle=\scriptsize\ttfamily,      % sets font
+  basicstyle=\small\ttfamily,      % sets font
   numbers=none,                   % where to put the line-numbers
@@
 
-Cache locality and hand-written AVX intrinsics are the reasons why the code is very fast. However, that also means that the code is 
+Cache locality and hand-written \href{http://en.wikipedia.org/wiki/Advanced_Vector_Extensions}{AVX} intrinsics are the reasons why the code is very fast. However, that also means that the code is 
 not very portable. I have tried my best to ensure that the codes work on Linux and MAC OSX. If it does not work for you, particularly 
@@
 The paper associated with the codes is being (slowly) written by me. I hope to submit the paper to Astronomy \& Computing and release 
-the codes simultaneously. Once I submit the paper to arXiv, there will be an actual paper arXiv link \href{http://adsabs.harvard.edu/abs/2014arXiv1405.5832L}{here}.
+the codes simultaneously. Once I submit the paper to arXiv, there will be an actual paper arXiv link \href{valid arXiv link}{here}.
 
@@
 or ii) Download the tar archive (\href{https://bitbucket.org/manodeep/corrfunc/downloads/corrfunc.1.0.0.tar.gz}{corrfunc.\$MAJOR.0.\$MINOR.tar.gz}) and 
-unpack it in the directory where you wish to keep the files (\texttt{tar xvzf corrfunc.\$MAJOR.0.\$MINOR.tar.gz}). Here, \$MAJOR and \$MINOR
-refer to the major and minor release versions (current \$MAJOR=1, \$MINOR=0). I will only change the \$MAJOR version if the API breaks.
+unpack it in the directory where you wish to keep the files (\texttt{tar xvzf}\\ \texttt{corrfunc.\$MAJOR.0.\$MINOR.tar.gz}). Here, \texttt{\$MAJOR}
+and \texttt{\$MINOR} refer to the major and minor release versions (current \texttt{\$MAJOR=1}, \texttt{\$MINOR=0}). I will only change 
+the \texttt{\$MAJOR} version in the highly unlikely event that the API changes.
 
+\subsection{Directory Structure}
 The directory structure for the code looks like this:
@@
 
-\subsection{Code Options}
+\subsection{Compilation Options}
 There are a few code options that control both the Science case and the code compilation. All of these 
@@
 \flushleft The \texttt{fast-food} file consists of a header:
-\begin{lstlisting}[language=C,numbers=none,caption={The header format for \texttt{fast-food} files}]
+\begin{lstlisting}[caption={The header format for \texttt{fast-food} files}]
 int idat[5];
@@
 \stdsection{Code Design}
-The following sections are taken directly from the associated \href{http://adsabs.harvard.edu/abs/2014arXiv1405.5832L}{paper}. The fundamental idea 
+The following sections are taken directly from the associated \href{valid arXiv link}{paper}. The fundamental idea 
 behind the code can be broken down into the following steps:
@@
 We ensure that the particle locations are contiguous by moving them into the following \texttt{C struct} in the order in which they arrive.
-\begin{lstlisting}[language=C,numbers=none,label={code:celldefn},caption={Definition of the cellarray structure. This 
+\begin{lstlisting}[label={code:celldefn},caption={Definition of the cellarray structure. This 
 structure contains the \texttt{X/Y/Z} positions of all the particles that are in one 3-D cell. }]
@@
 line in Listing~\ref{code:onedim_index}. Such an 1-D array for \texttt{struct lattice} gives much better OpenMP scaling.   
-\begin{lstlisting}[language=C,numbers=none,label={code:onedim_index},caption={Accessing as \texttt{lattice[index]} rather than \texttt{lattice[ix][iy][iz]}.},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:onedim_index},caption={Accessing as \texttt{lattice[index]} rather than \texttt{lattice[ix][iy][iz]}.}]
 int64_t totncells  = nmesh_x * nmesh_y * nmesh_z;
@@
 volume of each 3-D cell (assuming a random distribution). Then, we can allocate memory for each of the \texttt{X/Y/Z} arrays inside each 3-D cell:
-\begin{lstlisting}[language=C,numbers=none,label={code:malloc_cellarray},caption={Pre-allocating memory for the \texttt{X/Y/Z} arrays in \texttt{struct cellarray}.},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:malloc_cellarray},caption={Pre-allocating memory for the \texttt{X/Y/Z} arrays in \texttt{struct cellarray}.}]
 for (int64_t index=0;index<totncells;index++) {
@@
 particles. Now, we can begin to process the individual particles and assigning them to the 3-D cells (if enough memory has already been allocated to assign the new particle). 
-\begin{lstlisting}[language=C,numbers=none,label={code:assign_to_cellarray},caption={Assigning the particles to the \texttt{struct cellarray} in the cell.},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:assign_to_cellarray},caption={Assigning the particles to the \texttt{struct cellarray} in the cell.}]
 for (int64_t i=0;i<np;i++)  {
@@
 possible pairs, we first need to loop over all particles in the first data-set. 
-\begin{lstlisting}[language=C,numbers=none,label={code:loop_over_first_cells},caption={Looping over all cells in the first data-set.},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:loop_over_first_cells},caption={Looping over all cells in the first data-set.}]
 for (int64_t index1=0;index1<totncells;index1++) {
@@
 the 1-D index to the corresponding 3-D indices. 
-\begin{lstlisting}[language=C,numbers=none,label={code:finding_3D_index_first},caption={Reconstructing 3-D index for \texttt{first} cell in the first data-set.},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:finding_3D_index_first},caption={Reconstructing 3-D index for \texttt{first} cell in the first data-set.}]
 const int iz = index1 % nmesh_z ;
@@
 how the looping over neighbouring cells is done for the X dimension. Similar segments follow in the actual code for the Y/Z dimensions. 
-\begin{lstlisting}[language=C,numbers=none,label={code:looping_over_neighbouring_cells},caption={Looping over all the neighbouring cells and taking care 
-of PERIODIC boundary conditions},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:looping_over_neighbouring_cells},caption={Looping over all the neighbouring cells and taking care of PERIODIC boundary conditions.}]
 for(int iix=-bin_refine_factor;iix<=bin_refine_factor;iix++){
@@
 how the neighbouring cell and the associated \texttt{x2/y2/z2} array pointers are defined. 
-\begin{lstlisting}[language=C,numbers=none,label={code:dereference_second_cell},caption={Dereferencing the pointers for the neighbouring (\texttt{second}) cell under consideration.},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:dereference_second_cell},caption={Dereferencing the pointers for the neighbouring (\texttt{second}) cell under consideration.}]
 const cellarray *second = &lattice2[index2];
@@
 We also have another set of \texttt{x2/y2/z2} arrays with \texttt{second->nelements}
-elements representing the second data-set.
-
-\begin{lstlisting}[language=C,numbers=none,label={code:looping_over_first_particles},caption={Looping over all particles in the \texttt{first} cell and accounting for PERIODIC boundary conditions.},basicstyle=\scriptsize]
+elements representing the second data-set. Now, we have to compute all possible pair-wise separations between these two data-sets. We begin with a loop over the elements in 
+\texttt{first}:
+\begin{lstlisting}[label={code:looping_over_first_particles},caption={Looping over all particles in the \texttt{first} cell and accounting for PERIODIC boundary conditions.}]
 for(int64_t i=0;i<first->nelements;i++) {
@@
 \end{lstlisting}
+If PERIODIC boundary conditions are enabled, then \texttt{off\_xwrap}, \texttt{off\_ywrap} and \texttt{off\_zwrap} have been declared and initialized in Listing~\ref{code:looping_over_neighbouring_cells}.
+By wrapping the elements if the first data-set, we can avoid these operations in the following \texttt{j-loop} over all particles in \texttt{second}.
 
-\begin{lstlisting}[language=C,numbers=none,label={code:looping_over_second_particles},caption={AVX intrinsics for looping over all particles in the \texttt{second} cell.
-PERIODIC boundary conditions have
-already been accounted for in \texttt{x1pos,y1pos,z1pos} variables.},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:looping_over_second_particles},caption={AVX intrinsics for looping over all particles in the \texttt{second} cell. PERIODIC boundary conditions have 
+already been accounted for in \texttt{x1pos,y1pos,z1pos} variables.}]
 const AVX_FLOATS m_x1pos = AVX_SET_FLOAT(x1pos);
@@
   const AVX_FLOATS z2pos = AVX_LOAD_FLOATS_UNALIGNED(&z2[j]);
+
+  const AVX_FLOATS m_xdiff = AVX_SUBTRACT_FLOATS(m_x1pos,x2pos);
+  const AVX_FLOATS m_ydiff = AVX_SUBTRACT_FLOATS(m_y1pos,y2pos);
+  const AVX_FLOATS m_zdiff = AVX_SUBTRACT_FLOATS(m_z1pos,z2pos);
 \end{lstlisting}
+
 The three codes \xir, \xirppi and \wprp diverge somewhat after this point. For \xir, we need to calculate the full 3-D separation, whereas for \xirppi and \wprp 
@@
 \subsubsection{Pair-counting for \texorpdfstring{\xir}{xi(r)}}\label{section:pair_counting_r}
+Pair-counting for \xir is the most straight-forward. Once all the \texttt{NVEC} separations have been computed, we use bit-masks to check if any separations fall 
+within the range \texttt{sqr\_rpmin} and \texttt{sqr\_rpmax}. If not, we \texttt{continue} with the \texttt{j-loop}. 
+\begin{lstlisting}[label={code:masks_xi},caption={Bit-masks in \xir.}]
+m_mask_left = AVX_COMPARE_FLOATS(r2,m_sqr_rpmax,_CMP_LT_OS);
+if(AVX_TEST_COMPARISON(m_mask_left) == 0) {
+  continue;
+}
+
+const AVX_FLOATS m_mask = AVX_BITWISE_AND(m_mask_left, AVX_COMPARE_FLOATS(r2, m_sqr_rpmin, _CMP_GE_OS));
+if(AVX_TEST_COMPARISON(m_mask) == 0) {
+  continue;
+}
+r2 = AVX_BLEND_FLOATS_WITH_MASK(m_sqr_rpmax, r2, m_mask);
+m_mask_left = AVX_COMPARE_FLOATS(r2,m_sqr_rpmax,_CMP_LT_OS);
+\end{lstlisting}
+
 \subsubsection{Pair-counting for \texorpdfstring{\xirppi}{xi(rp,pi)}}\label{section:pair_counting_rp_pi}
+
 \subsubsection{Pair-counting for \texorpdfstring{\wprp}{wp(rp)}}\label{section:pair_counting_wp}
@@
 all possible pairs have already been accounted for. 
-\begin{lstlisting}[label={code:updating_histogram},caption={AVX intrinsics for updating the \texttt{npairs} histgram for \xir and \wprp.},basicstyle=\scriptsize]
+\begin{lstlisting}[label={code:updating_histogram},caption={AVX intrinsics for updating the \texttt{npairs} histgram for \xir and \wprp.}]
 for(int kbin=nrpbin-1;kbin>=1;kbin--) {
@@
 the corresponding function signature:
-\begin{lstlisting}[language=C,numbers=none,label={code:API_DD},basicstyle=\scriptsize,caption={API for the 3-D \xir}]
+\begin{lstlisting}[label={code:API_DD},caption={API for the 3-D \xir.}]
 results_countpairs * countpairs(
@@
 definition is:
-\begin{lstlisting}[language=C,numbers=none,label={code:API_DD_struct},basicstyle=\scriptsize,caption={Structure definition for the output of \xir}]
+\begin{lstlisting}[label={code:API_DD_struct},caption={Structure definition for the output of \xir.}]
 typedef struct{
@@
 the corresponding function signature:
-\begin{lstlisting}[language=C,numbers=none,label={code:API_DDrppi},basicstyle=\scriptsize,caption={API for the 2-D \xirppi}]
+\begin{lstlisting}[label={code:API_DDrppi},caption={API for the 2-D \xirppi}]
 results_countpairs_rp_pi * countpairs_rp_pi(
@@
 definition is:
-\begin{lstlisting}[language=C,numbers=none,label={code:API_DDrppi_struct},basicstyle=\scriptsize,,caption={Structure definition for the output of \xirppi}]
+\begin{lstlisting}[label={code:API_DDrppi_struct},caption={Structure definition for the output of \xirppi}]
 typedef struct{
@@
 \texttt{countpairs\_wp} function.}:
-\begin{lstlisting}[language=C,numbers=none,label={code:API_wp},basicstyle=\scriptsize,caption={API for the \wprp}]
+\begin{lstlisting}[label={code:API_wp},caption={API for the \wprp.}]
 results_countpairs_wp *countpairs_wp(
@@
 The output from \texttt{countpairs\_wp} is contained in \texttt{struct results\_countpairs\_wp}. The structure definition is:
-\begin{lstlisting}[language=C,numbers=none,label={code:API_wp_struct},basicstyle=\scriptsize,caption={Structure definition for the output of \wprp}]
+\begin{lstlisting}[label={code:API_wp_struct},caption={Structure definition for the output of \wprp}]
 typedef struct{

### 2015-01-28 xi_theory/tests/test_periodic.c

@@ -38,7 +38,7 @@
 #include "../xi_rp_pi/countpairs_rp_pi.c"
 #include "../wp/countpairs_wp.c"
 
-char tmpoutputfile[]="./tmp_output.txt";
+char tmpoutputfile[]="./test_periodic_output.txt";
 
 int test_periodic_DD(const char *correct_outputfile);
 int test_periodic_DDrppi(const char *correct_outputfile);
@@ -256,6 +256,9 @@ int main(int argc, char **argv)
 		system(execstring);
 	  } else {
 		fprintf(stderr,ANSI_COLOR_RED "FAILED: " ANSI_COLOR_MAGENTA "%s" ANSI_COLOR_RED ". Time taken = %8.2lf seconds " ANSI_COLOR_RESET "\n", testname,pair_time);
+		char execstring[MAXLEN];
+		my_snprintf(execstring,MAXLEN,"mv %s %s.%d",tmpoutputfile,tmpoutputfile,i);
+		system(execstring);
 		failed++;
 	  }
 	}
@@ -277,8 +280,11 @@ int main(int argc, char **argv)
 		if(status==EXIT_SUCCESS) {
 		  fprintf(stderr,ANSI_COLOR_GREEN "PASSED: " ANSI_COLOR_MAGENTA "%s" ANSI_COLOR_GREEN ". Time taken = %8.2lf seconds " ANSI_COLOR_RESET "\n", testname,pair_time);
 		} else {
-		  failed++;
 		  fprintf(stderr,ANSI_COLOR_RED "FAILED: " ANSI_COLOR_MAGENTA "%s" ANSI_COLOR_RED ". Time taken = %8.2lf seconds " ANSI_COLOR_RESET "\n", testname,pair_time);
+		  char execstring[MAXLEN];
+		  my_snprintf(execstring,MAXLEN,"mv %s %s.%d",tmpoutputfile,tmpoutputfile,this_test_num);
+		  system(execstring);
+		  failed++;
 		}
 	  }
 	}

### 2015-01-28 xi_theory/tests/test_nonperiodic.c
Latest version. Made some changes to tests suite in case of failures

@@ -38,7 +38,7 @@
 #include "../xi_rp_pi/countpairs_rp_pi.c"
 
 
-char tmpoutputfile[]="./tmp_output.txt";
+char tmpoutputfile[]="./test_nonperiodic_output.txt";
 
 int test_nonperiodic_DD(const char *correct_outputfile);
 int test_nonperiodic_DDrppi(const char *correct_outputfile);
@@ -232,8 +232,11 @@ int main(int argc, char **argv)
 		my_snprintf(execstring,MAXLEN,"rm -f %s",tmpoutputfile);
 		system(execstring);
 	  } else {
-		failed++;
 		fprintf(stderr,ANSI_COLOR_RED "FAILED: " ANSI_COLOR_MAGENTA "%s" ANSI_COLOR_RED ". Time taken = %8.2lf seconds " ANSI_COLOR_RESET "\n", testname,pair_time);
+		char execstring[MAXLEN];
+		my_snprintf(execstring,MAXLEN,"mv %s %s.%d",tmpoutputfile,tmpoutputfile,i);
+		system(execstring);
+		failed++;
 	  }
 	}
   } else {
@@ -254,8 +257,11 @@ int main(int argc, char **argv)
 		if(status==EXIT_SUCCESS) {
 		  fprintf(stderr,ANSI_COLOR_GREEN "PASSED: " ANSI_COLOR_MAGENTA "%s" ANSI_COLOR_GREEN ". Time taken = %8.2lf seconds " ANSI_COLOR_RESET "\n", testname,pair_time);
 		} else {
-		  failed++;
 		  fprintf(stderr,ANSI_COLOR_RED "FAILED: " ANSI_COLOR_MAGENTA "%s" ANSI_COLOR_RED ". Time taken = %8.2lf seconds " ANSI_COLOR_RESET "\n", testname,pair_time);
+		  char execstring[MAXLEN];
+		  my_snprintf(execstring,MAXLEN,"mv %s %s.%d",tmpoutputfile,tmpoutputfile,this_test_num);
+		  system(execstring);
+		  failed++;
 		}
 	  }
 	}

### 2015-01-28 paper/users-guide.tex
Changed variables to be consistent across codes

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{aas_macros}
@@
 \newcommand{\pimax}{\ensuremath{{\pi_{\rm max}}}\xspace}
-\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}\xspace}}
+\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}}\xspace}
 \newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
@@
 
-\title{User Guide for Corrfunc}
+\title{Users Guide for Corrfunc}
 %% Abuse the author macro to show all the info I want.
@@
 of data points), I have not implemented those optimizations for the auto-correlation calculations. 
-\subsection{Input Files}
+\subsection{Input File Formats}\label{section:formats}
 The codes currently can handle these types of input data files:
@@
 
-\flushleft The \texttt{fast-food} file consists of a header:
+The \texttt{fast-food} file consists of a header:
 \begin{lstlisting}[caption={The header format for \texttt{fast-food} files}]
@@
 \vspace{0.2in}
-\texttt{./logbins <rmin> <rmax> <nbins>   > <text filename>}
+\texttt{./logbins rmin rmax nbins   > filename}
 \vspace{0.2in}
@@
 \item \texttt{file1}   -- the file name for the first file.
-\item \texttt{format1} -- the file format for the first file.
+\item \texttt{format1} -- the file format for the first file. Options are \texttt{a,c,f} -- see Section~\ref{section:formats}.
 \item \texttt{file2}   -- the file name for the second file.
-\item \texttt{format2} -- the file format for the second file.
+\item \texttt{format2} -- the file format for the second file. Options are \texttt{a,c,f} -- see Section~\ref{section:formats}.
 \item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
@@
 \item \texttt{file}    -- file name.
-\item \texttt{format}  -- file format.
+\item \texttt{format}  -- file format. Options are \texttt{a,c,f} -- see Section~\ref{section:formats}.
 \item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
@@
 \subsection{Partitioning the Particles based on \texorpdfstring{\rmax}{rmax}}
+We need to compute pairwise distances to get the correlation function. A naive implementation of a correlation function would compute {\em all possible} 
+pairwise separations with a complexity $\mathcal{O}(N^2)$. However, for almost all correlation functions, we are only interested in separations less than 
+a certain \rmax, where \rmax is much smaller than the domain of the point distribution itself. We can then immediately see a way to prune pairs that can not
+{\em possibly} be within \rmax. If we impose a 3-d grid, with cell-size \rmax, then two points separated by more than one cell size (\rmax) in any one 
+dimension can not be within \rmax of each other (see Fig.~\ref{fig:grid} for a 2-D representation). Thus, given one point which is the target galaxy and a grid with cell-size \rmax, 
+immediately allows us to prune {\em all} of the points that are not within 1 cell offset in each dimension. However, even with this pruning, the actual 
+implementation of the algorithm matters. For instance, the non-AVX version of the codes tend to run $2-3\times$ slower than the AVX version. 
 \begin{figure}
@@
 If PERIODIC boundary conditions are enabled, then \texttt{off\_xwrap}, \texttt{off\_ywrap} and \texttt{off\_zwrap} have been declared and initialized in Listing~\ref{code:looping_over_neighbouring_cells}.
-By wrapping the elements if the first data-set, we can avoid these operations in the following \texttt{j-loop} over all particles in \texttt{second}.
-
+By wrapping the elements if the first data-set, we can avoid the wrapping operations in the \texttt{j-loop} over all particles in \texttt{second}. 
 \begin{lstlisting}[label={code:looping_over_second_particles},caption={AVX intrinsics for looping over all particles in the \texttt{second} cell. PERIODIC boundary conditions have 
@@
 \subsubsection{Pair-counting for \texorpdfstring{\xir}{xi(r)}}\label{section:pair_counting_r}
-Pair-counting for \xir is the most straight-forward. Once all the \texttt{NVEC} separations have been computed, we use bit-masks to check if any separations fall 
+Pair-counting for \xir is the most straight-forward. The squared 3-D separation, \texttt{r2}, is simply the sum of the squared differences in 
+each \texttt{X/Y/Z} dimensions. 
+\begin{lstlisting}[label={code:r2_xi},caption={Calculating squared separations in \xir.}]
+const AVX_FLOATS m_xdiff_sqr = AVX_SQUARE_FLOAT(m_xdiff);
+const AVX_FLOATS m_ydiff_sqr = AVX_SQUARE_FLOAT(m_ydiff);
+const AVX_FLOATS m_zdiff_sqr = AVX_SQUARE_FLOAT(m_zdiff);
+const AVX_FLOATS m_xydiff_sqr_sum = AVX_ADD_FLOATS(m_xdiff_sqr,m_ydiff_sqr);
+AVX_FLOATS r2 = AVX_ADD_FLOATS(m_zdiff_sqr,m_xydiff_sqr_sum);
+\end{lstlisting}
+Once all the \texttt{NVEC} separations have been computed, we use bit-masks to check if any separations fall 
 within the range \texttt{sqr\_rpmin} and \texttt{sqr\_rpmax}. If not, we \texttt{continue} with the \texttt{j-loop}. 
@@
 \end{lstlisting}
+If the code reaches past this bit-mask section, then at least one separation is within range. In that case, we use the 
+code in Listing~\ref{code:updating_histogram} to update the \texttt{npairs} array (and the \texttt{rpavg} array if 
+OUTPUT\_RPAVG is enabled).
+
+\subsubsection{Pair-separations for the projected correlation functions \texorpdfstring{\xirppi,\wprp}{xi(rp,pi), wp(rp)}}
+Listing~\ref{code:r2_projected} shows the squared distance calculation for \xirppi and \wprp. \texttt{r2} is simply 
+the projected separation in the \texttt{X-Y} plane.
+\begin{lstlisting}[label={code:r2_projected},caption={Calculating squared separations in \xirppi and \wprp.}]
+const AVX_FLOATS m_xdiff_sqr = AVX_SQUARE_FLOAT(m_xdiff);
+const AVX_FLOATS m_ydiff_sqr = AVX_SQUARE_FLOAT(m_ydiff);
+AVX_FLOATS r2 = AVX_ADD_FLOATS(m_xdiff_sqr,m_ydiff_sqr);
+\end{lstlisting}
 
 \subsubsection{Pair-counting for \texorpdfstring{\xirppi}{xi(rp,pi)}}\label{section:pair_counting_rp_pi}
+\begin{lstlisting}[label={code:masks_xirppi},caption={Bit-masks in \xirppi.}]
+const AVX_FLOATS m_mask_pimax = AVX_COMPARE_FLOATS(m_zdiff,m_pimax,_CMP_LT_OS);
+const int test = AVX_TEST_COMPARISON(m_mask_pimax);
+if(test == 0) {
+  continue;
+}
+
+const AVX_FLOATS m1 = AVX_COMPARE_FLOATS(r2,m_sqr_rpmin,_CMP_GE_OS);
+r2 = AVX_BLEND_FLOATS_WITH_MASK(m_sqr_rpmax,r2,m_mask_pimax);
+
+m_mask_left = AVX_COMPARE_FLOATS(r2,m_sqr_rpmax,_CMP_LT_OS);
+const AVX_FLOATS m_mask = AVX_BITWISE_AND(m1,m_mask_left);
+int test1 = AVX_TEST_COMPARISON(m_mask);
+if(test1 == 0) {
+  continue;
+}
+m_zdiff = AVX_BLEND_FLOATS_WITH_MASK(m_pimax, m_zdiff, m_mask);
+#ifdef OUTPUT_RPAVG
+union_mDperp.m_Dperp = AVX_SQRT_FLOAT(r2);
+#endif
+union_pibin.m_ibin = AVX_TRUNCATE_FLOAT_TO_INT(AVX_MULTIPLY_FLOATS(m_zdiff,m_inv_dpi));
+\end{lstlisting}
+In the \xirppi code, we have to update a two dimensional \texttt{npairs($r_p$,$\pi$)} array. This means we can not directly update 
+the \texttt{npairs} matrix and instead have to use a separate loop (this loop is typically only invoked for \xir and \wprp 
+when OUTPUT\_RPAVG is enabled). As a result of this extra loop, \xirppi is $2-3\times$ slower than \xir and \wprp.
+
+\begin{lstlisting}[label={code:updating_histogram_xirppi},caption={Updating the \texttt{npairs} matrix in \xirppi.}]
+for(int jj=0;jj<NVEC;jj++) {
+  int rpbin = union_rpbin.ibin[jj];
+  int pibin = union_pibin.ibin[jj];
+  int ibin = rpbin*(npibin+1) + pibin;
+  npairs[ibin]++;
+#ifdef OUTPUT_RPAVG
+  rpavg [ibin] += union_mDperp.Dperp[jj];
+#endif
+}
+\end{lstlisting}
 
 \subsubsection{Pair-counting for \texorpdfstring{\wprp}{wp(rp)}}\label{section:pair_counting_wp}
-Since the pair-counting in \wprp always assumes PERIODIC boundary conditions {\em and} only computes an auto-correlation, some more optimizations are possible in the
+Since the pair-counting in \wprp always assumes PERIODIC boundary conditions {\em and} only computes an auto-correlation, extra optimizations are possible in the
 \wprp calculation. First, each individual cell in the \texttt{struct lattice} elements are sorted based on their \texttt{Z} arrays. The \texttt{X/Y} arrays are 
 also re-ordered simultaneously. Once all the \texttt{Z} values inside a cell have been sorted, we can avoid double-counting the pairs by changing the \texttt{iiz loop}
-to be:\\
-\vspace{0.2in}
-\texttt{for(int iiz={\bf 0};iiz<=zbin\_refine\_factor;iiz++)} 
-\vspace{0.2in}
-\\instead of\\
-\vspace{0.2in}
-\texttt{for(int iiz={\bf -zbin\_refine\_factor};iiz<=zbin\_refine\_factor;iiz++)} 
-\vspace{0.2in}
-
+to be:
+\begin{lstlisting}[label={code:loop_optimization_wp},caption={Optimizing the loop over neighbouring cells \texttt{z} in \wprp.}]
+for(int iiz=0;iiz<=zbin_refine_factor;iiz++)
+
+(*@{\Large instead of }@*)
+
+for(int iiz=-zbin_refine_factor;iiz<=zbin_refine_factor;iiz++)
+\end{lstlisting}
 Another advantage of sorting the \texttt{Z} values is earlier termination inside the actual calculation. The Listing~\ref{code:calculating_separations_wp} shows 
@@
 In Fig.~\ref{fig:scaling_numpart}, we show the scaling for the three codes with the number of particles. For this scaling, we subsampled 
-the fiducial mock to attain 10 logarithmic steps in particle number ranging from $1.2\times10^4$ to $1.2\times10^6$.
-
+the fiducial mock to attain 10 logarithmic steps in particle number ranging from $1.2\times10^4$ to $1.2\times10^6$. All the timings are 
+generateed using 1 thread. 
 \begin{figure}[htbp]
 \includegraphics[clip=true,width=\linewidth]{timings_Mr19_numpart}%
-\caption{Scaling with particle number for \xir, \xirppi, \wprp }
+\caption{Scaling with particle number for \xir, \xirppi, \wprp. The timings are obtained 
+using 1 OpenMP thread. }
 \label{fig:scaling_numpart}
@@
 \subsection{Scaling with \texorpdfstring{\rmax}{rmax}}
+The code runtime increases drastically as the largest requested separation, \rmax increases. Roughly speaking, the runtimes scale 
+as $\mathcal{O}(r_{max}^2)$ with \xir showing the strongest dependence on \rmax. This is from the \pimax dependence of \xirppi and 
+\wprp. Even when \rmax is large, the effective number of particles per cell only grows as $r_{max}^2$ and not $r_{max}^3$ (as it does 
+for \xir). If we extend the benchmarks to larger \rmax, we will see a $\mathcal{O}(r_{max}^3)$ dependence for 
 \begin{figure}[htbp]
 \includegraphics[clip=true,width=\linewidth]{timings_Mr19_rmax}%
-\caption{Scaling with \rmax for \xir, \xirppi, \wprp }
+\caption{Scaling with \rmax for \xir, \xirppi, \wprp. The timings are obtained 
+with 4 OpenMP threads. }
 \label{fig:scaling_rmax}
@@
 \subsection{Scaling with OpenMP threads}
+All of the codes presented here scale reasonably well (efficiency $\gtrsim$ 80\%) up to 10 OpenMP threads. Beyond that, the work-load is 
+scaling efficiency starts dropping off and plateaus for \texttt{nthreads} $\gtrsim 20$. 
 \begin{figure}[htbp]
 \includegraphics[clip=true,width=\linewidth]{timings_Mr19_openmp}%
-\caption{OpenMP scaling for \xir, \xirppi, \wprp }
+\caption{OpenMP scaling for \xir, \xirppi, \wprp. The fiducial mock used here contains $\sim 10^6$ particles 
+in a $420.0$ \hMpc cube. }
 \label{fig:scaling_openmp}
 \end{figure}
+
+\begin{table}
+\centering
+\caption{\footnotesize OpenMP scaling for the three codes. Efficiencies are defined 
+as \texttt{Speedup/Nthreads}. Note, that \xir has super-linear scaling with \texttt{nthreads}.}
+\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{cccc} 
+\toprule
+\multirow{2}{*}{\textbf{Nthreads}}   &
+\multicolumn{3}{c}{\textbf{Efficiency[\%]}} \\
+\cmidrule(l{1.0em}r{1.0em}){2-4}
+&
+\multicolumn{1}{c}{$\boldsymbol{\xir}$}   &
+\multicolumn{1}{c}{$\boldsymbol{\xirppi}$} &
+\multicolumn{1}{c}{$\boldsymbol{\wprp}$}     \\
+\midrule
+\input{timings_Mr19_openmp} 
+\bottomrule
+\end{tabular}
+\end{adjustbox}
+\label{table:openmp}
+\end{table}
 

### 2015-01-30 utils/cellarray.h
Epic re-write of the io.c file. Can read in arbitrary numbers of fields of DOUBLE precision

@@ -1,0 +1,32 @@
+/* File: cellarray.h */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#pragma once
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <inttypes.h>
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "function_precision.h" //definition of DOUBLE depending on Makefile flag -DOUBLE_PREC
+
+#define NLATMAX   100      /* maximum grid dimension in X-Y plane */
+
+typedef struct{
+  DOUBLE *x;
+  DOUBLE *y;
+  DOUBLE *z;
+  int64_t nelements;
+} cellarray;
+
+#ifdef __cplusplus
+}
+#endif

### 2015-01-30 utils/gridlink.c

@@ -1,0 +1,185 @@
+/* File: gridlink.c */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+
+#include "defs.h"
+#include "gridlink.h"
+#include "utils.h"
+#include "function_precision.h"
+#include "cellarray.h"
+
+#define MEMORY_INCREASE_FAC   1.2
+
+double get_binsize(const double xmin,const double xmax, const double rmax, const int refine_factor, const int max_ncells, int *nlattice)  __attribute__((warn_unused_result));
+
+double get_binsize(const double xmin,const double xmax, const double rmax, const int refine_factor, const int max_ncells, int *nlattice)
+{
+  double xdiff = xmax-xmin;
+  int nmesh=(int)(refine_factor*xdiff/rmax) ;
+#ifdef PERIODIC
+  if (nmesh<(2*refine_factor+1))  {
+    fprintf(stderr,"linklist> ERROR:  nlattice = %d is so small that with periodic wrapping the same cells will be counted twice ....exiting\n",nmesh) ;
+    exit(EXIT_FAILURE) ;
+  }
+#endif
+  
+  if (nmesh>max_ncells)  nmesh=max_ncells;
+  double xbinsize = xdiff/nmesh;
+  *nlattice = nmesh;
+  return xbinsize;
+}
+
+void get_max_min(const int64_t ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
+				 DOUBLE *min_x, DOUBLE *min_y, DOUBLE *min_z, DOUBLE *max_x, DOUBLE *max_y, DOUBLE *max_z)
+{
+  DOUBLE xmin = *min_x, ymin = *min_y, zmin=*min_z;
+  DOUBLE xmax = *max_x, ymax = *max_y, zmax=*max_z;
+	
+  for(int64_t i=0;i<ND1;i++) {
+    if(X1[i] < xmin) xmin=X1[i];
+    if(Y1[i] < ymin) ymin=Y1[i];
[…]

### 2015-01-30 utils/sglib.h

@@ -1,0 +1,1960 @@
+/* File: sglib.h */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+/* 
+
+  This is SGLIB version 1.0.3
+
+  (C) by Marian Vittek, Bratislava, http://www.xref-tech.com/sglib, 2003-5
+
+  License Conditions: You can use a verbatim copy (including this
+  copyright notice) of sglib freely in any project, commercial or not.
+  You can also use derivative forms freely under terms of Open Source
+  Software license or under terms of GNU Public License.  If you need
+  to use a derivative form in a commercial project, or you need sglib
+  under any other license conditions, contact the author.
+
+
+
+*/
+
+
+#ifndef _SGLIB__h_
+#define _SGLIB__h_
+
+/* the assert is used exclusively to write unexpected error messages */
+#include <assert.h>
+
+
+/* ---------------------------------------------------------------------------- */
+/* ---------------------------------------------------------------------------- */
+/* -                            LEVEL - 0  INTERFACE                          - */
+/* ---------------------------------------------------------------------------- */
+/* ---------------------------------------------------------------------------- */
+
+
+/* ---------------------------------------------------------------------------- */
+/* ------------------------------ STATIC ARRAYS ------------------------------- */
+/* ---------------------------------------------------------------------------- */
+
+/* 
+
+  Basic algorithms  for sorting arrays. Multiple  depending arrays can
+  be rearranged using user defined 'elem_exchangers'
+
+*/
+
+/*               HEAP - SORT  (level 0)           */
+
[…]

### 2015-01-30 xi_theory/examples/run_correlations.c

@@ -1,0 +1,241 @@
+/* File: run_correlations.c */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+/*
+	Example code to show how to use the correlation function libraries
+	Author: Manodeep Sinha <manodeep@gmail.com>
+	Date: At some point in early 2015. 
+
+
+
+*/
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <assert.h>
+#include <sys/time.h>
+
+#include "function_precision.h"
+#include "io.h"
+#include "defs.h"
+#include "utils.h"
+
+/* Library proto-types + struct definitions in the ../include directory */
+#include "countpairs.h"
+#include "countpairs_rp_pi.h"
+#include "countpairs_wp.h"
+
+#ifndef MAXLEN
+#define MAXLEN 500
+#endif
+
+//Just to output some colors
+
+#define ANSI_COLOR_RED     "\x1b[31m"
+#define ANSI_COLOR_GREEN   "\x1b[32m"
+#define ANSI_COLOR_RESET   "\x1b[0m"
+#define ANSI_COLOR_BLUE    "\x1b[34m"
+#define ANSI_COLOR_MAGENTA "\x1b[35m"
+
+/* #define ANSI_COLOR_YELLOW  "\x1b[33m" */
+/* #define ANSI_COLOR_CYAN    "\x1b[36m" */
+
+
+
+void Printhelp(void);
+
+void Printhelp(void)
+{
+  fprintf(stderr,ANSI_COLOR_RED "=========================================================================\n") ;
+  fprintf(stderr,"   --- run_correlations file format binfile boxsize [Nthreads] \n") ;
+  fprintf(stderr,"   --- Measure the auto-correlation functions DD(r), DD(rp, pi) and wp(rp) for a single file\n");
+  fprintf(stderr,"     * file         = name of data file\n") ;
+  fprintf(stderr,"     * format       = format of data file  (a=ascii, f=fast-food)\n") ;
[…]

### 2015-01-30 xi_theory/tests/test_periodic.c

@@ -1,0 +1,317 @@
+/* File: test_periodic.c */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <assert.h>
+#include <string.h>
+#include <sys/time.h>
+
+#ifndef MAXLEN
+#define MAXLEN 500
+#endif
+
+#ifndef PERIODIC
+#define PERIODIC 
+#endif
+
+#ifndef DOUBLE_PREC
+#define DOUBLE_PREC
+#endif
+
+#ifndef OUTPUT_RPAVG
+#define OUTPUT_RPAVG
+#endif
+
+#ifndef SILENT
+#define SILENT
+#endif
+
+#include "function_precision.h"
+#include "io.h"
+#include "defs.h"
+#include "utils.h"
+
+
+//Including the C files directly
+#include "../utils/gridlink.c"
+#include "../io/io.c"
+#include "../io/ftread.c"
+#include "../xi_of_r/countpairs.c"
+#include "../xi_rp_pi/countpairs_rp_pi.c"
+#include "../wp/countpairs_wp.c"
+
+char tmpoutputfile[]="./test_periodic_output.txt";
+
+int test_periodic_DD(const char *correct_outputfile);
+int test_periodic_DDrppi(const char *correct_outputfile);
+int test_wp(const char *correct_outputfile);
+void read_data_and_set_globals(const char *firstfilename, const char *firstformat,const char *secondfilename,const char *secondformat);
+
+//Global variables
+int ND1;
+DOUBLE *X1=NULL,*Y1=NULL,*Z1=NULL;
+
+int ND2;
+DOUBLE *X2=NULL,*Y2=NULL,*Z2=NULL;
+
+char binfile[]="bins";
+DOUBLE pimax=40.0;
+double boxsize=420.0;
+#ifdef USE_OMP
+const int nthreads=4;
+#endif
+
+char current_file1[MAXLEN],current_file2[MAXLEN];
+
+//end of global variables
+
+
+int test_periodic_DD(const char *correct_outputfile)
+{
+  int autocorr = (X1==X2) ? 1:0;
+
[…]

### 2015-01-30 io/ftread.c
Epic re-write of the io.c file. Can read in arbitrary numbers of fields of DOUBLE precision

@@ -1,0 +1,67 @@
+/* File: ftread.c */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+/*	ftread reads unformatted data that has been written by fortran
+	or in fortran convention -- i.e. an integer with the number of
+	bytes in the record, the record data, and another integer with
+	the number of bytes in the record.  ftread does various error
+	checks, writing a message to stderr and returning a negative 
+	value if an error or warning occurs.  The call is identical to
+	the standard i/o library routine fread.
+*/
+
+#include "ftread.h"
+
+
+
+int my_ftread(void *ptr,size_t size,size_t nitems,FILE *stream)
+{
+  int err = ftread(ptr,size,nitems,stream);
+  if(err != 0){
+    fprintf(stderr,"ERROR: Read error in ftread ..exiting\n");
+    exit(EXIT_FAILURE);
+  }
+
+  return 0;
+}
+
+
+int ftread(void *ptr,size_t size,size_t nitems,FILE *stream)
+{
+  unsigned int nbyte1, nbyte2;
+  size_t nitem1 ;
+  int errno ;
+  errno = 0 ;
+  if ( fread(&nbyte1,sizeof(nbyte1),1,stream) != 1 )  {
+    errno = -10 ;
+    fprintf(stderr,"read error, file empty ? \n") ;
+  }
+  nitem1 = fread(ptr,size,nitems,stream) ;
+  if ( nitem1 != nitems ) {
+    errno = -20 ;
+    fprintf(stderr,"read error, %zu items expected, %zu items read. \n",
+	    nitems,nitem1) ;
+  }
+  if ( fread(&nbyte2,sizeof(nbyte2),1,stream) != 1 )  {
+    errno = -30 ;
+    fprintf(stderr,"read error, file too short ? \n") ;
+  }
+  if ( nbyte1 != nbyte2 ) {
+    errno = errno - 1 ;
+    fprintf(stderr,"read warning, byte # do not match, nbyte1 = %d, nbyte2 = %d \n",
+	    nbyte1,nbyte2) ;
+  }
+  if ( nbyte1 != size*nitems) {
+    errno = errno - 2 ;
[…]

### 2015-01-30 io/ftread.h

@@ -1,0 +1,26 @@
+/* File: ftread.h */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#ifndef __FTREAD_H
+#define __FTREAD_H
+
+#include <stdio.h>
+#include <stdlib.h>
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+int ftread(void *ptr,size_t size,size_t nitems,FILE *stream);
+int my_ftread(void *ptr,size_t size,size_t nitems,FILE *stream);
+
+#ifdef __cplusplus
+}
+#endif
+	
+#endif

### 2015-01-30 io/io.h

@@ -1,0 +1,25 @@
+/* File: io.h */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#pragma once
+#include <stdio.h>
+#include <stdlib.h>
+#include <inttypes.h>
+#include <stdarg.h>
+#include <string.h>
+#include <stddef.h>
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+int64_t read_positions(const char *filename, const char *format, const size_t size, const int num_fields, ...) __attribute__((warn_unused_result));
+
+#ifdef __cplusplus
+}
+#endif

### 2015-01-30 utils/defs.h

@@ -1,0 +1,14 @@
+/* File: defs.h */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#pragma once
+
+
+#define ADD_DIFF_TIME(t0,t1)     ((t1.tv_sec - t0.tv_sec) + 1e-6*(t1.tv_usec - t0.tv_usec))
+
+

### 2015-01-30 utils/gridlink.h

@@ -1,0 +1,40 @@
+/* File: gridlink.h */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#pragma once
+
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "function_precision.h"
+#include "cellarray.h"
+
+cellarray * gridlink(const int64_t np,
+					 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
+					 const DOUBLE xmin, const DOUBLE xmax,
+					 const DOUBLE ymin, const DOUBLE ymax,
+					 const DOUBLE zmin, const DOUBLE zmax,
+					 const DOUBLE max_x_size,
+					 const DOUBLE max_y_size,
+					 const DOUBLE max_z_size,
+					 const int xbin_refine_factor,
+					 const int ybin_refine_factor,
+					 const int zbin_refine_factor,
+					 int *nlattice_x,
+					 int *nlattice_y,
+					 int *nlattice_z) __attribute__((warn_unused_result)); 
+
+void get_max_min(const int64_t ND1, const DOUBLE * restrict X1, const DOUBLE * restrict Y1, const DOUBLE * restrict Z1,
+				 DOUBLE *min_x, DOUBLE *min_y, DOUBLE *min_z, DOUBLE *max_x, DOUBLE *max_y, DOUBLE *max_z);
+
+	
+#ifdef __cplusplus
+}
+#endif

### 2015-01-30 xi_theory/logbins.c

@@ -1,0 +1,58 @@
+/* File: logbins.c */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+/* PROGRAM logbins
+
+   --- logbins xmin xmax nbin
+   --- Outputs the bins for a logarithmic bin spacing
+
+      * xmin   = inner radius of smallest bin
+      * xmax   = outer radius of largest bin
+      * nbin  = number of bins (logarithmically spaced in theta)
+      ----------------------------------------------------------------------
+*/
+
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+#include <assert.h>
+
+int main(int argc, char *argv[])
+{
+  /*---Arguments-------------------------*/
+  int nbin ;
+  double xmin,xmax,logxmin,logxmax,dlogx;
+
+  if(argc<4) {
+    fprintf(stderr,"Usage `%s' rmin rmax nbins \n",argv[0]);
+    exit(EXIT_FAILURE);
+  } else {
+    /*---Read-arguments-----------------------------------*/
+    sscanf(argv[1],"%lf",&xmin) ;
+    sscanf(argv[2],"%lf",&xmax) ;
+    sscanf(argv[3],"%d",&nbin) ;
+  }
+	assert(xmin > 0 && "rmin has to be non-zero, otherwise log is undefined");
+
+  /*---Setup-Pairs-arrays-------------------------------*/
+  logxmin = log10(xmin) ;
+  logxmax = log10(xmax) ;
+  dlogx = (logxmax-logxmin)/(double)nbin ;
+
+  double left_edge=logxmin;
+  for(int i=0;i<nbin;i++){
+    double right_edge=left_edge+dlogx; 
+    fprintf(stdout,"%20.12lf %20.12lf\n",pow(10.0,left_edge),pow(10.0,right_edge)) ;
+    left_edge=right_edge;
+  }
+  exit(EXIT_SUCCESS);
+}
+
+
+
+

### 2015-01-30 xi_theory/wp/countpairs_wp.h

@@ -1,0 +1,43 @@
+/* File: countpairs_wp.h */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#pragma once
+
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "function_precision.h"
+#include <inttypes.h>
+
+//define the results structure
+typedef struct{
+	uint64_t *npairs;
+	DOUBLE *wp;
+	DOUBLE *rupp;
+	DOUBLE *rpavg;
+	DOUBLE pimax;
+	int nbin;
+} results_countpairs_wp;
+
+results_countpairs_wp *countpairs_wp(const int64_t ND1, DOUBLE * restrict X1, DOUBLE * restrict Y1, DOUBLE * restrict Z1,
+									 const double boxsize, 
+#ifdef USE_OMP
+									 const int numthreads,
+#endif
+									 const char *binfile,
+									 const double pimax) __attribute__((warn_unused_result));
+
+
+void free_results_wp(results_countpairs_wp **results);
+
+#ifdef __cplusplus
+}
+#endif
+	

### 2015-01-30 xi_theory/xi_rp_pi/countpairs_rp_pi.c

@@ -1,0 +1,533 @@
+/* File: countpairs_rp_pi.c */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#include <stdio.h>
+#include <math.h>
+#include <stdlib.h>
+
+#include "countpairs_rp_pi.h" //function proto-type
+#include "gridlink.h"//function proto-type for gridlink
+#include "cellarray.h" //definition of struct cellarray
+#include "utils.h" //all of the utilities
+#include "progressbar.h" //for the progressbar
+
+#ifdef USE_AVX
+#include "avx_calls.h"
+#endif
+
+#ifdef USE_OMP
+#include <omp.h>
+#endif
+
+
+void free_results_rp_pi(results_countpairs_rp_pi **results)
+{
+	if(results==NULL)
+		return;
+
+	if(*results==NULL)
+		return;
+	
+	results_countpairs_rp_pi *tmp = *results;
+
+	free(tmp->npairs);
+	free(tmp->rupp);
+	free(tmp->rpavg);
+	free(tmp);
+	tmp = NULL;
+}
+
+
+results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
+											const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
+#ifdef USE_OMP
+											const int numthreads,
+#endif
+											const int autocorr,
+											const char *binfile,
+											const double pimax)
+{
+  int bin_refine_factor=1;
+	int zbin_refine_factor=2;
+	if(autocorr==1) {
+		bin_refine_factor=2;
+		zbin_refine_factor=2;
+	} else {
+		bin_refine_factor=1;
+		zbin_refine_factor=1;
+	}
+#ifdef USE_OMP
+	if(numthreads > 1) {
+		if(autocorr==1) {
+			bin_refine_factor=2;
[…]

### 2015-01-30 xi_theory/xi_rp_pi/countpairs_rp_pi.h

@@ -1,0 +1,41 @@
+/* File: countpairs_rp_pi.h */
+/*
+		This file is a part of the Corrfunc package
+		Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+		License: MIT LICENSE. See LICENSE file under the top-level
+		directory at https://bitbucket.org/manodeep/corrfunc/
+*/
+
+#pragma once
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "function_precision.h" //for definition of DOUBLE
+#include <inttypes.h> //for uint64_t
+
+//define the results structure
+typedef struct{
+	uint64_t *npairs;
+	DOUBLE *rupp;
+	DOUBLE *rpavg;
+	DOUBLE pimax;
+	int nbin;
+	int npibin;
+} results_countpairs_rp_pi;
+
+results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
+											const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
+#ifdef USE_OMP
+											const int numthreads,
+#endif
+											const int autocorr,
+											const char *binfile,
+											const double pimax)  __attribute__((warn_unused_result));
+
+void free_results_rp_pi(results_countpairs_rp_pi **results);
+
+#ifdef __cplusplus
+}
+#endif

### 2015-01-30 io/io.c
Epic re-write of the io.c file. Can read in arbitrary numbers of fields of DOUBLE precision

@@ -9,15 +9,19 @@
 #include "io.h"
 #include "ftread.h"
 #include "utils.h"
+#include "function_precision.h"
 
 #ifndef MEMORY_INCREASE_FAC
 #define MEMORY_INCREASE_FAC 1.1
 #endif
 
-int64_t read_positions(const char *filename, const char *format, void **xpos, void **ypos, void **zpos, const size_t size)
+int64_t read_positions(const char *filename, const char *format, const size_t size, const int num_fields, ...)
 {
   int64_t np;
-  void *x,*y,*z;
+  assert(num_fields >= 1 && "You have to request at least one field to read-in");
+  assert(size == sizeof(DOUBLE) && "Requested size of an item does not match sizeof(DOUBLE)");
+
+  DOUBLE *data[num_fields];
   if(strncmp(format,"f",1)==0) { /*Read-in fast-food file*/
 	//read fast-food file
 	size_t bytes=sizeof(int) + sizeof(float)*9 + sizeof(int);//skip fdat
@@ -25,13 +29,13 @@ int64_t read_positions(const char *filename, const char *format, void **xpos, vo
 	int idat[5];
 	FILE *fp = my_fopen(filename,"r");
 	my_ftread(idat,sizeof(int),5,fp);
-	np = (int64_t) idat[1]; //idat[1] is int
+	np = (int64_t) idat[1]; //idat[1] is int. 
 	
 	assert((size == 4 || size == 8) && "Size of each position element can be either 4 (float) or 8 (double)");
 	
-	x = my_malloc(size,np);
-	y = my_malloc(size,np);
-	z = my_malloc(size,np);
+	for(int i=0;i<num_fields;i++) {
+	  data[i] = my_malloc(size,np);
+	}
 	
 	my_fseek(fp,bytes,SEEK_CUR);
 	//Check that the file was written with the requested precision
@@ -43,9 +47,9 @@ int64_t read_positions(const char *filename, const char *format, void **xpos, vo
[…]

### 2015-01-30 io/io.h

@@ -10,13 +10,15 @@
 #include <stdio.h>
 #include <stdlib.h>
 #include <inttypes.h>
-
+#include <stdarg.h>
+#include <string.h>
+#include <stddef.h>
 
 #ifdef __cplusplus
 extern "C" {
 #endif
 
-int64_t read_positions(const char *filename, const char *format, void **xpos, void **ypos, void **zpos, const size_t size);
+int64_t read_positions(const char *filename, const char *format, const size_t size, const int num_fields, ...) __attribute__((warn_unused_result));
 
 #ifdef __cplusplus
 }

### 2015-01-30 xi_theory/examples/run_correlations.c

@@ -121,7 +121,7 @@ int main(int argc, char **argv)
 	fprintf(stderr,"\t\t -------------------------------------" ANSI_COLOR_RESET "\n");
 
 	//Read-in the data
-	const int64_t ND1 = read_positions(file,fileformat,(void **) &x1,(void **) &y1,(void **) &z1,sizeof(DOUBLE));
+	const int64_t ND1 = read_positions(file,fileformat,sizeof(DOUBLE),3, &x1, &y1, &z1);
 
 	int autocorr=1;
 	DOUBLE *x2 = x1;

### 2015-01-30 xi_theory/xi_of_r/DD.c

@@ -94,7 +94,7 @@ int main(int argc, char *argv[])
 
   /*---Read-data1-file----------------------------------*/
   gettimeofday(&t0,NULL);
-  ND1=read_positions(file1,fileformat1,(void **) &x1,(void **) &y1,(void **) &z1,sizeof(DOUBLE));
+  ND1=read_positions(file1,fileformat1, sizeof(DOUBLE), 3, &x1, &y1, &z1);
   gettimeofday(&t1,NULL);
   read_time += ADD_DIFF_TIME(t0,t1);
 
@@ -106,7 +106,7 @@ int main(int argc, char *argv[])
   gettimeofday(&t0,NULL);  
   if (autocorr==0) {
     /*---Read-data2-file----------------------------------*/
-		ND2=read_positions(file2,fileformat2,(void **) &x2,(void **) &y2,(void **) &z2,sizeof(DOUBLE));
+	ND2=read_positions(file2,fileformat2, sizeof(DOUBLE), 3, &x2, &y2, &z2);
     gettimeofday(&t1,NULL);
     read_time += ADD_DIFF_TIME(t0,t1);
   } else {

### 2015-01-30 xi_theory/xi_rp_pi/DDrppi.c

@@ -128,14 +128,14 @@ int main(int argc, char *argv[])
 
   gettimeofday(&t0,NULL);
   /*---Read-data1-file----------------------------------*/
-  ND1=read_positions(file1,fileformat1,(void **) &x1,(void **) &y1,(void **) &z1,sizeof(DOUBLE));
+  ND1=read_positions(file1,fileformat1,sizeof(DOUBLE), 3, &x1, &y1, &z1);
   gettimeofday(&t1,NULL);
   read_time += ADD_DIFF_TIME(t0,t1);
   gettimeofday(&t0,NULL);  
 
   if (autocorr==0) {
     /*---Read-data2-file----------------------------------*/
-		ND2=read_positions(file2,fileformat2,(void **) &x2,(void **) &y2,(void **) &z2,sizeof(DOUBLE));
+	ND2=read_positions(file2,fileformat2,sizeof(DOUBLE), 3, &x2, &y2, &z2);
     gettimeofday(&t1,NULL);
     read_time += ADD_DIFF_TIME(t0,t1);
 

### 2015-02-14 paper/users-guide.tex
It was impressed upon me that a data correlation function is also required. So, I am adding those as well. Means restructuring of the code directories. io has been moved one level up.
Fairly large internal re-organization for the lattice structure. In some cases, it is faster to store the positions in AVX chunks (8 for floats/4 for doubles).

@@
 .2 paper.
+.2 io\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files for reading in data{.} 
+  \end{minipage}.
+.2 utils\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files for creating grids and helper routines{.} 
+  \end{minipage}.
+.2 bin\DotFill 
+  \begin{minipage}[t]{8cm}
+    Will be created to copy executable files when you run `make install'{.} 
+  \end{minipage}.
+.2 include\DotFill 
+  \begin{minipage}[t]{8cm}
+    Header files for static libraries{.} 
+  \end{minipage}.
+.2 lib\DotFill 
+  \begin{minipage}[t]{8cm}
+    Will be created to copy static libraries and python library after you run `make'{.} 
+  \end{minipage}.
 .2 xi\_theory.
@@
   \end{minipage}.
-.3 bin\DotFill 
-  \begin{minipage}[t]{8cm}
-    Will be created to copy executable files when you run `make install'{.} 
-  \end{minipage}.
 .3 examples\DotFill 
@@
   \end{minipage}.
-.3 include\DotFill 
-  \begin{minipage}[t]{8cm}
-    Header files for static libraries{.} 
-  \end{minipage}.
-.3 io\DotFill 
-  \begin{minipage}[t]{8cm}
-    Source files for reading in data{.} 
-  \end{minipage}.
-.3 lib\DotFill 
-  \begin{minipage}[t]{8cm}
-    Will be created to copy static libraries and python library after you run `make'{.} 
-  \end{minipage}.
 .3 python\_bindings\DotFill 
@@
   \end{minipage}.
-.3 utils\DotFill 
-  \begin{minipage}[t]{8cm}
-    Source files for creating 3-D grid and helper routines{.} 
-  \end{minipage}.
 .3 xi\_of\_r\DotFill 
   \begin{minipage}[t]{8cm}
-    Source files for $\xi(r)${.} 
+    Source files for getting pair-counts to compute $\xi(r)${.} 
   \end{minipage}.
@@
     Source files for $w_p(r_p)${.} 
+  \end{minipage}.
+.3 xi\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files for $\xi(r)$ directly on a periodic cube{.} 
+  \end{minipage}.
+.3 vpf\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files to compute the counts-in-spheres statistic{.} 
+  \end{minipage}.
+
+.2 xi\_mocks.
+.3 benchmarks\DotFill
+  \begin{minipage}[t]{8cm}
+    IDL scripts to run benchmarks for mocks{.} 
+  \end{minipage}.
+.3 examples\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files for example C bindings using the static libraries for mocks{.} 
+  \end{minipage}.
+.3 python\_bindings\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files to generate python bindings for mocks{.} 
+  \end{minipage}.
+.3 tests\DotFill 
+  \begin{minipage}[t]{8cm}
+    Correct outputs for tests for mocks{.}
+  \end{minipage}.
+.4 data\DotFill 
+  \begin{minipage}[t]{8cm}
+    Mock galaxy catalogs for tests on a sphere{.} 
+  \end{minipage}.
+.3 DDrppi\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files for getting pair-counts to compute $\xi(r_p,\pi)$ on mocks{.} 
+  \end{minipage}.
+.3 wtheta\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files for getting pair-counts to compute $\omega(\theta)$ {.} 
+  \end{minipage}.
+.3 vpf\DotFill 
+  \begin{minipage}[t]{8cm}
+    Source files to compute the counts-in-spheres statistic on mocks{.} 
   \end{minipage}.
@@
 \caption{\footnotesize OpenMP scaling for the three codes. Efficiencies are defined 
-as \texttt{Speedup/Nthreads}. Note, that \xir has super-linear scaling with \texttt{nthreads}.}
+as \texttt{Speedup/Nthreads}. Note, that \xir has super-linear scaling with \texttt{nthreads}
+up to $\sim$ 8 threads.}
 \begin{adjustbox}{max width=\textwidth}

### 2015-02-14 xi_mocks/wtheta/countpairs_theta_mocks.c

@@ -14,6 +14,7 @@
 #include <stdlib.h>
 #include <math.h>
 
+#include "defs.h"
 #include "gridlink_mocks.h"//function proto-type for gridlink
 #include "countpairs_theta_mocks.h" //function proto-type
 #include "cellarray_mocks.h" //definition of struct cellarray_mocks
@@ -91,7 +92,6 @@ results_countpairs_theta * countpairs_theta_mocks(const int64_t ND1, DOUBLE *phi
 																									const char *binfile)
 {
 
-  fprintf(stderr,"Running in `%s' mode \n", autocorr == 1 ? "auto-correlation":"cross-correlation");
 #if defined(LINK_IN_RA) && !defined(LINK_IN_DEC)
 #error LINK_IN_DEC Makefile option must be enabled before LINK_IN_RA is selected
 #endif 
@@ -127,6 +127,12 @@ results_countpairs_theta * countpairs_theta_mocks(const int64_t ND1, DOUBLE *phi
 
 #ifdef LINK_IN_DEC
   int rbin_refine_factor=2;
+#ifdef USE_OMP
+	if(numthreads > 1) {
+		rbin_refine_factor=numthreads;
+	}
+#endif
+	
 #ifdef LINK_IN_RA
 	//roughly occurs around
 	int phi_bin_refine_factor=2;
@@ -192,7 +198,7 @@ results_countpairs_theta * countpairs_theta_mocks(const int64_t ND1, DOUBLE *phi
   x2=my_malloc(sizeof(*x2),ND2);
   y2=my_malloc(sizeof(*y2),ND2);
   z2=my_malloc(sizeof(*z2),ND2);
-  
+
   for(int i=0;i<ND2;i++) {
     x2[i] = COSD(theta2[i])*COSD(phi2[i]) ;
     y2[i] = COSD(theta2[i])*SIND(phi2[i]) ;
@@ -204,13 +210,36 @@ results_countpairs_theta * countpairs_theta_mocks(const int64_t ND1, DOUBLE *phi
     if(theta2[i] > dec_max)
       dec_max = theta2[i];
 #endif
-
   }
 
+	DOUBLE *x1,*y1,*z1;
 
+	if (autocorr==0) {
+		x1 = my_malloc(sizeof(*x1),ND1);
+		y1 = my_malloc(sizeof(*y1),ND1);
+		z1 = my_malloc(sizeof(*z1),ND1);
+		
+		for(int i=0;i<ND1;i++) {
[…]

### 2015-02-14 xi_theory/tests/test_periodic.c

@@ -198,7 +198,8 @@ int test_vpf(const char *correct_outputfile)
 
 int test_xi(const char *correct_outputfile)
 {
-  results_countpairs_xi *results = countpairs_xi(ND1,X1,Y1,Z1,
+
+	results_countpairs_xi *results = countpairs_xi(ND1,X1,Y1,Z1,
 																								 boxsize,
 #ifdef USE_OMP
 																								 nthreads,

### 2015-02-14 xi_theory/vpf/countspheres.c

@@ -74,7 +74,7 @@ results_countspheres * countspheres(const int64_t np, const DOUBLE * restrict X,
   //First create the 3-d linklist
   int nmesh_x=0,nmesh_y=0,nmesh_z=0;
   gettimeofday(&t0,NULL);
-	cellarray *lattice = gridlink(np, X, Y, Z, xmin, xmax, ymin, ymax, zmin, zmax, rmax, rmax, rmax, bin_refine_factor, bin_refine_factor, bin_refine_factor, &nmesh_x, &nmesh_y, &nmesh_z);
+	cellarray_nvec *lattice = gridlink_nvec(np, X, Y, Z, xmin, xmax, ymin, ymax, zmin, zmax, rmax, rmax, rmax, bin_refine_factor, bin_refine_factor, bin_refine_factor, &nmesh_x, &nmesh_y, &nmesh_z);
   gettimeofday(&t1,NULL);
 	int64_t totncells = (int64_t) nmesh_x * (int64_t) nmesh_y * (int64_t) nmesh_z; 
 
@@ -188,24 +188,28 @@ results_countspheres * countspheres(const int64_t np, const DOUBLE * restrict X,
 #endif
 
 					const int64_t index=iiix*nmesh_y*nmesh_z + iiiy*nmesh_z + iiiz;
-					const cellarray *first = &(lattice[index]);
-					const DOUBLE *x2 = first->x;
-					const DOUBLE *y2 = first->y;
-					const DOUBLE *z2 = first->z;
+					const cellarray_nvec *first = &(lattice[index]);
+					DOUBLE *x2 = first->pos;
+					DOUBLE *y2 = first->pos + NVEC;
+					DOUBLE *z2 = first->pos + 2*NVEC;
 #ifndef USE_AVX
+
 					for(int64_t j=0;j<first->nelements;j+=NVEC) {
 						int block_size=first->nelements - j;
 						if(block_size > NVEC) block_size=NVEC;
 						for(int jj=0;jj<block_size;jj++) {
 							int ibin;
-							DOUBLE dx=x2[j+jj]-newxpos;
-							DOUBLE dy=y2[j+jj]-newypos;
-							DOUBLE dz=z2[j+jj]-newzpos;
+							DOUBLE dx=x2[jj]-newxpos;
+							DOUBLE dy=y2[jj]-newypos;
[…]

### 2015-02-14 xi_theory/xi/countpairs_xi.c

@@ -10,6 +10,7 @@
 #include <math.h>
 #include <stdlib.h>
 
+#include "defs.h"
 #include "gridlink.h"//function proto-type for gridlink
 #include "countpairs_xi.h" //function proto-type
 #include "cellarray.h" //definition of struct cellarray
@@ -58,7 +59,8 @@ results_countpairs_xi *countpairs_xi(const int64_t ND1, DOUBLE * restrict X1, DO
 	//How many bins to subdivide rmax into -> affects runtime on O(20-30%) levels.
 	//Check with your typical use-case and set appropriately. Values of 1,2 and 3 are 
 	//all you might need to check. 
-  int bin_refine_factor=2;
+  int bin_refine_factor=1;
+
 	
   /***********************
    *initializing the  bins
@@ -70,7 +72,6 @@ results_countpairs_xi *countpairs_xi(const int64_t ND1, DOUBLE * restrict X1, DO
   assert(rpmin > 0.0 && rpmax > 0.0 && rpmin < rpmax && "[rpmin, rpmax] are valid inputs");
   assert(nrpbin > 0 && "Number of rp bins must be > 0");
   
-
   /*---Create 3-D lattice--------------------------------------*/
   int nmesh_x=0,nmesh_y=0,nmesh_z=0;
   const DOUBLE xmin = 0.0, xmax=boxsize;
@@ -91,29 +92,46 @@ results_countpairs_xi *countpairs_xi(const int64_t ND1, DOUBLE * restrict X1, DO
 		lattice = gridlink(ND1, X1, Y1, Z1, xmin, xmax, ymin, ymax, zmin, zmax, rpmax, rpmax, rpmax, bin_refine_factor, bin_refine_factor, bin_refine_factor, &nmesh_x, &nmesh_y, &nmesh_z);
   }
   const int64_t totncells = (int64_t) nmesh_x * (int64_t) nmesh_y * (int64_t) nmesh_z;  		
+
+
 #ifdef USE_OMP
   omp_set_num_threads(numthreads);
 #pragma omp parallel for schedule(dynamic)
 #endif
 	for(int64_t icell=0;icell<totncells;icell++) {
 		const cellarray *first=&(lattice[icell]);
-		DOUBLE *x = first->x;
-		DOUBLE *y = first->y;
-		DOUBLE *z = first->z;
[…]

### 2015-02-14 utils/gridlink.h
Fairly large internal re-organization for the lattice structure. In some cases, it is faster to store the positions in AVX chunks (8 for floats/4 for doubles).

@@ -17,20 +17,37 @@ extern "C" {
 #include "cellarray.h"
 
 cellarray * gridlink(const int64_t np,
-					 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
-					 const DOUBLE xmin, const DOUBLE xmax,
-					 const DOUBLE ymin, const DOUBLE ymax,
-					 const DOUBLE zmin, const DOUBLE zmax,
-					 const DOUBLE max_x_size,
-					 const DOUBLE max_y_size,
-					 const DOUBLE max_z_size,
-					 const int xbin_refine_factor,
-					 const int ybin_refine_factor,
-					 const int zbin_refine_factor,
-					 int *nlattice_x,
-					 int *nlattice_y,
-					 int *nlattice_z) __attribute__((warn_unused_result)); 
+										 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
+										 const DOUBLE xmin, const DOUBLE xmax,
+										 const DOUBLE ymin, const DOUBLE ymax,
+										 const DOUBLE zmin, const DOUBLE zmax,
+										 const DOUBLE max_x_size,
+										 const DOUBLE max_y_size,
+										 const DOUBLE max_z_size,
+										 const int xbin_refine_factor,
+										 const int ybin_refine_factor,
+										 const int zbin_refine_factor,
+										 int *nlattice_x,
+										 int *nlattice_y,
+										 int *nlattice_z) __attribute__((warn_unused_result)); 
+										 
 
+cellarray_nvec * gridlink_nvec(const int64_t np,
+															 const DOUBLE *x,const DOUBLE *y,const DOUBLE *z,
+															 const DOUBLE xmin, const DOUBLE xmax,
+															 const DOUBLE ymin, const DOUBLE ymax,
+															 const DOUBLE zmin, const DOUBLE zmax,
+															 const DOUBLE max_x_size,
+															 const DOUBLE max_y_size,
+															 const DOUBLE max_z_size,
+															 const int xbin_refine_factor,
+															 const int ybin_refine_factor,
+															 const int zbin_refine_factor,
+															 int *nlattice_x,
+															 int *nlattice_y,
+															 int *nlattice_z) __attribute__((warn_unused_result)); 
+
+	
[…]

### 2015-02-14 utils/gridlink_mocks.c

@@ -71,7 +71,7 @@ void get_max_min_data(const int64_t ND1, const DOUBLE * restrict cz,
 
 
 cellarray_mocks *gridlink1D(const int64_t np,const DOUBLE czmin,const DOUBLE czmax, const DOUBLE rcell,
-														const DOUBLE *theta, const DOUBLE *phi, const DOUBLE *cz, 
+														const DOUBLE *dec, const DOUBLE *ra, const DOUBLE *cz,
 														int *ngrid,int *max_in_cell,
 														const int zbin_refine_factor)
 {
@@ -85,21 +85,23 @@ cellarray_mocks *gridlink1D(const int64_t np,const DOUBLE czmin,const DOUBLE czm
   const int64_t totncells = nmesh;
 
   int64_t expected_n=(int64_t)((np/(double) nmesh)  *MEMORY_INCREASE_FAC);
-  expected_n=expected_n <= 1 ? 2:expected_n;
+  expected_n=expected_n < NVEC ? NVEC:expected_n;
+	while(expected_n % NVEC != 0) {
+		expected_n++;
+	}
+	
 #ifndef SILENT
   fprintf(stderr,"%s> Allocating %0.2g (MB) memory for the lattice, expected_n = %"PRId64" nmesh = %d np=%"PRId64" \n",__FUNCTION__,(3*4)*expected_n*nmesh/(1024.*1024.),expected_n,nmesh,np);
 #endif
   cellarray_mocks *lattice = my_malloc(sizeof(cellarray_mocks), totncells);
 
   /*
-	Allocate memory for each of the fields in cellarray. Since we haven't processed the data yet, 
-	expected_n is a reasonable guess as to the number of points in the cell. 
+	Allocate memory for each of the fields in cellarray. Since we haven't processed the data yet,
+	expected_n is a reasonable guess as to the number of points in the cell.
   */
   for (int64_t index=0;index<totncells;index++) {
-		lattice[index].x  = my_malloc(sizeof(DOUBLE),expected_n);
-		lattice[index].y  = my_malloc(sizeof(DOUBLE),expected_n);
-		lattice[index].z  = my_malloc(sizeof(DOUBLE),expected_n);
-		lattice[index].cz = my_malloc(sizeof(DOUBLE),expected_n);
[…]

### 2015-02-14 xi_mocks/vpf/countspheres_mocks.c

@@ -46,26 +46,31 @@ int count_neighbors(const DOUBLE xcen,const DOUBLE ycen,const DOUBLE zcen,const 
       const int max_iz = iz + bin_refine_factor > ngrid-1 ? ngrid-1:iz + bin_refine_factor;
       
       for(int iiz=min_iz;iiz<=max_iz;iiz++) {
-		DOUBLE newzpos = zcen;
-		const int64_t index=iix*ngrid*ngrid + iiy*ngrid + iiz;
-		cellstruct = &(lattice[index]);
-		const DOUBLE *x2 = cellstruct->x;
-		const DOUBLE *y2 = cellstruct->y;
-		const DOUBLE *z2 = cellstruct->z;
-#if  __INTEL_COMPILER
-#pragma simd reduction(+:numngb) vectorlengthfor(DOUBLE)
-#endif
-		for(int i=0;i<cellstruct->nelements;i++) {
-		  const DOUBLE dx=x2[i]-newxpos;
-		  const DOUBLE dy=y2[i]-newypos;
-		  const DOUBLE dz=z2[i]-newzpos;
-		  const DOUBLE r2 = dx*dx + dy*dy + dz*dz;
-		  if (r2 < rmax_sqr) numngb++;
-		}
-		
-		if(numngb > nthreshold)
-		  return numngb;
-      }
+				DOUBLE newzpos = zcen;
+				const int64_t index=iix*ngrid*ngrid + iiy*ngrid + iiz;
+				cellstruct = &(lattice[index]);
+				DOUBLE *x2 = cellstruct->pos;
+				DOUBLE *y2 = cellstruct->pos + NVEC;
+				DOUBLE *z2 = cellstruct->pos + 2*NVEC;
+				
+				for(int i=0;i<cellstruct->nelements;i+=NVEC) {
+					int block_size = cellstruct->nelements - i ;
+					if(block_size > NVEC) block_size = NVEC;
+					for(int ii=0;ii<block_size;ii++) {
+						const DOUBLE dx=x2[ii]-newxpos;
+						const DOUBLE dy=y2[ii]-newypos;
+						const DOUBLE dz=z2[ii]-newzpos;
+						const DOUBLE r2 = dx*dx + dy*dy + dz*dz;
+						if (r2 < rmax_sqr) numngb++;
+					}
+					if(numngb > nthreshold) return numngb;
+					
+					x2 += 3*NVEC;
+					y2 += 3*NVEC;
+					z2 += 3*NVEC;
+
[…]

### 2015-02-14 utils/avx_calls.h
Fairly large internal re-organization for the lattice structure. In some cases, it is faster to store the positions in AVX chunks (8 for floats/4 for doubles).

@@ -12,6 +12,7 @@
 #include <stdlib.h>
 #include <immintrin.h>
 
+#include "function_precision.h"
 
 #define PREFETCH(mem)        asm ("prefetcht0 %0"::"m"(mem))
 
@@ -66,7 +67,12 @@
 #define AVX_BLEND_FLOATS_WITH_MASK(FALSEVALUE,TRUEVALUE,MASK) _mm256_blendv_ps(FALSEVALUE,TRUEVALUE,MASK)
 
 //Trig
+#ifdef  __INTEL_COMPILER
 #define AVX_ARC_COSINE(X)                 _mm256_acos_ps(X)
+#else
+//Other compilers do not have the vectorized arc-cosine
+#define AVX_ARC_COSINE(X)                  inv_cosine(X)
+#endif
 
 //Max
 #define AVX_MAX_FLOATS(X,Y)               _mm256_max_ps(X,Y)
@@ -114,7 +120,11 @@
 #define AVX_BLEND_FLOATS_WITH_MASK(FALSEVALUE,TRUEVALUE,MASK) _mm256_blendv_pd(FALSEVALUE,TRUEVALUE,MASK)
 
 //Trig
+#ifdef  __INTEL_COMPILER
 #define AVX_ARC_COSINE(X)                 _mm256_acos_pd(X)
+#else
+#define AVX_ARC_COSINE(X)                  inv_cosine(X)
+#endif
 
 //Max
 #define AVX_MAX_FLOATS(X,Y)               _mm256_max_pd(X,Y)
@@ -127,6 +137,30 @@
 #define AVX_STREAMING_STORE_FLOATS(X,Y)   _mm256_stream_pd(X,Y)
 #define AVX_STREAMING_STORE_INTS(X,Y)     _mm_stream_si128(X,Y)
 
-
-
 #endif
+
+
+#ifndef  __INTEL_COMPILER
+static inline AVX_FLOATS inv_cosine(const AVX_FLOATS X) 
+{																																				
+	union cos{
+		AVX_FLOATS m;
+		DOUBLE x[NVEC];
+	};
+	union cos union_costheta;
+	union cos union_returnvalue;
+	union_costheta.m = X;
+	
+	for(int ii=0;ii<NVEC;ii++) {														
+		const DOUBLE costheta = union_costheta.x[ii];
+		if(costheta <= 1.0 && costheta >= -1.0) {
+			union_returnvalue.x[ii] = ACOS(costheta);
+		} else {
[…]

### 2015-02-14 utils/cellarray_mocks.h

@@ -21,18 +21,15 @@ extern "C" {
 #define NLATMAX   100      /* maximum grid dimension in X-Y plane */
 
 typedef struct{
-  DOUBLE *x;
-  DOUBLE *y;
-  DOUBLE *z;
+  DOUBLE *pos;
+  /* DOUBLE *y; */
+  /* DOUBLE *z; */
   int nelements;
 	int nallocated;
 } cellarray;
 
 typedef struct{
-  DOUBLE *x;
-  DOUBLE *y;
-  DOUBLE *z;
-  DOUBLE *cz;
+  DOUBLE *pos;
   int nelements;
   int nallocated;
 } cellarray_mocks;

### 2015-02-14 xi_mocks/tests/tests_mocks.c

@@ -33,7 +33,7 @@
 #endif
 
 #if !(defined(__INTEL_COMPILER)) && defined(USE_AVX)
-#error Test suite for mocks will not work without Intel ICC while USE_AVX is set. Disable USE_AVX to use the test suite with non-Intel compilers. 
+#warning Test suite for mocks will be slow without Intel ICC while USE_AVX is set. 
 #endif
 
 #ifndef SILENT
@@ -287,17 +287,17 @@ int main(int argc, char **argv)
   const char firstfilename[][MAXLEN] = {"../tests/data/Mr19_mock_northonly.rdcz.dat",
 																				"../tests/data/Mr19_mock_northonly.rdcz.dat",
 																				"../tests/data/Mr19_mock_northonly.rdcz.dat",
-																				"../tests/data/Mr19_randoms_northonly.rdcz.dat",
-																				"../tests/data/Mr19_randoms_northonly.rdcz.dat",
-																				"../tests/data/Mr19_randoms_northonly.rdcz.dat"};
-  const char firstfiletype[][MAXLEN]  = {"a","a","a","a","a","a"};
+																				"../tests/data/Mr19_randoms_northonly.rdcz.ff",
+																				"../tests/data/Mr19_randoms_northonly.rdcz.ff",
+																				"../tests/data/Mr19_randoms_northonly.rdcz.ff"};
+  const char firstfiletype[][MAXLEN]  = {"a","a","a","f","f","f"};
   const char secondfilename[][MAXLEN] = {"../tests/data/Mr19_mock_northonly.rdcz.dat",
 																				 "../tests/data/Mr19_mock_northonly.rdcz.dat",
 																				 "../tests/data/Mr19_mock_northonly.rdcz.dat",
 																				 "../tests/data/Mr19_mock_northonly.rdcz.dat",
 																				 "../tests/data/Mr19_mock_northonly.rdcz.dat",
-																				 "../tests/data/Mr19_randoms_northonly.rdcz.dat"};
-  const char secondfiletype[][MAXLEN] = {"a","a","a","a","a","a"};
+																				 "../tests/data/Mr19_randoms_northonly.rdcz.ff"};
+  const char secondfiletype[][MAXLEN] = {"a","a","a","a","a","f"};
   const DOUBLE allpimax[]             = {40.0,40.0,40.0,40.0,40.0,40.0};
 
   int (*allfunctions[]) (const char *) = {test_DDrppi_mocks,test_wtheta_mocks,test_vpf_mocks};

### 2015-02-14 xi_theory/xi_rp_pi/countpairs_rp_pi.c

@@ -44,13 +44,13 @@ void free_results_rp_pi(results_countpairs_rp_pi **results)
 
 
 results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1, const DOUBLE *Y1, const DOUBLE *Z1,
-											const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
+																						const int64_t ND2, const DOUBLE *X2, const DOUBLE *Y2, const DOUBLE *Z2,
 #ifdef USE_OMP
-											const int numthreads,
+																						const int numthreads,
 #endif
-											const int autocorr,
-											const char *binfile,
-											const double pimax)
+																						const int autocorr,
+																						const char *binfile,
+																						const double pimax)
 {
   int bin_refine_factor=1;
 	int zbin_refine_factor=2;
@@ -61,17 +61,17 @@ results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1,
 		bin_refine_factor=1;
 		zbin_refine_factor=1;
 	}
-#ifdef USE_OMP
-	if(numthreads > 1) {
-		if(autocorr==1) {
-			bin_refine_factor=2;
-			zbin_refine_factor=2;
-		} else {
-			bin_refine_factor=1;
-			zbin_refine_factor=1;
-		}
-	}
-#endif
+/* #ifdef USE_OMP */
+/* 	if(numthreads > 1) { */
+/* 		if(autocorr==1) { */
+/* 			bin_refine_factor=2; */
+/* 			zbin_refine_factor=2; */
+/* 		} else { */
+/* 			bin_refine_factor=1; */
+/* 			zbin_refine_factor=1; */
+/* 		} */
+/* 	} */
+/* #endif */
 	const int npibin = (int) pimax;
   /***********************
    *initializing the  bins
@@ -185,7 +185,7 @@ results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1,
 		for(int i=0;i<totnbins;i++) rpavg[i] = 0.0;
 #endif			
 
-#pragma omp for  schedule(dynamic)
+#pragma omp for  schedule(dynamic) nowait
 #endif
[…]

### 2015-02-20 xi_mocks/DDrppi/countpairs_rp_pi_mocks.c
Changed the rbin_refine_factor in wtheta and added some alignment keywords

@@ -11,7 +11,8 @@
 #include <stdlib.h>
 #include <gsl/gsl_interp.h>
 
-#include "sglib.h"
+#include "defs.h"
+/* #include "sglib.h" */
 #include "utils.h"
 #include "cellarray_mocks.h"
 #include "gridlink_mocks.h"
@@ -324,10 +325,10 @@ results_countpairs_mocks * countpairs_mocks(const int64_t ND1, DOUBLE *phi1, DOU
 #pragma omp parallel shared(numdone)
 	{
 		const int tid = omp_get_thread_num();
-		uint64_t npairs[totnbins] __attribute__ ((aligned(64)));
+		uint64_t npairs[totnbins] __attribute__ ((aligned(ALIGNMENT)));
 		for(int i=0;i<totnbins;i++) npairs[i] = 0;
 #ifdef OUTPUT_RPAVG
-		DOUBLE rpavg[totnbins];
+		DOUBLE rpavg[totnbins] __attribute__ ((aligned(ALIGNMENT)));
 		for(int i=0;i<totnbins;i++) rpavg[i] = 0.0;
 #endif
 

### 2015-02-25 utils/function_precision.h
Added some protection for C++ compilers. Switched the objects files to the front for the Makefiles

@@ -8,6 +8,12 @@
 
 #pragma once
 
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
 #ifndef M_PI
 #define M_PI            3.14159265358979323846264338327950288 
 #endif
@@ -72,4 +78,8 @@
 #define ASIN   asinf
 #define POW    powf
 #define ABS    fabsf
+#endif
+
+#ifdef __cplusplus
+}
 #endif

### 2015-02-25 paper/ms.tex
Epic re-write of the io.c file. Can read in arbitrary numbers of fields of DOUBLE precision
Added some protection for C++ compilers. Switched the objects files to the front for the Makefiles

@@
 \usepackage{url}
+\usepackage{adjustbox}
 \journal{Astronomy and Computing}
@@
 \newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
-\newcommand{\xir}{\ensuremath{{\xi(r)}}\xspace}
+\newcommand{\xir}{\ensuremath{{DD(r)}}\xspace}
+\newcommand{\xiofr}{\ensuremath{{\xi(r)}}\xspace}
 \newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
@@
 \subsection{Partitioning the Particles based on \rmax}
-
+Fig.~\ref{fig:grid} shows a schematic 2-D grid that corresponds to our partitioning scheme.
 \begin{figure}[htbp]
@@
 A circle with radius \rmax is also drawn to shown the actual pairs that will eventually count in the correlation function.} 
+\label{fig:grid}
 \end{figure}
@@
 
+\begin{table}
+\centering
+\caption{\footnotesize OpenMP scaling for the three codes. Efficiencies are defined 
+as \texttt{Speedup/Nthreads}. Note, that \xir has super-linear scaling with \texttt{nthreads}
+up to $\sim$ 10 threads.}
+\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{ccccc} 
+\toprule
+\multirow{2}{*}{\textbf{Nthreads}}   &
+\multicolumn{4}{c}{\textbf{Efficiency[\%]}} \\
+\cmidrule(l{1.0em}r{1.0em}){2-5}
+&
+\multicolumn{1}{c}{$\boldsymbol{\xir}$}    &
+\multicolumn{1}{c}{$\boldsymbol{\xirppi}$} &
+\multicolumn{1}{c}{$\boldsymbol{\wprp}$}   & 
+\multicolumn{1}{c}{$\boldsymbol{\xiofr}$}   \\
+\midrule
+\input{timings_Mr19_openmp} 
+\bottomrule
+\end{tabular}
+\end{adjustbox}
+\label{table:openmp}
+\end{table}
+
 \section{Conclusions}

### 2015-02-25 utils/avx_calls.h

@@ -12,6 +12,11 @@
 #include <stdlib.h>
 #include <immintrin.h>
 
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+
 #include "function_precision.h"
 
 #define PREFETCH(mem)        asm ("prefetcht0 %0"::"m"(mem))
@@ -168,3 +173,6 @@ static inline AVX_FLOATS inv_cosine(const AVX_FLOATS X)
 }
 #endif
 
+#ifdef __cplusplus
+}
+#endif

### 2015-02-25 paper/ms.tex
Added a joint OpenMP scaling table for both mocks and theory (by copy-pasting the two individual OpenMP scaling files)
Fixed the Makefile. Updated the paper with the mocks plots

@@
 \includegraphics[clip=true,width=\linewidth]{timings_Mr19_numpart}%
-\caption{Scaling with particle number for \xir, \xirppi, \wprp }
+\caption{Scaling with particle number for \xir, \xirppi, \wprp and \xiofr}
 \label{fig:scaling_numpart}
@@
 \begin{figure}[htbp]
-\includegraphics[clip=true,width=\linewidth]{timings_Mr19_rmax}%
-\caption{Scaling with \rmax for \xir, \xirppi, \wprp }
+\includegraphics[clip=true,width=0.5\linewidth]{timings_Mr19_rmax}%
+\includegraphics[clip=true,width=0.5\linewidth]{timings_Mr19_mocks_rmax}
+\caption{Scaling with \rmax for \xir, \xirppi, \wprp and \xiofr}
 \label{fig:scaling_rmax}
@@
 \begin{figure}[htbp]
-\includegraphics[clip=true,width=\linewidth]{timings_Mr19_openmp}%
-\caption{OpenMP scaling for \xir, \xirppi, \wprp }
+\includegraphics[clip=true,width=0.5\linewidth]{timings_Mr19_openmp}%
+\includegraphics[clip=true,width=0.5\linewidth]{timings_Mr19_mocks_openmp}
+\caption{OpenMP scaling for \xir, \xirppi, \wprp and \xiofr}
 \label{fig:scaling_openmp}
@@
 \begin{adjustbox}{max width=\textwidth}
-\begin{tabular}{ccccc} 
+\begin{tabular}{ccccccccc} 
 \toprule
-\multirow{2}{*}{\textbf{Nthreads}}   &
-\multicolumn{4}{c}{\textbf{Efficiency[\%]}} \\
+\multirow{3}{*}{\textbf{Nthreads}}   &
+\multicolumn{8}{c}{\textbf{Efficiency[\%]}} \\
+\cmidrule(l{3em}r{3em}){2-9}
+
+                                     &
+\multicolumn{4}{c}{\textbf{Periodic Boxes}}  &
+\multicolumn{4}{c}{\textbf{Spherical Geometry}}  \\
 \cmidrule(l{1.0em}r{1.0em}){2-5}
+\cmidrule(l{1.0em}r{1.0em}){6-9}
                                                &
@@
 \multicolumn{1}{c}{$\boldsymbol{\wprp}$}       & 
-\multicolumn{1}{c}{$\boldsymbol{\xiofr}$}   \\
+\multicolumn{1}{c}{$\boldsymbol{\xiofr}$}      &
+\multicolumn{1}{c}{$\boldsymbol{DD(r_p,\pi)}$} &
+\multicolumn{1}{c}{$\boldsymbol{DD(\theta)}$}  &
+\multicolumn{1}{c}{$\boldsymbol{DR(r_p,\pi)}$} & 
+\multicolumn{1}{c}{$\boldsymbol{DR(\theta)}$}   \\
 \midrule
-\input{timings_Mr19_openmp} 
+\input{timings_Mr19_openmp_theory_and_mocks}
 \bottomrule

### 2015-03-25 paper/ms.tex
Fixed Makefile and silenced clang static analyzer

@@
 
-Modern cpu's have a hierarchy of memory locations; the smallest and fastest are closest to the cores while the largest and slowest are the
+Modern cpu's have a hierarchy of memory locations; the smallest (and fastest) are closest to the cores while the largest (and slowest) are the
 farthest. All cpu instructions need to be carried out from cpu registers - there are $\sim 10-30$ registers typically available and the 

### 2015-03-25 xi_theory/xi_rp_pi/countpairs_rp_pi.c

@@ -529,6 +529,7 @@ results_countpairs_rp_pi * countpairs_rp_pi(const int64_t ND1, const DOUBLE *X1,
 		results->rupp[i] = rupp[i];
 		for(int j=0;j<npibin;j++) {
 			int index = i*(npibin+1) + j;
+			assert(index < totnbins && "index must be within range");
 			results->npairs[index] = npairs[index];
 #ifdef OUTPUT_RPAVG
 			results->rpavg[index] = rpavg[index];

### 2015-03-27 paper/ms.tex
Updated the logic in DDrppi to remove using the comoving distance array. Now, I check for max. separation before proceeding with the division

--- 2015-02-25 paper/ms.tex
@@
 \newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
+\newcommand{\cis}{\ensuremath{{pN(r)}}\xspace}
+\newcommand{\ddrppi}{\ensuremath{{DD(r_p,\pi)}}\xspace}
+%% \newcommand{\ddrppi}{\ensuremath{{DD(r_p,\pi)}}\xspace}
+\newcommand{\wtheta}{\ensuremath{{DD(\theta)}}\xspace}
 \newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
@@
 \begin{abstract}
-We have entered the era of `Big Data' where we are gathering a tremendous amount of data that
-needs to be processed as well as modeled accurately. In particular, large galaxy surveys
-like the Sloan Digitial Sky Survey requires computing correlation functions. While measuring the
-correlation function in the data is not a bottle-neck, modeling the observed galaxy distribution correctly
-{\em} requires a MCMC chain and repeated measurements of $\xi(r)$ and/or $\xi(r_p,\pi)$. However, 
-the architecture of modern CPU's means that best performance is only obtained when cache misses 
-are kept to a minimum. Here, I present a suite of 3 correlation function codes that exploit 
-current CPU architecture and hand-written AVX
-
+In the era of `Big Data', we need special computing techniques to both process the raw data as well 
+generate and evaluate models to test our theories. In particular, large galaxy surveys
+like the Sloan Digitial Sky Survey requires computing a variety of $n$-point statistics, e.g., the 
+2-pt projected correlation function, $\wprp$. While measuring the correlation function in the data is not a bottle-neck, 
+modeling the observed galaxy distribution correctly
+{\em requires} a MCMC chain and repeated measurements of $\wprp$. To efficiently 
+measure the correlation functions, we need to write code that accounts for cache hierarchies while 
+simultaneously exploits the wide vector registers present in the relatively modern CPUs.  
+Here, I present a suite of OpenMP parallelized, clustering codes that exploit current CPU micro-architecture with hand-written AVX 
+intrinsics. Given a 3-D distribution of points in a cartesian system, these codes compute spatial clustering statistics 3-D correlation functions \xir, \xiofr, \xirppi; 2-point 
+projected correlation function \wprp; counts-in-spheres \cis. For a 3-D particle distribution on the sky, the codes can compute 
+projected correlation function \ddrppi, angular correlation function \wtheta as well as \cis. These codes are designed to be blazing fast 
+and can compute \wprp for $\mathcal{O}$(1 million) galaxies in $\sim 6$ seconds on a post-2011 CPU, which is a factor of few faster than 
+the existing public correlation function routines.  
 The accompanying codes are publicly available at \url{https://bitbucket.org/manodeep/Corrfunc/}. 
@@
 
-Modern cpu's have a hierarchy of memory locations; the smallest and fastest are closest to the cores while the largest and slowest are the
-farthest. All cpu instructions need to be carried out from cpu registers - there are $\sim 10-30$ registers typically available and the 
+Modern cpu's have a hierarchy of memory locations; the smallest (and fastest) are physically located close to the computing cores 
+while the largest (and slowest) are the farthest. All cpu instructions need to be carried out from cpu registers - there are $\sim 100$ 
+registers typically available and the 
 access times can be thought of as instantaneous. Next up is the $L1$ cache divided into $L1D$ for data and $L1I$ for instructions cache. Since 
 the cpu always necessarily executes instructions that are close together, we will ignore the instruction cache from now on. Typical $L1$ cache 
-sizes range from 64KB to 128 KB. Next level up is the $L2$ cache, typically $\sim$ 256KB to 1 MB. The last level cache or the $L3$ cache 
-is usually shared across all cores on the socket and can be 10 MB to 40 MB.
+sizes range from 64KB to 128 KB (shared between instruction and data). 
+Next level up is the $L2$ cache, typically $\sim$ 256KB to 1 MB. The last level cache or the $L3$ cache is usually shared across all 
+cores on the socket and can be 10 MB to 40 MB.
 
@@
 \subsection{\wprp}
+\subsection{}
 

### 2015-03-27 xi_mocks/DDrppi/countpairs_rp_pi_mocks.c

@@ -158,7 +158,8 @@ results_countpairs_mocks * countpairs_mocks(const int64_t ND1, DOUBLE *phi1, DOU
   setup_bins(binfile,&rpmin,&rpmax,&nrpbin,&rupp);
   assert(rpmin > 0.0 && rpmax > 0.0 && rpmin < rpmax && "[rpmin, rpmax] are valid inputs");
   assert(nrpbin > 0 && "Number of rp bins is valid");
-
+  
+  const DOUBLE sqr_max_sep = rpmax*rpmax + pimax*pimax;
       
   //Change cz into co-moving distance
   DOUBLE *d1     = my_malloc(sizeof(*d1),ND1);
@@ -406,17 +407,16 @@ results_countpairs_mocks * countpairs_mocks(const int64_t ND1, DOUBLE *phi1, DOU
 					DOUBLE *x2  = cellstruct->pos;
 					DOUBLE *y2  = cellstruct->pos + NVEC;
 					DOUBLE *z2  = cellstruct->pos + 2*NVEC;
-					DOUBLE *cz2 = cellstruct->pos + 3*NVEC;
-					
-					const DOUBLE TWO=2.0;
-					const DOUBLE sqr_d1 = d1[i]*d1[i];
+/* 					DOUBLE *cz2 = cellstruct->pos + 3*NVEC; */
+/* 					const DOUBLE TWO=2.0; */
+/* 					const DOUBLE sqr_d1 = d1[i]*d1[i]; */
 		  
 #ifndef USE_AVX
 
 					DOUBLE *localx2  = x2;
 					DOUBLE *localy2  = y2;
 					DOUBLE *localz2  = z2;
-					DOUBLE *localcz2 = cz2;
+/* 					DOUBLE *localcz2 = cz2; */
 
 					for(int j=0;j<cellstruct->nelements;j+=NVEC) {
 						int block_size2=cellstruct->nelements - j;
@@ -449,17 +449,17 @@ results_countpairs_mocks * countpairs_mocks(const int64_t ND1, DOUBLE *phi1, DOU
 						}//end of jj-loop
 
 						//increment localx2/localy2 etc
-						localx2   += 4*NVEC;//this might actually exceed the allocated range but we will never dereference that
-						localy2   += 4*NVEC;
[…]

### 2016-02-02 paper/users-guide.tex
Added some protection for C++ compilers. Switched the objects files to the front for the Makefiles
Updated the logic in DDrppi to remove using the comoving distance array. Now, I check for max. separation before proceeding with the division

@@
 \subsection{Getting the Source}
-You can obtain the source in two ways: i) Clone the mercurial repo (\texttt{hg clone} \url{https://bitbucket.org/manodeep/corrfunc/}) 
-or ii) Download the tar archive (\href{https://bitbucket.org/manodeep/corrfunc/downloads/corrfunc.1.0.0.tar.gz}{corrfunc.\$MAJOR.0.\$MINOR.tar.gz}) and 
-unpack it in the directory where you wish to keep the files (\texttt{tar xvzf}\\ \texttt{corrfunc.\$MAJOR.0.\$MINOR.tar.gz}). Here, \texttt{\$MAJOR}
-and \texttt{\$MINOR} refer to the major and minor release versions (current \texttt{\$MAJOR=1}, \texttt{\$MINOR=0}). I will only change 
+You can obtain the source in two ways: i) Clone the git repo (\texttt{git clone} \url{https://github/manodeep/Corrfunc/}) 
+or ii) Download the tar archive (\href{https://github.com/manodeep/Corrfunc/releases/latest}{Corrfunc.\$MAJOR.\$MINOR.\$PATCH.tar.gz}) and 
+unpack it in the directory where you wish to keep the files (\texttt{tar
+  xvzf}\\ \texttt{Corrfunc.\$MAJOR.\$MINOR.\$PATCH.tar.gz}). Here,
+\texttt{\$MAJOR}, \texttt{\$MINOR} and \texttt{\$PATCH} refer to the major,
+minor and patch release versions (current \texttt{\$MAJOR=0},
+\texttt{\$MINOR=2}, \texttt{\$PATCH=0}). I will only change 
 the \texttt{\$MAJOR} version in the highly unlikely event that the API changes.
@@
   \end{minipage}.
-
 .2 xi\_mocks.
@@
 as \texttt{Speedup/Nthreads}. Note, that \xir has super-linear scaling with \texttt{nthreads}
-up to $\sim$ 8 threads.}
+up to $\sim$ 10 threads.}
 \begin{adjustbox}{max width=\textwidth}
-\begin{tabular}{cccc} 
+\begin{tabular}{ccccccccc} 
 \toprule
-\multirow{2}{*}{\textbf{Nthreads}}   &
-\multicolumn{3}{c}{\textbf{Efficiency[\%]}} \\
-\cmidrule(l{1.0em}r{1.0em}){2-4}
+\multirow{3}{*}{\textbf{Nthreads}}   &
+\multicolumn{8}{c}{\textbf{Efficiency[\%]}} \\
+\cmidrule(l{3em}r{3em}){2-9}
+
+                                     &
+\multicolumn{4}{c}{\textbf{Periodic Boxes}}  &
+\multicolumn{4}{c}{\textbf{Spherical Geometry}}  \\
+\cmidrule(l{1.0em}r{1.0em}){2-5}
+\cmidrule(l{1.0em}r{1.0em}){6-9}
                                                &
@@
 \multicolumn{1}{c}{$\boldsymbol{\xirppi}$}     &
-\multicolumn{1}{c}{$\boldsymbol{\wprp}$}     \\
+\multicolumn{1}{c}{$\boldsymbol{\wprp}$}       & 
+\multicolumn{1}{c}{$\boldsymbol{\xiofr}$}      &
+\multicolumn{1}{c}{$\boldsymbol{DD(r_p,\pi)}$} &
+\multicolumn{1}{c}{$\boldsymbol{DD(\theta)}$}  &
+\multicolumn{1}{c}{$\boldsymbol{DR(r_p,\pi)}$} & 
+\multicolumn{1}{c}{$\boldsymbol{DR(\theta)}$}   \\
 \midrule
-\input{timings_Mr19_openmp} 
+\input{timings_Mr19_openmp_theory_and_mocks}
 \bottomrule

### 2016-06-08 paper/ms.tex
Updating paper. Pushing the changes

@@
 
-\title{Cache is King: Presenting a Suite of Fast, Correlation Function Codes}
+\title{Cache is King: Presenting a Suite of Fast Correlation Function Codes}
 
@@
 like the Sloan Digitial Sky Survey requires computing a variety of $n$-point statistics, e.g., the 
-2-pt projected correlation function, $\wprp$. While measuring the correlation function in the data is not a bottle-neck, 
-modeling the observed galaxy distribution correctly
-{\em requires} a MCMC chain and repeated measurements of $\wprp$. To efficiently 
+2-pt projected correlation function, $\wprp$. 
+Since the clustering measure on the data is only done once, time taken to do
+this measurement is not a bottleneck. However, modeling the observed galaxy distribution correctly
+{\em requires} a MCMC chain and repeated measurements of some clustering statistic. To efficiently 
 measure the correlation functions, we need to write code that accounts for cache hierarchies while 
@@
 Here, I present a suite of OpenMP parallelized, clustering codes that exploit current CPU micro-architecture with hand-written AVX 
-intrinsics. Given a 3-D distribution of points in a cartesian system, these codes compute spatial clustering statistics 3-D correlation functions \xir, \xiofr, \xirppi; 2-point 
-projected correlation function \wprp; counts-in-spheres \cis. For a 3-D particle distribution on the sky, the codes can compute 
+intrinsics. Given a 3-D distribution of points in a cartesian system (i.e.,
+generated from cosmological simulations), these codes compute spatial clustering statistics 3-D correlation functions \xir, \xiofr, \xirppi; 2-point 
+projected correlation function \wprp; counts-in-spheres \cis. For a 3-D
+particle distribution on the sky (i.e., mock galaxy catalogs, real observed galaxies), the codes can compute 
 projected correlation function \ddrppi, angular correlation function \wtheta as well as \cis. These codes are designed to be blazing fast 
 and can compute \wprp for $\mathcal{O}$(1 million) galaxies in $\sim 6$ seconds on a post-2011 CPU, which is a factor of few faster than 
-the existing public correlation function routines.  
-The accompanying codes are publicly available at \url{https://bitbucket.org/manodeep/Corrfunc/}. 
+the existing public correlation function routines.  The accompanying codes are
+publicly available at \url{https://github.com/manodeep/Corrfunc/} while
+benchmarks with existing codes are available at \url{https://gist.github.com/manodeep/cffd9a5d77510e43ccf0}.
+
 \end{abstract}

### 2016-06-22 paper/ms.tex
latest paper changes

--- 2015-03-25 paper/ms.tex
@@
-\documentclass[preprint,authoryear,12pt]{elsarticle}
+\documentclass[preprint, 12pt, authoryear]{elsarticle}
 
@@
 \newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
+\newcommand{\cis}{\ensuremath{{pN(r)}}\xspace}
+\newcommand{\ddrppi}{\ensuremath{{DD(r_p,\pi)}}\xspace}
+\newcommand{\wtheta}{\ensuremath{{DD(\theta)}}\xspace}
 \newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
@@
 
-\title{Cache is King: Presenting a Suite of Fast, Correlation Function Codes}
+\title{Cache is King: Presenting {\tt Corrfunc} -- a Suite of Fast Correlation Function Codes}
 
@@
 \begin{abstract}
-We have entered the era of `Big Data' where we are gathering a tremendous amount of data that
-needs to be processed as well as modeled accurately. In particular, large galaxy surveys
-like the Sloan Digitial Sky Survey requires computing correlation functions. While measuring the
-correlation function in the data is not a bottle-neck, modeling the observed galaxy distribution correctly
-{\em} requires a MCMC chain and repeated measurements of $\xi(r)$ and/or $\xi(r_p,\pi)$. However, 
-the architecture of modern CPU's means that best performance is only obtained when cache misses 
-are kept to a minimum. Here, I present a suite of 3 correlation function codes that exploit 
-current CPU architecture and hand-written AVX
-
-The accompanying codes are publicly available at \url{https://bitbucket.org/manodeep/Corrfunc/}. 
+In the era of `Big Data', we need special computing techniques to both process the raw data as well 
+generate and evaluate models to test our theories. In particular, large galaxy surveys
+like the Sloan Digitial Sky Survey requires computing a variety of $n$-point statistics, e.g., the 
+2-pt projected correlation function, $\wprp$. 
+Since the clustering measure on the data is only done once, time taken to do
+such a measurement is not a bottleneck. However, modeling the observed galaxy distribution correctly
+{\em requires} a Monte-Carlo Markov Chain (MCMC) and repeated measurements of one or multiple
+clustering statistics. In order to efficiently compute {\em any} clustering statistic,
+we need to write code that is tuned for the underlying hardware. For
+modern CPUs, such tuning involves proper utilization of the cache hierarchy {\em and}
+vectorization with Single Instruction Multiple Data (SIMD) capable wide vector registers. 
+In this paper, I present {\tt Corrfunc} --  a suite of OpenMP parallelized, clustering
+codes that exploit current CPU micro-architecture with hand-written Advanced
+Vector Extensions (AVX) and  Streaming SIMD Extensions (SSE) intrinsics. 
+Given a 3-D distribution of points in a cartesian system (i.e.,
+generated from cosmological simulations), {\tt Corrfunc} can compute spatial
+3-D correlation functions \xir, \xiofr, \xirppi; 2-point 
+projected correlation function \wprp; counts-in-spheres \cis. For a 3-D
+particle distribution on the sky (i.e., mock galaxy catalogs, observed galaxies), the codes can compute 
+projected correlation function \ddrppi, angular correlation function \wtheta as
+well as \cis.  By design, {\tt Corrfunc} is highly optimized 
+and can compute \wprp for $\mathcal{O}$(1 million) galaxies in
+$\sim 6$ seconds on a post-2011 CPU, which is a factor of few faster than 
+existing public correlation function routines.  {\tt Corrfunc} is
+publicly available at \url{https://github.com/manodeep/Corrfunc/} while
+benchmarks with existing codes are available at
+\url{https://gist.github.com/manodeep/cffd9a5d77510e43ccf0}.
 \end{abstract}
@@
 \section{INTRODUCTION}
-%% %% MS: I have not encountered this in the past -- but the footnotecounter needs to be reset. Otherwise, 
-%% %% the numbering starts at Nauthors+1. Presumably \altaffiltext uses \footnote internally and does not
-%% %% reset the counter like a responsible person!
-%% \setcounter{footnote}{0}
 The large-scale structure of the Universe can be now measured using $\mathcal{O}$(million) galaxies 
-using data from current surveys like SDSS/BOSS surveys. Upcoming surveys like LSST will probe even 
+using data from current surveys like the Sloan Digitial Sky Survey (SDSS). Upcoming
+surveys on the Large Synoptic Survey Telescope will probe even 
 deeper and wider and aim to target on $\mathcal{O}$(10's of million) galaxies. With such a large galaxy data, 
@@
 
-Modern cpu's have a hierarchy of memory locations; the smallest (and fastest) are closest to the cores while the largest (and slowest) are the
-farthest. All cpu instructions need to be carried out from cpu registers - there are $\sim 10-30$ registers typically available and the 
-access times can be thought of as instantaneous. Next up is the $L1$ cache divided into $L1D$ for data and $L1I$ for instructions cache. Since 
-the cpu always necessarily executes instructions that are close together, we will ignore the instruction cache from now on. Typical $L1$ cache 
-sizes range from 64KB to 128 KB. Next level up is the $L2$ cache, typically $\sim$ 256KB to 1 MB. The last level cache or the $L3$ cache 
-is usually shared across all cores on the socket and can be 10 MB to 40 MB.
+\section{BACKGROUND}
+\subsection{HARDWARE ARCHITECTURE}
+Modern cpu's have a hierarchy of memory locations; the smallest (and fastest) are physically located close to the computing cores 
+while the largest (and slowest) are the farthest. All cpu instructions need to be carried out from cpu registers - there are $\sim 100$ 
+registers typically available and the access times are essentially
+instantaneous. Next up is the {\em L1} cache 
+divided into {\em L1D} for data and {\em L1I} for instructions cache. Since 
+the cpu always necessarily executes instructions that are close together, we will ignore the instruction cache from now on. Typical L1 cache 
+sizes range from 64KB to 128 KB (shared between instruction and data). 
+Next level up is the L2 cache, typically $\sim$ 256KB to 1 MB. The last level cache or the L3 cache is usually shared across all 
+cores on the socket and can be 10 MB to 40 MB. 
+\subsection{VECTORIZATION}
+
+\subsection{OPTIMIZATION STRATEGIES}
+
+\section{PACKAGE SUMMARY}
+{\tt Corrfunc} is written primarily in {\tt C} and comes with convenient
+{\tt python} wrappers for every clustering statistic. The fundamental design
+principles for {\tt Corrfunc} are the following:
+\begin{enumerate}
+  \item Correct -- {\tt Corrfunc} has a base set of correct outputs for every statistic
+    generated either through slow, brute-force methods or independent, 
+    external codes. Within {\tt Corrfunc}, every clustering statistic is covered with at least one
+    test case that {\em requires} reproducing this `known-correct' result {\em
+      exactly}. The `PASS/FAIL' status for each test is the direct output of {\tt diff -q
+      test\_output known\_correct\_output}.  
+  \item High Performance -- Performance is the overarching goal for {\tt
+    Corrfunc}. The code-base has been repeatedly analyzed with Intel VTune
+    (\url{https://software.intel.com/en-us/intel-vtune-amplifier-xe})
+    on a variety of platforms to optimize the innermost loops. 
+  \item Portable -- {\tt Corrfunc} is written in {\tt ISO/IEC 9899:1999}
+    compliant {\tt C}. All instruction set specific codes are protected via compile-time constant
+    definitions.
+\end{enumerate}
+Every clustering statistic in {\tt Corrfunc} can either be accessed through the
+API call (e.g., via {\tt python}), linking with the static libraries) or as
+a command-line executable.
 

### 2016-07-20 paper/ms.tex
Added the speedup plots into the text [ci skip]

@@
 \newcommand{\emcee}{\texttt emcee}
-\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
+\newcommand{\rmax}{\ensuremath{{{\mathcal{R}}_{max}}}\xspace}
+\newcommand{\rmaxcubed}{\ensuremath{{\mathcal{R}^{3}_{max}}}\xspace}
 \newcommand{\xir}{\ensuremath{{DD(r)}}\xspace}
@@
 \newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
+\newcommand{\avx}{\texttt AVX}
+\newcommand{\sse}{\texttt SSE}
 
@@
 codes that exploit current CPU micro-architecture with hand-written Advanced
-Vector Extensions (AVX) and  Streaming SIMD Extensions (SSE) intrinsics. 
+Vector Extensions (\avx) and  Streaming SIMD Extensions (\sse) intrinsics. 
 Given a 3-D distribution of points in a cartesian system (i.e.,
@@
 {\tt Corrfunc} is written primarily in {\tt C} and comes with convenient
-{\tt python} wrappers for every clustering statistic. The fundamental design
+{\tt python} wrappers for typical clustering statistics. The fundamental design
 principles for {\tt Corrfunc} are the following:
@@
 \section{Methods}
-We need to compute pairwise distances to get the correlation function. A naive implementation of a correlation function would compute {\em all possible} 
+\subsection{Complexity of the Code}
+We need to compute pairwise~\citep{} distances to get the correlation function. A naive implementation of a correlation function would compute {\em all possible} 
 pairwise separations with a complexity $\mathcal{O}(N^2)$. However, for almost all correlation functions, we are only interested in separations less than 
@@
 dimension can not be within \rmax of each other. Thus, given one point which is the target galaxy and a grid with cell-size \rmax, 
-immediately allows us to prune {\em all} of the points that are not within 1 cell offset in each dimension. However, even with this pruning, the actual 
+immediately allows us to prune {\em all} of the points that are not within 1 cell offset along each dimension. However, even with this pruning, the actual 
 implementation of the algorithm matters. For instance, a linked-list in each cell performs $\sim 60\times$ worse than the algorithm described here.\footnote{The 
-usual linked-list is very cache-unfriendly. Each dereference requires a read from a new region of memory and an almost guaranteed cache miss. }
+usual linked-list is very cache-unfriendly. Each dereference requires a read
+from a new region of memory and an almost guaranteed cache miss. } 
+We can now compute the theoretical complexity of the proposed algorithm. Let
+$\mathcal{L}$ be the side-length of the cube over which $\mathcal{N}$ points are distributed, then
+each cell contains $r_{\mathrm max}^3\times \mathcal{N/V}$ particles. To compute the
+correlation function, we have to loop over each point, and then {\em all} of
+the points in the neighboring cells. This results in a complexity of 
+$\mathcal{O(N{\dfrac{\rmax}/L}^3}$. A typical \rmax is $\sim 10-20\% \times
+\mathcal{L}$, therefore ordering the particles in cells of size \rmax should
+result in a speedup of $(\rmax/\mathcal{L})^3 \sim 125-1000$ compared to the
+brute-force algorithm.
 
@@
 
+\subsection{Speedup from SIMD code}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[clip=true,scale=0.7]{wp_speedup_AVX}%
+\caption{Speedup from AVX}
+\label{fig:speedup_avx}
+\end{figure}
+\begin{figure}[htbp]
+\centering
+\includegraphics[clip=true,scale=0.7]{wp_speedup_SSE_4_2}
+\caption{Speedup from AVX}
+\label{fig:speedup_sse}
+\end{figure}
+
 \section{Conclusions}
@@
 \end{itemize}
-
 \citet{foreman-mackey_etal_13}
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2016-07-20 paper/ms.tex
Working on the complexity

@@
 \newcommand{\emcee}{\texttt emcee}
+\newcommand{\corrfunc}{{\tt Corrfunc}\xspace}
 \newcommand{\rmax}{\ensuremath{{{\mathcal{R}}_{max}}}\xspace}
 \newcommand{\rmaxcubed}{\ensuremath{{\mathcal{R}^{3}_{max}}}\xspace}
+\newcommand{\lbox}{\ensuremath{\mathcal{L}}\xspace}
+\newcommand{\npart}{\ensuremath{\mathcal{N}}\xspace}
 \newcommand{\xir}{\ensuremath{{DD(r)}}\xspace}
@@
 
-\title{Cache is King: Presenting {\tt Corrfunc} -- a Suite of Fast Correlation Function Codes}
+\title{Cache is King: Presenting \corrfunc -- a Suite of Fast Correlation Function Codes}
 
@@
 vectorization with Single Instruction Multiple Data (SIMD) capable wide vector registers. 
-In this paper, I present {\tt Corrfunc} --  a suite of OpenMP parallelized, clustering
+In this paper, I present \corrfunc --  a suite of OpenMP parallelized, clustering
 codes that exploit current CPU micro-architecture with hand-written Advanced
@@
 Given a 3-D distribution of points in a cartesian system (i.e.,
-generated from cosmological simulations), {\tt Corrfunc} can compute spatial
+generated from cosmological simulations), \corrfunc can compute spatial
 3-D correlation functions \xir, \xiofr, \xirppi; 2-point 
@@
 projected correlation function \ddrppi, angular correlation function \wtheta as
-well as \cis.  By design, {\tt Corrfunc} is highly optimized 
+well as \cis.  By design, \corrfunc is highly optimized 
 and can compute \wprp for $\mathcal{O}$(1 million) galaxies in
 $\sim 6$ seconds on a post-2011 CPU, which is a factor of few faster than 
-existing public correlation function routines.  {\tt Corrfunc} is
+existing public correlation function routines.  \corrfunc is
 publicly available at \url{https://github.com/manodeep/Corrfunc/} while
@@
 \section{PACKAGE SUMMARY}
-{\tt Corrfunc} is written primarily in {\tt C} and comes with convenient
+\corrfunc is written primarily in {\tt C} and comes with convenient
 {\tt python} wrappers for typical clustering statistics. The fundamental design
-principles for {\tt Corrfunc} are the following:
+principles for \corrfunc are the following:
 \begin{enumerate}
-  \item Correct -- {\tt Corrfunc} has a base set of correct outputs for every statistic
+  \item Correct -- \corrfunc has a base set of correct outputs for every statistic
     generated either through slow, brute-force methods or independent, 
-    external codes. Within {\tt Corrfunc}, every clustering statistic is covered with at least one
+    external codes. Within \corrfunc, every clustering statistic is covered with at least one
     test case that {\em requires} reproducing this `known-correct' result {\em
@@
     on a variety of platforms to optimize the innermost loops. 
-  \item Portable -- {\tt Corrfunc} is written in {\tt ISO/IEC 9899:1999}
+  \item Portable -- \corrfunc is written in {\tt ISO/IEC 9899:1999}
     compliant {\tt C}. All instruction set specific codes are protected via compile-time constant
@@
 \end{enumerate}
-Every clustering statistic in {\tt Corrfunc} can either be accessed through the
+Every clustering statistic in \corrfunc can either be accessed through the
 API call (e.g., via {\tt python}), linking with the static libraries) or as
@@
 We can now compute the theoretical complexity of the proposed algorithm. Let
-$\mathcal{L}$ be the side-length of the cube over which $\mathcal{N}$ points are distributed, then
+\lbox be the side-length of the cube over which $\mathcal{N}$ points are distributed, then
 each cell contains $r_{\mathrm max}^3\times \mathcal{N/V}$ particles. To compute the
@@
 the points in the neighboring cells. This results in a complexity of 
-$\mathcal{O(N{\dfrac{\rmax}/L}^3}$. A typical \rmax is $\sim 10-20\% \times
-\mathcal{L}$, therefore ordering the particles in cells of size \rmax should
-result in a speedup of $(\rmax/\mathcal{L})^3 \sim 125-1000$ compared to the
-brute-force algorithm.
+$\mathcal{O(NM)}$, where $\mathcal{M} = \mathcal{N}\times\left(\rmax/\lbox\right)^3$. Typical \rmax is $\sim 0.10-0.2\times
+\lbox$, therefore ordering the particles in cells of size \rmax should
+result in a speedup of $(\rmax/\lbox)^3 \sim 125-1000$ compared to the
+brute-force algorithm. Since every possible pair within \rmax has to be
+examined, the algorithm deteriorates to $\mathcal{O(N}^2)$ for \rmax comparable
+to \lbox. Space partitioning tree based approaches are far more suitable for
+cases where \rmax is $\gtrsim \lbox/3$~\citep[e.g.,][]{mlpack2013,
+  feng_kdcount}.
 

### 2016-07-21 paper/ms.tex
paper so far

@@
 \newcommand{\rmaxcubed}{\ensuremath{{\mathcal{R}^{3}_{max}}}\xspace}
+\newcommand{\rsep}{\ensuremath{{{\mathcal{R}}_{sep}}}\xspace}
 \newcommand{\lbox}{\ensuremath{\mathcal{L}}\xspace}
@@
 to \lbox. Space partitioning tree based approaches are far more suitable for
-cases where \rmax is $\gtrsim \lbox/3$~\citep[e.g.,][]{mlpack2013,
+cases where \rmax is $\gtrsim 0.5\times\lbox$~\citep[e.g.,][]{mlpack2013,
   feng_kdcount}.
@@
 \subsection{Partitioning the Particles based on \rmax}
-Fig.~\ref{fig:grid} shows a schematic 2-D grid that corresponds to our partitioning scheme.
+Fig.~\ref{fig:grid} shows a schematic 2-D grid for the partitioning scheme. The
+red circles represents the reference Poisson distributed points while the blue filled circle
+shows the query point. The grid is constructed in the exact same way for both
+datasets such that the 2-D (the index is in 3-D for the actual code)  index for the query point is identical to that of
+the reference dataset. Once the central cell is determined, {\em any} reference
+point that satisfies the distance inequality $\rsep < \rmax$, {\em must} lie
+within the shaded 9 cells (corresponds to 27 in 3-D). 
 \begin{figure}[htbp]

### 2016-09-13 Corrfunc/call_correlation_functions.py
Renamed the xi_theory and xi_mocks directories

@@ -41,7 +41,7 @@ def main():
     nthreads = 4
     pimax = 40.0
     binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                    "../xi_theory/tests/", "bins")
+                    "../theory/tests/", "bins")
     autocorr = 1
     periodic = 1
 

### 2016-09-13 Corrfunc/theory/DDrppi.py

@@ -3,7 +3,7 @@
 
 """
 Python wrapper around the C extension for the pair counter in
-``xi_theory/xi_rp_pi``. `~Corrfunc.theory.DDrppi`
+``theory/xi_rp_pi``. `~Corrfunc.theory.DDrppi`
 """
 
 from __future__ import (division, print_function, absolute_import,
@@ -25,7 +25,7 @@ def DDrppi(autocorr, nthreads, pimax, binfile, X1, Y1, Z1,
 
     Note, that this module only returns pair counts and not the actual
     correlation function :math:`\\xi(r_p, \pi)`. See the
-    ``xi_theory/xi_rp_pi/wprp.c`` for computing :math:`wp(r_p)` from
+    ``theory/xi_rp_pi/wprp.c`` for computing :math:`wp(r_p)` from
     the pair counts returned.
      
     Parameters
@@ -127,7 +127,7 @@ def DDrppi(autocorr, nthreads, pimax, binfile, X1, Y1, Z1,
     >>> import Corrfunc
     >>> from Corrfunc.theory import DDrppi
     >>> binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                        "../xi_theory/tests/", "bins")
+                        "../theory/tests/", "bins")
     >>> N = 10000
     >>> boxsize = 420.0
     >>> nthreads = 4
@@ -217,7 +217,7 @@ if __name__ == '__main__':
     import time
     from os.path import dirname, abspath, join as pjoin
     binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                    "../xi_theory/tests/", "bins")
+                    "../theory/tests/", "bins")
     
     N = 100000
     boxsize = 420.0

### 2016-09-13 Corrfunc/theory/wp.py

@@ -116,7 +116,7 @@ def wp(boxsize, pimax, nthreads, binfile, X, Y, Z, verbose=False,
     >>> import Corrfunc
     >>> from Corrfunc.theory import wp
     >>> binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                        "../xi_theory/tests/", "bins")
+                        "../theory/tests/", "bins")
     >>> N = 10000
     >>> boxsize = 420.0
     >>> pimax = 40.0
@@ -185,7 +185,7 @@ if __name__ == '__main__':
     import Corrfunc
     import time
     binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                    "../xi_theory/tests/", "bins")
+                    "../theory/tests/", "bins")
     
     N = 100000
     boxsize = 420.0

### 2016-09-13 mocks/python_bindings/_countpairs_mocks.c

@@ -86,7 +86,7 @@ static PyMethodDef module_methods[] = {
      "are set to 1.0 Mpc/h.\n"
      "\n"
      "Note, that this module only returns pair counts and not the actual correlation function\n"
-     ""XI_UNICODE"(rp, "PI_UNICODE"). See the xi_mocks/DDrppi/wprp_mocks.c for computing wp(rp) from DD(rp, "PI_UNICODE").\n"
+     ""XI_UNICODE"(rp, "PI_UNICODE"). See the mocks/DDrppi/wprp_mocks.c for computing wp(rp) from DD(rp, "PI_UNICODE").\n"
      "\n"
      "parameters (all parameters are required)\n"
      "----------------------------------------\n"
@@ -152,17 +152,17 @@ static PyMethodDef module_methods[] = {
      "is the upper limit of the "PI_UNICODE" values that were considered in that (rp, "PI_UNICODE") bin.\n"
      "``npairs``contains the number of pairs in that bin and can be used to compute the\n"
      "actual "XI_UNICODE"(rp, "PI_UNICODE") by combining with RR counts.\n"
-     "(The C struct is identical to the one in `xi_theory/xi_rp_pi/countpairs_rp_pi.h`)\n"
+     "(The C struct is identical to the one in `theory/xi_rp_pi/countpairs_rp_pi.h`)\n"
      "\n"
      "example\n"
      "-------\n"
      "import numpy as np\n"
      "from Corrfunc._countpairs_mocks import countpairs_rp_pi_mocks\n"
-     "ra,dec,cz = np.genfromtxt('../xi_mocks/tests/data/Mr19_mock_northonly.rdcz.dat',dtype=np.float,unpack=True)\n"
+     "ra,dec,cz = np.genfromtxt('../mocks/tests/data/Mr19_mock_northonly.rdcz.dat',dtype=np.float,unpack=True)\n"
      "cosmology=1\n"
      "autocorr=1\n"
      "nthreads=4\n"
-     "binfile='../xi_mocks/tests/bins'\n"
+     "binfile='../mocks/tests/bins'\n"
      "pimax=40.0\n"
      "DD=countpairs_rp_pi_mocks(autocorr, cosmology,nthreads,pimax,binfile,ra,dec,cz,ra,dec,cz)\n"
      "\n"
@@ -178,7 +178,7 @@ static PyMethodDef module_methods[] = {
      "(i.e., seriously consider using some other code for real survey data).\n"
      "\n"
[…]

### 2016-09-13 paper/get_speedups.py

@@ -41,7 +41,7 @@ class nf(float):
 
 
 def main():
-    base_dir = '../xi_theory/wp/'
+    base_dir = '../theory/wp/'
     base_string = 'wp'
     files = ['timings_naive', 'timings_sse', 'timings_avx']
     files = [base_dir + f for f in files]

### 2016-09-13 paper/scripts/generate_figures.py

@@ -28,7 +28,7 @@ def benchmark_theory_threads_all(min_threads=1, max_threads=16,
     print("Benchmarking routines = {0}".format(keys))
     x, y, z = read_catalog()
     binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                    "../xi_theory/tests/", "bins")
+                    "../theory/tests/", "bins")
     autocorr = 1
     pimax = 40.0
     boxsize = 420.0

### 2016-09-13 Corrfunc/mocks/DDrppi_mocks.py
Renamed the xi_theory and xi_mocks directories

@@ -3,7 +3,7 @@
 
 """
 Python wrapper around the C extension for the pair counter in
-``xi_mocks/DDrppi``. This python wrapper is in `~Corrfunc.mocks.DDrppi_mocks`
+``mocks/DDrppi``. This python wrapper is in `~Corrfunc.mocks.DDrppi_mocks`
 """
 
 from __future__ import (division, print_function, absolute_import,
@@ -34,7 +34,7 @@ def DDrppi_mocks(autocorr, cosmology, nthreads, pimax, binfile,
 
     Note, that this module only returns pair counts and not the actual
     correlation function :math:`\\xi(r_p, \pi)`. See the
-    ``xi_mocks/DDrppi/wprp_mocks.c`` for computing :math:`wp(r_p)` from
+    ``mocks/DDrppi/wprp_mocks.c`` for computing :math:`wp(r_p)` from
     the pair counts returned.
      
     Parameters
@@ -197,7 +197,7 @@ def DDrppi_mocks(autocorr, cosmology, nthreads, pimax, binfile,
     >>> import math
     >>> from os.path import dirname, abspath, join as pjoin
     >>> binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                        "../xi_theory/tests/", "bins")
+                        "../mocks/tests/", "bins")
     >>> N = 100000
     >>> boxsize = 420.0
     >>> X = np.random.uniform(-0.5*boxsize, 0.5*boxsize, N)
@@ -297,7 +297,7 @@ if __name__ == '__main__':
 
     print("\nRunning 2-D correlation function for mocks DD(rp,pi)")
     binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                    "../xi_theory/tests/", "bins")
+                    "../mocks/tests/", "bins")
     
     N = 100000
     boxsize = 420.0

### 2016-09-13 Corrfunc/mocks/vpf_mocks.py

@@ -3,7 +3,7 @@
 
 """
 Python wrapper around the C extension for the counts-in-cells
-for positions on the sky. Corresponding C codes are in `xi_mocks/vpf`
+for positions on the sky. Corresponding C codes are in `mocks/vpf`
 while the python wrapper is in `~Corrfunc.mocks.vpf_mocks`
 """
 
@@ -58,7 +58,7 @@ def vpf_mocks(rmax, nbins, nspheres, numpN,
     threshold_ngb: integer
        Minimum number of random points needed in a ``rmax`` sphere such that it
        is considered to be entirely within the mock footprint. The
-       command-line version, ``xi_mocks/vpf/vpf_mocks.c``, assumes that the
+       command-line version, ``mocks/vpf/vpf_mocks.c``, assumes that the
        minimum number of randoms can be at most a 1-sigma deviation from
        the expected random number density.
 
@@ -200,7 +200,7 @@ def vpf_mocks(rmax, nbins, nspheres, numpN,
     >>> threshold_ngb = 1  # does not matter since we have the centers
     >>> cosmology = 1  # LasDamas cosmology
     >>> centers_file = pjoin(dirname(abspath(Corrfunc.__file__)),
-                             "../xi_mocks/tests/data/",
+                             "../mocks/tests/data/",
                              "Mr19_centers_xyz_forVPF_rmax_10Mpc.txt")
     >>> N = 100000
     >>> boxsize = 420.0
@@ -287,7 +287,7 @@ if __name__ == '__main__':
 
     # file containing the centers of random spheres
     centers_file = pjoin(dirname(abspath(Corrfunc.__file__)),
-                         "../xi_mocks/tests/data/",
+                         "../mocks/tests/data/",
                          "Mr19_centers_xyz_forVPF_rmax_10Mpc.txt")
     
     N = 100000

### 2016-09-13 Corrfunc/theory/xi.py

@@ -4,7 +4,7 @@
 """
 Python wrapper around the C extension for the theoretical 3-D
 real-space correlation function, :math:`\\xi(r)`. Corresponding
-C routines are in xi_theory/xi/, python interface is `~Corrfunc.theory.xi`
+C routines are in theory/xi/, python interface is `~Corrfunc.theory.xi`
 """
 
 from __future__ import (division, print_function, absolute_import,
@@ -110,7 +110,7 @@ def xi(boxsize, nthreads, binfile, X, Y, Z, verbose=False,
     >>> import Corrfunc
     >>> from Corrfunc.theory import xi
     >>> binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                        "../xi_theory/tests/", "bins")
+                        "../theory/tests/", "bins")
     >>> N = 100000
     >>> boxsize = 420.0
     >>> nthreads = 4
@@ -178,7 +178,7 @@ if __name__ == '__main__':
     import time
     from os.path import dirname, abspath, join as pjoin
     binfile = pjoin(dirname(abspath(Corrfunc.__file__)),
-                    "../xi_theory/tests/", "bins")
+                    "../theory/tests/", "bins")
     
     N = 100000
     boxsize = 420.0

### 2016-11-01 theory/wp/wp.c
Fixed typo in docs [ci skip]

@@ -120,7 +120,7 @@ int main(int argc, char *argv[])
     gettimeofday(&t1,NULL);
     read_time += ADD_DIFF_TIME(t0,t1);
 
-    //check that theee positions are within limits
+    //check that the positions are within limits
     for(int i=0;i<ND1;i++) {
         assert(x1[i] >= 0.0 && x1[i] <= boxsize && "xpos is within limits [0, boxsize]");
         assert(y1[i] >= 0.0 && y1[i] <= boxsize && "ypos is within limits [0, boxsize]");
@@ -130,6 +130,7 @@ int main(int argc, char *argv[])
     /*---Count-pairs--------------------------------------*/
     gettimeofday(&t0,NULL);
     struct config_options options = get_config_options();
+    options.verbose=0;
     /* If you want to change the bin refine factors */
     /* const int bf[] = {2, 2, 1}; */
     /* set_bin_refine_factors(&options, bf); */

### 2019-05-21 Corrfunc/mocks/DDrppi_mocks.py
AVX512F kernels, min separation optimizations, not duplicating particle positions, array of cell-pairs (#173)
* WIP: Added the machinery for quicker exits

* WIP: Adding in early exits to the theory routines

* Break out of next j-loop if any dz values in current iteration are larger than 'max_dz'

* [WIP] Started implementing the min sep optimizations for mocks

* Preparing for v2.3  (#170)

* started basic work on avx512 (completely wrong results currently)

* Fixed the number of bits set in the mask. npairs now agrees with the test, rpavg and weightavg are still wrong

* Trying mask loads.

@@ -3,7 +3,7 @@
 
 """
 Python wrapper around the C extension for the pair counter in
-``mocks/DDrppi_mocks/``. This python wrapper is 
+``mocks/DDrppi_mocks/``. This python wrapper is
 :py:mod:`Corrfunc.mocks.DDrppi_mocks`
 """
 
@@ -22,6 +22,7 @@ def DDrppi_mocks(autocorr, cosmology, nthreads, pimax, binfile,
                  fast_divide_and_NR_steps=0,
                  xbin_refine_factor=2, ybin_refine_factor=2,
                  zbin_refine_factor=1, max_cells_per_dim=100,
+                 copy_particles=True, enable_min_sep_opt=True,
                  c_api_timer=False, isa=r'fastest', weight_type=None):
     """
     Calculate the 2-D pair-counts corresponding to the projected correlation
@@ -31,7 +32,7 @@ def DDrppi_mocks(autocorr, cosmology, nthreads, pimax, binfile,
     counted. The input positions are expected to be on-sky co-ordinates.
     This module is suitable for calculating correlation functions for mock
     catalogs.
-    
+
     If ``weights`` are provided, the resulting pair counts are weighted.  The
     weighting scheme depends on ``weight_type``.
 
@@ -42,8 +43,8 @@ def DDrppi_mocks(autocorr, cosmology, nthreads, pimax, binfile,
     .. note:: that this module only returns pair counts and not the actual
        correlation function :math:`\\xi(r_p, \pi)` or :math:`wp(r_p)`. See the
        utilities :py:mod:`Corrfunc.utils.convert_3d_counts_to_cf` and
-       :py:mod:`Corrfunc.utils.convert_rp_pi_counts_to_wp` for computing 
-       :math:`\\xi(r_p, \pi)` and :math:`wp(r_p)` respectively from the 
+       :py:mod:`Corrfunc.utils.convert_rp_pi_counts_to_wp` for computing
+       :math:`\\xi(r_p, \pi)` and :math:`wp(r_p)` respectively from the
        pair counts.
 
 
@@ -75,7 +76,7 @@ def DDrppi_mocks(autocorr, cosmology, nthreads, pimax, binfile,
 
     pimax : double
         A double-precision value for the maximum separation along
-        the Z-dimension. 
+        the Z-dimension.
 
         Distances along the :math:`\\pi` direction are binned with unit
[…]

### 2019-05-21 Corrfunc/mocks/DDtheta_mocks.py

@@ -3,8 +3,8 @@
 
 """
 Python wrapper around the C extension for the angular correlation function
-:math:`\\omega(\\theta)`. Corresponding C routines are in 
-``mocks/DDtheta_mocks/``, while the python interface is 
+:math:`\\omega(\\theta)`. Corresponding C routines are in
+``mocks/DDtheta_mocks/``, while the python interface is
 :py:mod:`Corrfunc.mocks.DDtheta_mocks`
 """
 
@@ -22,6 +22,7 @@ def DDtheta_mocks(autocorr, nthreads, binfile,
                   verbose=False, output_thetaavg=False,
                   fast_acos=False, ra_refine_factor=2,
                   dec_refine_factor=2, max_cells_per_dim=100,
+                  copy_particles=True, enable_min_sep_opt=True,
                   c_api_timer=False, isa=r'fastest', weight_type=None):
     """
     Function to compute the angular correlation function for points on
@@ -29,14 +30,14 @@ def DDtheta_mocks(autocorr, nthreads, binfile,
 
     Returns a numpy structured array containing the pair counts for the
     specified angular bins.
-    
+
     If ``weights`` are provided, the resulting pair counts are weighted.  The
     weighting scheme depends on ``weight_type``.
 
 
     .. note:: This module only returns pair counts and not the actual
-       correlation function :math:`\\omega(\theta)`. See 
-       :py:mod:`Corrfunc.utils.convert_3d_counts_to_cf` for computing 
+       correlation function :math:`\\omega(\theta)`. See
+       :py:mod:`Corrfunc.utils.convert_3d_counts_to_cf` for computing
        :math:`\\omega(\theta)` from the pair counts returned.
 
 
@@ -48,7 +49,7 @@ def DDtheta_mocks(autocorr, nthreads, binfile,
         then the second set of particle positions are not required.
 
     nthreads : integer
-       Number of threads to use.
+        Number of threads to use.
 
     binfile: string or an list/array of floats. Units: degrees.
         For string input: filename specifying the ``theta`` bins for
@@ -60,7 +61,7 @@ def DDtheta_mocks(autocorr, nthreads, binfile,
         bin-edges. For example,
         ``np.logspace(np.log10(0.1), np.log10(10.0), 15)`` is a valid
         input specifying **14** (logarithmic) bins between 0.1 and 10.0
-        degrees. This array does not need to be sorted.         
+        degrees. This array does not need to be sorted.
 
[…]

### 2019-05-21 Corrfunc/mocks/vpf_mocks.py

@@ -21,7 +21,7 @@ def vpf_mocks(rmax, nbins, nspheres, numpN,
               verbose=False, is_comoving_dist=False,
               xbin_refine_factor=1, ybin_refine_factor=1,
               zbin_refine_factor=1, max_cells_per_dim=100,
-              c_api_timer=False, isa=r'fastest'):
+              copy_particles=True, c_api_timer=False, isa=r'fastest'):
     """
     Function to compute the counts-in-cells on points on the sky. Suitable
     for mock catalogs and observed galaxies.
@@ -33,55 +33,55 @@ def vpf_mocks(rmax, nbins, nspheres, numpN,
     ----------
 
     rmax : double
-       Maximum radius of the sphere to place on the particles
+        Maximum radius of the sphere to place on the particles
 
     nbins : integer
-       Number of bins in the counts-in-cells. Radius of first shell
-       is rmax/nbins
+        Number of bins in the counts-in-cells. Radius of first shell
+        is rmax/nbins
 
     nspheres : integer (>= 0)
-       Number of random spheres to place within the particle distribution.
-       For a small number of spheres, the error is larger in the measured
-       pN's.
+        Number of random spheres to place within the particle distribution.
+        For a small number of spheres, the error is larger in the measured
+        pN's.
 
     numpN : integer (>= 1)
-       Governs how many unique pN's are to returned. If ``numpN`` is set to 1,
-       then only the vpf (p0) is returned. For ``numpN=2``, p0 and p1 are
-       returned.
+        Governs how many unique pN's are to returned. If ``numpN`` is set to 1,
+        then only the vpf (p0) is returned. For ``numpN=2``, p0 and p1 are
+        returned.
 
-       More explicitly, the columns in the results look like the following:
+        More explicitly, the columns in the results look like the following:
 
-         ======   ==========================
-         numpN    Columns in output
-         ======   ==========================
-            1      p0
-            2      p0      p1
-            3      p0      p1     p2
-            4      p0      p1     p2     p3
-         ======   ==========================
+          ======   ==========================
+          numpN    Columns in output
+          ======   ==========================
+             1      p0
+             2      p0      p1
+             3      p0      p1     p2
+             4      p0      p1     p2     p3
+          ======   ==========================
 
-       and so on...
+        and so on...
 
-       Note: ``p0`` is the vpf
+        Note: ``p0`` is the vpf
 
     threshold_ngb : integer
[…]

### 2019-05-21 Corrfunc/theory/DD.py

@@ -17,19 +17,20 @@ def DD(autocorr, nthreads, binfile, X1, Y1, Z1, weights1=None, periodic=True,
        X2=None, Y2=None, Z2=None, weights2=None, verbose=False, boxsize=0.0,
        output_ravg=False, xbin_refine_factor=2, ybin_refine_factor=2,
        zbin_refine_factor=1, max_cells_per_dim=100,
+       copy_particles=True, enable_min_sep_opt=True,
        c_api_timer=False, isa=r'fastest', weight_type=None):
     """
     Calculate the 3-D pair-counts corresponding to the real-space correlation
     function, :math:`\\xi(r)`.
-    
+
     If ``weights`` are provided, the resulting pair counts are weighted.  The
     weighting scheme depends on ``weight_type``.
 
 
     .. note:: This module only returns pair counts and not the actual
-       correlation function :math:`\\xi(r)`. See 
-       :py:mod:`Corrfunc.utils.convert_3d_counts_to_cf` for computing 
-       for computing :math:`\\xi(r)` from the pair counts returned.
+        correlation function :math:`\\xi(r)`. See
+        :py:mod:`Corrfunc.utils.convert_3d_counts_to_cf` for computing
+        :math:`\\xi(r)` from the pair counts returned.
 
 
     Parameters
@@ -53,30 +54,31 @@ def DD(autocorr, nthreads, binfile, X1, Y1, Z1, weights1=None, periodic=True,
         bin-edges. For example,
         ``np.logspace(np.log10(0.1), np.log10(10.0), 15)`` is a valid
         input specifying **14** (logarithmic) bins between 0.1 and 10.0. This
-        array does not need to be sorted.         
-    
+        array does not need to be sorted.
+
     X1/Y1/Z1: array_like, real (float/double)
         The array of X/Y/Z positions for the first set of points.
         Calculations are done in the precision of the supplied arrays.
-        
+
     weights1: array_like, real (float/double), optional
-        A scalar, or an array of weights of shape (n_weights, n_positions) or (n_positions,).
-        `weight_type` specifies how these weights are used; results are returned
-        in the `weightavg` field.  If only one of weights1 and weights2 is
-        specified, the other will be set to uniform weights.
+        A scalar, or an array of weights of shape (n_weights, n_positions) or
+        (n_positions,). ``weight_type`` specifies how these weights are used;
+        results are returned in the ``weightavg`` field.  If only one of
+        weights1 and weights2 is specified, the other will be set to uniform
+        weights.
 
     periodic: boolean
[…]

### 2019-05-21 Corrfunc/theory/DDrppi.py

@@ -18,6 +18,7 @@ def DDrppi(autocorr, nthreads, pimax, binfile, X1, Y1, Z1, weights1=None,
            verbose=False, boxsize=0.0, output_rpavg=False,
            xbin_refine_factor=2, ybin_refine_factor=2,
            zbin_refine_factor=1, max_cells_per_dim=100,
+           copy_particles=True, enable_min_sep_opt=True,
            c_api_timer=False, isa=r'fastest', weight_type=None):
     """
     Calculate the 3-D pair-counts corresponding to the real-space correlation
@@ -25,17 +26,17 @@ def DDrppi(autocorr, nthreads, pimax, binfile, X1, Y1, Z1, weights1=None,
     separated by less than the ``rp`` bins (specified in ``binfile``) in the
     X-Y plane, and less than ``pimax`` in the Z-dimension are
     counted.
-    
+
     If ``weights`` are provided, the resulting pair counts are weighted.  The
     weighting scheme depends on ``weight_type``.
 
 
     .. note:: that this module only returns pair counts and not the actual
-       correlation function :math:`\\xi(r_p, \pi)` or :math:`wp(r_p)`. See the
-       utilities :py:mod:`Corrfunc.utils.convert_3d_counts_to_cf` and
-       :py:mod:`Corrfunc.utils.convert_rp_pi_counts_to_wp` for computing 
-       :math:`\\xi(r_p, \pi)` and :math:`wp(r_p)` respectively from the 
-       pair counts.
+        correlation function :math:`\\xi(r_p, \pi)` or :math:`wp(r_p)`. See the
+        utilities :py:mod:`Corrfunc.utils.convert_3d_counts_to_cf` and
+        :py:mod:`Corrfunc.utils.convert_rp_pi_counts_to_wp` for computing
+        :math:`\\xi(r_p, \pi)` and :math:`wp(r_p)` respectively from the
+        pair counts.
 
 
     Parameters
@@ -50,15 +51,14 @@ def DDrppi(autocorr, nthreads, pimax, binfile, X1, Y1, Z1, weights1=None,
         enabled during library compilation.
 
     pimax: double
-       A double-precision value for the maximum separation along
-       the Z-dimension. 
+        A double-precision value for the maximum separation along
+        the Z-dimension.
 
-       Distances along the :math:``\\pi`` direction are binned with unit
-       depth. For instance, if ``pimax=40``, then 40 bins will be created
-       along the ``pi`` direction.
[…]

### 2019-05-21 mocks/DDrppi_mocks/countpairs_rp_pi_mocks.c

@@ -19,10 +19,14 @@ void free_results_mocks(results_countpairs_mocks *results)
     if(results==NULL)
         return;
 
-    free(results->npairs);
-    free(results->rupp);
-    free(results->rpavg);
-    free(results->weightavg);
+    free(results->npairs);results->npairs = NULL;
+    free(results->rupp);results->rupp = NULL;
+    free(results->rpavg);results->rpavg = NULL;
+    free(results->weightavg);results->weightavg = NULL;
+
+    results->nbin = 0;
+    results->npibin = 0;
+    results->pimax = 0.0;
 }
 
 
@@ -46,7 +50,7 @@ int countpairs_mocks(const int64_t ND1, void *phi1, void *theta1, void *czD1,
         fprintf(stderr,"Error: Do not know this API version = `%s'. Expected version = `%s'\n", options->version, STR(VERSION));
         return EXIT_FAILURE;
     }
-    
+
     if(options->float_type == sizeof(float)) {
         return countpairs_mocks_float(ND1, (float *) phi1, (float *) theta1, (float *) czD1,
                                       ND2, (float *) phi2, (float *) theta2, (float *) czD2,

### 2019-05-21 mocks/tests/tests_mocks.c

@@ -31,10 +31,10 @@ int test_DDsmu_mocks(const char *correct_outputfile);
 void read_data_and_set_globals(const char *firstfilename, const char *firstformat,const char *secondfilename,const char *secondformat);
 
 //Global variables
-int ND1;
+int64_t ND1;
 double *RA1=NULL,*DEC1=NULL,*CZ1=NULL,*weights1=NULL;
 
-int ND2;
+int64_t ND2;
 double *RA2=NULL,*DEC2=NULL,*CZ2=NULL,*weights2=NULL;
 
 const int cosmology_flag=1;
@@ -92,7 +92,7 @@ int test_DDrppi_mocks(const char *correct_outputfile)
                 }
                 int floats_equal = AlmostEqualRelativeAndAbs_double(rpavg, results.rpavg[index], maxdiff, maxreldiff);
                 int weights_equal = AlmostEqualRelativeAndAbs_double(weightavg, results.weightavg[index], maxdiff, maxreldiff);
-                
+
                 //Check for exact equality of npairs and float "equality" for rpavg
                 if(npairs == results.npairs[index] && floats_equal == EXIT_SUCCESS && weights_equal == EXIT_SUCCESS) {
                     ret = EXIT_SUCCESS;
@@ -107,10 +107,10 @@ int test_DDrppi_mocks(const char *correct_outputfile)
             }
         }
         fclose(fp);
-    END_INTEGRATION_TEST_SECTION;
-        
+    END_INTEGRATION_TEST_SECTION(free_results_mocks(&results));
+
     /* If the test failed, then write the output into a temporary file */
-    if(ret != EXIT_SUCCESS) {
+    if(ret != EXIT_SUCCESS && results.nbin > 0) {
         FILE *fp=my_fopen(tmpoutputfile,"w");
         if(fp == NULL) {
             free_results_mocks(&results);
@@ -136,7 +136,7 @@ int test_DDsmu_mocks(const char *correct_outputfile)
 {
     results_countpairs_mocks_s_mu results;
     int ret = EXIT_FAILURE;
-    
+
     assert(RA1 != NULL && DEC1 != NULL && CZ1 != NULL && "ERROR: In test suite for DDsmu ra/dec/cz can not be NULL pointers");
     int autocorr = (RA1==RA2) ? 1:0;
 
@@ -184,7 +184,7 @@ int test_DDsmu_mocks(const char *correct_outputfile)
                 }
[…]

### 2019-05-21 mocks/vpf_mocks/countspheres_mocks.c

@@ -22,6 +22,11 @@ void free_results_countspheres_mocks(results_countspheres_mocks *results)
         return;
 
     matrix_free((void **) results->pN, results->nbin);
+    results->pN = NULL;
+    results->rmax = 0.0;
+    results->nbin = 0;
+    results->nc = 0;
+    results->num_pN = 0;
 }
 
 
@@ -48,10 +53,10 @@ int countspheres_mocks(const int64_t Ngal, void *xgal, void *ygal, void *zgal,
         fprintf(stderr,"Error: Do not know this API version = `%s'. Expected version = `%s'\n", options->version, STR(VERSION));
         return EXIT_FAILURE;
     }
-    
+
     if(options->float_type == sizeof(float)) {
         return countspheres_mocks_float(Ngal, (float *) xgal, (float *)ygal, (float *)zgal,
-                                        Nran, (float * restrict) xran, (float * restrict) yran, (float * restrict) zran,
+                                        Nran, (float *) xran, (float *) yran, (float *) zran,
                                         threshold_neighbors,
                                         rmax, nbin, nc,
                                         num_pN,
@@ -62,7 +67,7 @@ int countspheres_mocks(const int64_t Ngal, void *xgal, void *ygal, void *zgal,
                                         extra);
     } else {
         return countspheres_mocks_double(Ngal, (double *) xgal, (double *)ygal, (double *)zgal,
-                                         Nran, (double * restrict) xran, (double * restrict) yran, (double * restrict) zran,
+                                         Nran, (double *) xran, (double *) yran, (double *) zran,
                                          threshold_neighbors,
                                          rmax, nbin, nc,
                                          num_pN,

### 2019-05-21 paper/scripts/generate_numpart_scaling.py

@@ -4,8 +4,10 @@ from __future__ import print_function
 import numpy as np
 
 import Corrfunc
-
 from Corrfunc.io import read_catalog
+
+from utils import convert_numpy_bytes_to_unicode
+
 from os.path import join as pjoin, abspath, dirname
 import time
 import sys
@@ -76,7 +78,7 @@ def benchmark_theory_threads_all(numpart_frac=[0.001, 0.005, 0.01, 0.05, 0.1, 0.
     from Corrfunc.theory import DD, DDrppi, wp, xi
     allkeys = [#'DD', 'DDrppi',
                'wp', 'xi']
-    allisa = ['avx', 'sse42', 'fallback']
+    allisa = ['avx512f', 'avx', 'sse42', 'fallback']
     if keys is None:
         keys = allkeys
     else:
@@ -111,8 +113,8 @@ def benchmark_theory_threads_all(numpart_frac=[0.001, 0.005, 0.01, 0.05, 0.1, 0.
     nthreads = max_threads
     
     dtype = np.dtype([('repeat', np.int),
-                      ('name', 'S16'),
-                      ('isa', 'S16'),
+                      ('name', 'U16'),
+                      ('isa', 'U16'),
                       ('rmax', np.float),
                       ('ndata',np.int),
                       ('nrand',np.int),
@@ -253,7 +255,7 @@ def benchmark_mocks_threads_all(numpart_frac=[0.001, 0.005, 0.01, 0.05, 0.1, 0.2
                'DDrppi (DR)',
                'DDtheta (DR)'
               ]
-    allisa = ['avx', 'sse42', 'fallback']
+    allisa = ['avx512f', 'avx', 'sse42', 'fallback']
     if keys is None:
         keys = allkeys
     else:
@@ -294,8 +296,8 @@ def benchmark_mocks_threads_all(numpart_frac=[0.001, 0.005, 0.01, 0.05, 0.1, 0.2
 
     nthreads = max_threads
     dtype = np.dtype([('repeat', np.int),
-                      ('name', 'S16'),
-                      ('isa', 'S16'),
+                      ('name', 'U16'),
+                      ('isa', 'U16'),
                       ('rmax', np.float),
                       ('ndata',np.int),
                       ('nrand',np.int),
@@ -439,17 +441,17 @@ def benchmark_mocks_threads_all(numpart_frac=[0.001, 0.005, 0.01, 0.05, 0.1, 0.2
[…]

### 2019-05-21 paper/scripts/utils.py

@@ -1,0 +1,20 @@
+def convert_numpy_bytes_to_unicode(runtimes):
+    '''
+    Our old runtime arrays use the S dtype descriptor, which
+    in Python 3 becomes bytes, making str comparison
+    more cumbersome.  We can just change the dtype of
+    the string fields to be Unicode.  Doing so on
+    a structured arrays is also cumbersome, hence
+    this function.
+
+    Our new runtime arrays just use Unicode by default.
+    '''
+    runtimes_descr = runtimes.dtype.descr
+    runtimes_descr = [list(d) for d in runtimes_descr]
+    for d in runtimes_descr:
+        if 'S' in d[1]:
+            d[1] = d[1].replace('S','U')
+    runtimes_descr = [tuple(d) for d in runtimes_descr]
+    runtimes = runtimes.astype(runtimes_descr)
+
+    return runtimes

### 2019-05-21 theory/DD/DD.c

@@ -4,7 +4,7 @@
   Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
   License: MIT LICENSE. See LICENSE file under the top-level
   directory at https://github.com/manodeep/Corrfunc/
-  
+
 --- DD file1 format1 file2 format2 binfile numthreads [weight_method weights_file1 weights_format1 [weights_file2 weights_format2]] > DDfile
 --- Measure the cross-correlation function DD(r) for two different
    data files (or autocorrelation if file1=file2).
@@ -51,7 +51,7 @@ int main(int argc, char *argv[])
     char *fileformat1=NULL,*fileformat2=NULL,*weights_fileformat1=NULL,*weights_fileformat2=NULL;
     char *binfile=NULL;
     char *weight_method_str=NULL;
-    
+
     weight_method_t weight_method = NONE;
     int num_weights = 0;
 
@@ -64,7 +64,7 @@ int main(int argc, char *argv[])
     const char argnames[][30]={"file1","format1","file2","format2","binfile","Nthreads"};
 #endif
     const char optargnames[][30]={"weight_method", "weights_file1","weights_format1","weights_file2","weights_format2"};
-    
+
     int nargs=sizeof(argnames)/(sizeof(char)*30);
     int noptargs=sizeof(optargnames)/(sizeof(char)*30);
 
@@ -91,7 +91,7 @@ int main(int argc, char *argv[])
             fprintf(stderr,"\t\t %s = `?'\n",argnames[i-1]);
         return EXIT_FAILURE;
     }
-  
+
     /* Validate optional arguments */
     int noptargs_given = argc - (nargs + 1);
     if(noptargs_given != 0 && noptargs_given != 3 && noptargs_given != 5){
@@ -121,7 +121,7 @@ int main(int argc, char *argv[])
         return EXIT_FAILURE;
     }
 #endif
-  
+
     if(noptargs_given >= 3){
        weight_method_str = argv[nargs + 1];
        int wstatus = get_weight_method_by_name(weight_method_str, &weight_method);
@@ -130,7 +130,7 @@ int main(int argc, char *argv[])
          return EXIT_FAILURE;
        }
        num_weights = get_num_weights_by_method(weight_method);
-      
+
        weights_file1 = argv[nargs + 2];
        weights_fileformat1 = argv[nargs + 3];
     }
@@ -164,14 +164,14 @@ int main(int argc, char *argv[])
[…]

### 2019-05-21 theory/python_bindings/_countpairs.c

@@ -87,8 +87,8 @@ static PyMethodDef module_methods[] = {
      "countpairs(autocorr, nthreads, binfile, X1, Y1, Z1, weights1=None, weight_type=None, periodic=True,\n"
      "           X2=None, Y2=None, Z2=None, weights2=None, verbose=False, boxsize=0.0,\n"
      "           output_ravg=False, xbin_refine_factor=2, ybin_refine_factor=2,\n"
-     "           zbin_refine_factor=1, max_cells_per_dim=100, c_api_timer=False,\n"
-     "           isa=-1)\n"
+     "           zbin_refine_factor=1, max_cells_per_dim=100, copy_particles=True,\n"
+     "           enable_min_sep_opt=True, c_api_timer=False, isa=-1)\n"
      "\n"
      "Calculate the 3-D pair-counts, "XI_CHAR"(r), auto/cross-correlation \n"
      "function given two sets of points represented by X1/Y1/Z1 and X2/Y2/Z2 \n"
@@ -105,114 +105,124 @@ static PyMethodDef module_methods[] = {
      "Every parameter can be passed as a keyword of the corresponding name.\n\n"
 
      "autocorr : boolean\n"
-     "   Boolean flag for auto/cross-correlation. If autocorr is set to 1,\n"
-     "    are not used (but must still be passed, perhaps again as X1/Y1/Z1).\n"
-     "\n"
+     "    Boolean flag for auto/cross-correlation. If autocorr is set to 1,\n"
+     "    are not used (but must still be passed, perhaps again as X1/Y1/Z1).\n\n"
+
      "nthreads : integer\n"
-     "   The number of OpenMP threads to use. Has no effect if OpenMP was not\n"
-     "   enabled during library compilation.\n\n"
+     "    The number of OpenMP threads to use. Has no effect if OpenMP was not\n"
+     "    enabled during library compilation.\n\n"
 
      "binfile : string\n"
-     "   Filename specifying the ``r`` bins for ``DD``. The file should\n"
-     "   contain white-space separated values  of (rmin, rmax)  for each\n"
-     "   ``r`` wanted. The bins do not need to be contiguous but must be in\n"
-     "   increasing order (smallest bins come first). \n\n"
+     "    Filename specifying the ``r`` bins for ``DD``. The file should\n"
+     "    contain white-space separated values  of (rmin, rmax)  for each\n"
+     "    ``r`` wanted. The bins do not need to be contiguous but must be in\n"
+     "    increasing order (smallest bins come first). \n\n"
 
[…]

### 2019-05-21 theory/wp/countpairs_wp.c

@@ -20,11 +20,14 @@ void free_results_wp(results_countpairs_wp *results)
     if(results == NULL)
         return;
 
-    free(results->npairs);
-    free(results->rupp);
-    free(results->wp);
-    free(results->rpavg);
-    free(results->weightavg);
+    free(results->rupp);results->rupp = NULL;
+    free(results->npairs);results->npairs = NULL;
+    free(results->rpavg);results->rpavg = NULL;
+    free(results->weightavg);results->weightavg = NULL;
+    free(results->wp);results->wp = NULL;
+
+    results->nbin = 0;
+    results->pimax = 0.0;
 }
 
 int countpairs_wp(const int64_t ND, void * restrict X, void * restrict Y, void * restrict Z,
@@ -41,14 +44,14 @@ int countpairs_wp(const int64_t ND, void * restrict X, void * restrict Y, void *
                 __FUNCTION__, options->float_type);
         return EXIT_FAILURE;
     }
-    
+
     if( strncmp(options->version, STR(VERSION), sizeof(options->version)/sizeof(char)-1 ) != 0) {
         fprintf(stderr,"Error: Do not know this API version = `%s'. Expected version = `%s'\n", options->version, STR(VERSION));
         return EXIT_FAILURE;
     }
 
     if(options->float_type == sizeof(float)) {
-      return countpairs_wp_float(ND, (float * restrict) X, (float * restrict) Y, (float * restrict) Z,
+      return countpairs_wp_float(ND, (float *) X, (float *) Y, (float *) Z,
                                  boxsize,
                                  numthreads,
                                  binfile,
@@ -57,7 +60,7 @@ int countpairs_wp(const int64_t ND, void * restrict X, void * restrict Y, void *
                                  options,
                                  extra);
     } else {
-      return countpairs_wp_double(ND, (double * restrict) X, (double * restrict) Y, (double * restrict) Z,
+      return countpairs_wp_double(ND, (double *) X, (double *) Y, (double *) Z,
                                   boxsize,
                                   numthreads,
                                   binfile,

### 2019-05-21 theory/xi/countpairs_xi.c

@@ -20,11 +20,13 @@ void free_results_xi(results_countpairs_xi *results)
     if(results == NULL)
         return;
 
-    free(results->rupp);
-    free(results->xi);
-    free(results->npairs);
-    free(results->ravg);
-    free(results->weightavg);
+    free(results->rupp);results->rupp = NULL;
+    free(results->xi);results->xi = NULL;
+    free(results->npairs);results->npairs = NULL;
+    free(results->ravg);results->ravg = NULL;
+    free(results->weightavg);results->weightavg = NULL;
+
+    results->nbin = 0;
 }
 
 
@@ -45,9 +47,9 @@ int countpairs_xi(const int64_t ND, void * restrict X, void * restrict Y, void *
         fprintf(stderr,"Error: Do not know this API version = `%s'. Expected version = `%s'\n", options->version, STR(VERSION));
         return EXIT_FAILURE;
     }
-    
+
     if(options->float_type == sizeof(float)) {
-        return countpairs_xi_float(ND, (float * restrict) X, (float * restrict) Y, (float * restrict) Z,
+        return countpairs_xi_float(ND, (float *) X, (float *) Y, (float *) Z,
                                    boxsize,
                                    numthreads,
                                    binfile,
@@ -55,7 +57,7 @@ int countpairs_xi(const int64_t ND, void * restrict X, void * restrict Y, void *
                                    options,
                                    extra);
     } else {
-        return countpairs_xi_double(ND, (double * restrict) X, (double * restrict) Y, (double * restrict) Z,
+        return countpairs_xi_double(ND, (double *) X, (double *) Y, (double *) Z,
                                     boxsize,
                                     numthreads,
                                     binfile,

### 2019-05-21 utils/avx512_calls.h

@@ -1,0 +1,417 @@
+/* File: avx512_calls.h */
+/*
+  This file is a part of the Corrfunc package
+  Copyright (C) 2015-- Manodeep Sinha (manodeep@gmail.com)
+  License: MIT LICENSE. See LICENSE file under the top-level
+  directory at https://github.com/manodeep/Corrfunc/
+*/
+
+#pragma once
+
+#if defined(__AVX512F__)
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <immintrin.h>
+
+#ifdef __cplusplus
+extern "C" {
+#endif
+
+#include "function_precision.h" 
+
+#define PREFETCH(mem)        asm ("prefetcht0 %0"::"m"(mem))
+
+#if defined(__GNUC__) || defined(__GNUG__)
+#define AVX512_BIT_COUNT_INT(X)                __builtin_popcount(X)
+#else
+#define AVX512_BIT_COUNT_INT(X)                _popcnt32(X)
+#endif
+
+
+#define AVX512_BIT_COUNT_LONG(X)               _popcnt64(X)
+#define AVX512_BIT_COUNT_UNSIGNED_INT(X)       _mm_popcnt_u32(X)
+#define AVX512_BIT_COUNT_UNSIGNED_LONG(X)      _mm_popcnt_u64(X)
+
+
+#define AVX512_MASK_BITWISE_AND(X,Y)              _mm512_kand(X,Y)
+#define AVX512_MASK_BITWISE_OR(X,Y)               _mm512_kor(X,Y)
+#define AVX512_MASK_BITWISE_XOR_FLOATS(X,Y)       _mm512_kxor(X,Y)
+#define AVX512_MASK_BITWISE_AND_NOT(X,Y)          _mm512_kandn(X,Y)  //~X & Y
+#define AVX512_MASK_BITWISE_NOT(X,Y)              _mm512_knot(X) //~X
+
+    
+
+  /* For setting up the array that contains the number bits set in a mask */
+  // Can be used to generate up to 16 bit lookup tables
+#   define B2(n)  n,      n+1,      n+1,      n+2
+#   define B4(n)  B2(n),  B2(n+1),  B2(n+1),  B2(n+2)
+#   define B6(n)  B4(n),  B4(n+1),  B4(n+1),  B4(n+2)
+#   define B8(n)  B6(n),  B6(n+1),  B6(n+1),  B6(n+2)
[…]

### 2019-05-21 utils/macros.h

@@ -2,9 +2,13 @@
 
 #define NLATMAX           100
 
-#define BOOST_CELL_THRESH   10
+#define BOOST_CELL_THRESH    10
 #define BOOST_NUMPART_THRESH 250
-#define BOOST_BIN_REF       1
+#define BOOST_BIN_REF        1
+
+//what fraction of particles have to be sorted
+//to switch from quicksort to heapsort
+#define FRACTION_SORTED_REQD_TO_HEAP_SORT   0.6
 
 #define ADD_DIFF_TIME(t0,t1)            ((t1.tv_sec - t0.tv_sec) + 1e-6*(t1.tv_usec - t0.tv_usec))
 #define REALTIME_ELAPSED_NS(t0, t1)     ((t1.tv_sec - t0.tv_sec)*1000000000.0 + (t1.tv_nsec - t0.tv_nsec))
@@ -13,6 +17,10 @@
 
 #define STRINGIFY(x)   #x
 #define STR(x) STRINGIFY(x)
+
+
+#define MIN(X, Y)                        (( (X) < (Y)) ? (X):(Y))
+#define MAX(X, Y)                        (( (X) > (Y)) ? (X):(Y))
 
 
 #define ASSIGN_CELL_TIMINGS(thread_timings, nx1, nx2, timediff, tid, first_cellid, second_cellid) \
